### 2025-02-26 manuscript_GNM.qmd
Search and selection section done

---
title: "WORKING TITLE: The adaptive value of green nest material across birds: A systematic review and meta-analysis"
authors:
  - name: Shreya Dimri
    affiliation: Department of Evolutionary Biology, Bielefeld University, Bielefeld, Germany
    corresponding: true
  - name: Tuba Rizvi
    affiliation: Department of Evolutionary Biology, Bielefeld University, Bielefeld, Germany
  - name: Júlio Segovia
    affiliation: Department of Evolutionary Biology, Bielefeld University, Bielefeld, Germany
  - name: Alfredo Sánchez-Tójar
    affiliation: Department of Evolutionary Biology, Bielefeld University, Bielefeld, Germany

bibliography: references.bib
---

```{r Setup, include=FALSE, echo=FALSE}

pacman::p_load(readr)

```

# Methods (\~2200 words)

## Registration and Reporting

We pre-registered our study protocol on Open Science Framework [@dimri2024] prior to data exploration and analysis. The pre-registration details study objective, a priori hypothesis and predictions, methods and planned data analysis. Unless stated otherwise, we adhered to the registered protocol. We followed the PRISMA-EcoEvo guidelines [@odea2021] for reporting this study (in @figsupp-checklist). All resources associated with this study, including data and code, are available at \[<https://github.com/shreyadimri/Green_Nest_Material>\] (insert Zenodo citation later).

## Systematic Review

### Study selection and eligibility criteria

```{r load search data, include=FALSE, echo=FALSE}

unique_reference_list <- read_csv("data/unique_reference_list.csv")
full_reference_before_deduplication<-read_csv("data/full_reference_before_deduplication.csv")
selected_abstract_screening<-read_csv("data/selected_abstract_screening.csv")

# For updated search
unique_reference_list_repeat <- read_csv("data/unique_reference_repeat_12082024.csv")
full_reference_before_deduplication_repeat<-read_csv("data/full_reference_repeat_12082024.csv")
selected_abstract_screening_repeat<-read_csv("data/selected_abstract_screening_repeat.csv")

# For fulltext screening
excluded_fulltext <- read_csv("data/excluded_fulltext.csv")
```

We conducted a systematic search on Web of science Core Collection and Scopus citation database on 07.09.2022. The search query was designed in three blocks to included terms for predictor/treatment ("green\*" OR "herb\*" OR "aromatic\*") AND "nest\*"), population of interest ("bird\*" OR "aves" OR "avian" OR "ornithol\*" OR "passerine\*" OR "passeriform\*" OR "songbird\*" OR list of all bird genera) and only experimental studies ("experiment\*" OR "manipulat\*") (for complete search string tailored to specific database see @figsupp-string). We searched the title, abstract and keywords in English-language only with no restrictions on publication date or study type. We found `{r} nrow(unique_reference_list)` unique articles after removing `{r} nrow(full_reference_before_deduplication)-nrow(unique_reference_list)` duplicates using R package 'revtools' v0.4.1 (@westgate2019). We validated our search using an initial library of 15 key articles (\@figsupp-own-library), ensuring all were retrieved. Additionally, we used R package 'litsearchr' v1.0.0 (@grames2019) to confirm our keyword selection, which did not yield any new relevant terms.

All title and abstract screening were screened independently by atleast two screeners (\~66% each by SD, TR, JMGS) following a pre-designed decision tree (@figsupp-decision-tree) in revtools. We included all studies on birds that experimentally manipulated the green nest material (including aromatics or herbs). `{r} nrow(selected_abstract_screening)` articles passed title and abstract screening and further evaluated in a fulltext screening by a single screener (\~33% each by SD, TR, JMGS) and double checked by another. We used the same criteria as above with an additional evaluation of whether the variable of interest in the article was a fitness proxy. `{r} nrow(excluded_fulltext)` articles were excluded and all conflicts in screening decision (\~1% for title and abstract screening and none for fulltext screening) were resolved through discussion.

We repeated our search using the same search string on 12.08.2024 and found `{r} nrow(full_reference_before_deduplication_repeat)` additional articles. After deduplicating, SD screened the title and abstract for the `{r} nrow(unique_reference_list_repeat)` unique articles using the same decision tree. We found `{r} sum(selected_abstract_screening_repeat$screened_abstracts=="selected", na.rm=T)` article by updating our search. In addition, we identified `{r}  nrow(selected_abstract_screening_repeat)-(sum(selected_abstract_screening_repeat$screened_abstracts=="selected", na.rm=T))` articles through other sources that were not a part of our search. These studies met our inclusion criteria and were added to our meta-analysis. The exact number of screened and included studies along with reason of exclusion are shown in @figsupp-prisma-flowchart. We finally extracted data for `{r}  nrow(selected_abstract_screening_repeat) + (nrow(selected_abstract_screening)-nrow(excluded_fulltext))` articles and 2 unpublished studies. These were identified through our efforts to contact the authors of all included studies, and one dataset was provided by a co-author of this paper. <!--# Fix the code for generating this number later by using the data file directly to calculate the number of article -->

### Data Extraction

-   How many papers in final meta-analysis and how many effect sizes

<!-- -->

-   What did we extract in priority order

-   Wrote to authors for missing info, used methods to deal with missing info

-   Figures extraction

-   Moderators extracted

### Effect size calculation

### Meta-analysis and meta-regressions

-   main effect model

-   meta-regressions for testing hypothesis

-   meta-regressions for exploratory analyses

Senstivity analysis and Publication Bias Test

### Deviations from registration

## Supplementary Appendix

<details>

<summary>Supplementary materials</summary>

{{< include supplementary_appendix.qmd >}}

</details>