### 2025-03-17 docs/manuscript_GNM.qmd
Added a little more to methods

---
title: "WORKING TITLE: The adaptive value of green nest material across birds: A systematic review and meta-analysis"
authors:
  - name: Shreya Dimri
    affiliation: Department of Evolutionary Biology, Bielefeld University, Bielefeld, Germany
    corresponding: true
  - name: Tuba Rizvi
    affiliation: Department of Evolutionary Biology, Bielefeld University, Bielefeld, Germany
  - name: Júlio Segovia
    affiliation: Department of Evolutionary Biology, Bielefeld University, Bielefeld, Germany
  - name: Alfredo Sánchez-Tójar
    affiliation: Department of Evolutionary Biology, Bielefeld University, Bielefeld, Germany

bibliography: references.bib
---

```{r Setup, include=FALSE, echo=FALSE}

pacman::p_load(readr,tidyverse)

```

# Methods (\~2200-3800 words)

## Registration and Reporting

We pre-registered our study protocol on Open Science Framework [@dimri2024] prior to data exploration and analysis. The pre-registration details study objective, a priori hypothesis and predictions, methods and planned data analysis. Unless stated otherwise, we adhered to the registered protocol. We followed the PRISMA-EcoEvo guidelines [@odea2021] for reporting this study (in @figsupp-checklist). All resources associated with this study, including data and code, are available at \[<https://github.com/shreyadimri/Green_Nest_Material>\] (insert Zenodo citation later).

## Study selection and eligibility criteria

```{r load search data, include=FALSE, echo=FALSE}

unique_reference_list <- read_csv("data/unique_reference_list.csv")
full_reference_before_deduplication<-read_csv("data/full_reference_before_deduplication.csv")
selected_abstract_screening<-read_csv("data/selected_abstract_screening.csv")

# For updated search
unique_reference_list_repeat <- read_csv("data/unique_reference_repeat_12082024.csv")
full_reference_before_deduplication_repeat<-read_csv("data/full_reference_repeat_12082024.csv")
selected_abstract_screening_repeat<-read_csv("data/selected_abstract_screening_repeat.csv")

# For fulltext screening
excluded_fulltext <- read_csv("data/excluded_fulltext.csv")
```

We conducted a systematic search on Web of science Core Collection and Scopus citation database on 07.09.2022. The search query was designed in three blocks to included terms for predictor/treatment ("green\*" OR "herb\*" OR "aromatic\*") AND "nest\*"), population of interest ("bird\*" OR "aves" OR "avian" OR "ornithol\*" OR "passerine\*" OR "passeriform\*" OR "songbird\*" OR list of all bird genera) and only experimental studies ("experiment\*" OR "manipulat\*") (for complete search string tailored to specific database see @figsupp-string). We searched the title, abstract and keywords in English-language only with no restrictions on publication date or study type. We found `{r} nrow(unique_reference_list)` unique articles after removing `{r} nrow(full_reference_before_deduplication)-nrow(unique_reference_list)` duplicates using R package 'revtools' v0.4.1 (@westgate2019). We validated our search using an initial library of 15 relevant articles (\@figsupp-own-library), ensuring all were retrieved. Additionally, we used R package 'litsearchr' v1.0.0 (@grames2019) to confirm our keyword selection, which did not yield any new relevant terms.

All title and abstract screening were screened independently by atleast two screeners (\~66% each by SD, TR, JMGS) following a pre-designed decision tree (@figsupp-decision-tree) in revtools. We included all studies on birds that experimentally manipulated the green nest material (including aromatics or herbs). `{r} nrow(selected_abstract_screening)` articles passed title and abstract screening and further evaluated in a fulltext screening by a single screener (\~33% each by SD, TR, JMGS) and double checked by another. We used the same criteria as above with an additional evaluation of whether the variable of interest in the article was a fitness proxy. `{r} nrow(excluded_fulltext)` articles were excluded and all conflicts in screening decision (\~1% for title and abstract screening and none for fulltext screening) were resolved through discussion.

We repeated our search using the same search string on 12.08.2024 and found `{r} nrow(full_reference_before_deduplication_repeat)` additional articles. After deduplicating, SD screened the title and abstract for the `{r} nrow(unique_reference_list_repeat)` unique articles using the same decision tree. We found `{r} sum(selected_abstract_screening_repeat$screened_abstracts=="selected", na.rm=T)` article by updating our search. In addition, we identified `{r}  nrow(selected_abstract_screening_repeat)-(sum(selected_abstract_screening_repeat$screened_abstracts=="selected", na.rm=T))` articles through other sources that were not a part of our search. These studies met our inclusion criteria and were added to our meta-analysis. The exact number of screened and included studies along with reason of exclusion are shown in @figsupp-prisma-flowchart.

## Data Extraction

```{r load data extraction, include=FALSE, echo=FALSE}

merged_data_extraction <- read_csv("data/03_data_cleaning/merged_data_extraction.csv")

dataset_after_cleaning <- merged_data_extraction %>% filter(fulltext_screening!="excluded")
included_papers_no<-length(unique(dataset_after_cleaning$paper_ID))
```

### Data collection process

Data was extracted evenly by one of the three authors (\~33% each SD, TR and JMGS) and was checked by another to ensure correctness of the data. We extracted the primary data (i.e., measures of central tendency and variance) from text whenever possible. In cases where primary data was only available in figures (such as boxplots), we used 'metaDigitise' v.1.0.1 (@pick2018). If primary data was not available, we extracted any inferential statistics (t-test, F-statistics) that could be used to calculate an effect size. In cases where the values could not be extracted from the article but datasets were provided, we calculated the means, SD and sample sizes from the raw data.

To obtain any missing data and to locate any grey literature, we wrote to the corresponding author of the articles that passed fulltext screening using a standardized email template (\@figsupp-email). We obtained one unpublished dataset ... \[calculate estimates here\].. Another unpublished dataset was provided by a co-author of this paper. We extracted data for `{r} included_papers_no - 2` published articles and 2 unpublished studies.

### Extracted variables

We extracted mean, variance and sample size for a control-treatment pairwise comparison. Controls included non-aromatic materials (e.g., grass, moss), lesser amounts of aromatic material, or no material. Treatments consisted of aromatic materials added by authors, birds, or both. In cases comparing non-aromatic material to a blank nest, the non-aromatic material was considered the treatment. We recorded the type of comparison being made as a moderator. Each pairwise comparison was given a unique Observation ID, a paper ID (for each article reporting this data), an experiment ID (for estimates coming from the same experimental setup), a group ID (for estimates from the same group of individual) and a repeated trait ID (for estimates that were measured repeatedly across time).

For each study we also extracted information related to the study subjects (species of birds, location of the study population), the methodology of the study (plant species used as treatment and control, trait measured) and information about the extracted data (location or source of the data recorded). We categorized the trait measured into 6 broad categories - physiology, morphology, reproduction, behaviour, parasite and pathogenic load, phenology. In cases where the trait measured was a type of parasite or pathogen, we categorised the parasite studied into arthropods or micro-organisms. For a complete list of variables extracted see @figsupp-variables .

## Effect size calculation

```{r load effect size, include=FALSE, echo=FALSE}

inferential_stats<-dataset_after_cleaning%>%filter(statistics_type=="t-test"|statistics_type=="ANCOVA")

contingency_table<-dataset_after_cleaning%>%filter(proxy_decision=="contingency table")

no_ES<-dataset_after_cleaning%>%filter(proxy_comment == "use 0 as ES")

negative_lnRR<-dataset_after_cleaning%>%
  filter(measure_central_tendency_experiment<=0 | measure_central_tendency_control<=0)

```

We calculated two effect sizes for each pairwise comparision-log response ratio (lnRR) and standardised mean difference with heteroscedasticity correction (SMDH), using `escalc` function (`measure= ROM` and `SMDH` respectively, `vtype= "LS"`) in `metafor` package (@metafor). We specified treatment group as the numerator and the control group as the denominator for lnRR and treatment minus control for SMDH, such that the positive values indicate the trait value increased in the treatment group. In cases such as measures of parasite load or prevalence, laying or hatching dates, courtship time, scab score etc., the sign was inverted to ensure comparability across all fitness proxies. In all cases, the effect size were coded such that the fitness proxies were positively correlated with the overall fitness.

For studies reporting only inferential statistics (e.g., t-tests, ANCOVA), we estimated SMD values (n=`{r} nrow(inferential_stats)`, k=`{r} length(unique(inferential_stats$paper_ID))`) using the equations provided by @thehand2019 . In cases of dichotomous variables (e.g., successful nests, n=`{r} nrow(contingency_table)`, k=`{r} length(unique(contingency_table$paper_ID))`) we used the data to calculate log odds ratio using `escalc` function (`measure=OR`) and convert it to SMD using equations from @thehand2019. Since lnRR cannot be calculated from inferential statistics or dichotomous variables, only SMD under the assumption of homoskedasticity was derived from such data. We excluded these effect sizes along with any zero or negative values (n=`{r} nrow(negative_lnRR)`, k=`{r} length(unique(negative_lnRR$paper_ID))`) when calculating lnRR.

The sample sizes used was the number of nests in each group, and not the number of individuals, whenever possible. In cases where the control or treatment group was used repeatedly across multiple comparisons, sample sizes were adjusted by dividing them by the number of times each group was used in the comparison, to account for shared group non-independence. For the cases where the authors reported no significant difference between the two groups (n=`{r} nrow(no_ES)`, k=`{r} length(unique(no_ES$paper_ID))`) but did not provide the values, we used zero as the effect size. To impute the SD for these cases, we used the missing case approach following @nakagawa2022.

## Meta-analysis

We used rma.mv function from metafor pacjage using 'REML' method to fit all the models. To estimate the overall effect of green nest material addition to the nest on fitness across bird species that perform this behaviour, we fitted two separate multilevel intercept-only meta-analytical using lnRR or SMDH as dependent variables. Sampling variances were modelled as variance-covariance matrices assuming a 0.5 correlation (ρ) between sampling variances from the same study [@noble2017].

*To understand the sources of variation, we calculated Q and total heterogeneity (𝜎2), its source (or relative heterogeneity; I2) and two different metrics to understand its magnitude – a mean-standardized metric (CV) and a variance-mean-standardized metric (M; more in* @yang2025*).*

*To understand the sources of variation or heterogeneity, for the intercept-only models, we will use the pluralistic approach recently suggested by which consists of calculating four metrics of heterogeneity (σ2, I2, CV, M) to comprehensively understand both total and random-effect specific heterogeneity. For the meta-regressions, we will calculate Rmarginal2 to quantify the heterogeneity explained by the moderator(s) in proportion to the total variation (Nakagawa & Schielzeth, 2013).*

## Meta-regression for testing hypothesis

For understanding the effect of green nest material on reproductive success (courtship hypothesis) and on protective effects and health benefits for the nestling (Parental care hypothesis), we ran a multilevel meta-regression with 'hypothesis' (Courtship hypothesis, Parental care hypothesis, Both) as a moderator.

*For meta-regression models, we estimated the percentage of heterogeneity explained by the moderators using R2marginal (Nakagawa & Schielzeth 2013). We graphically presented the results as orchard plots using the R package “orchaRd” v.2.0 (Nakagawa et al. 2023).*

## Meta-regression for exploratory analyses

## Publication Bias Test

Publication bias was assessed through multilevel meta-regressions maintaining the same random-effects structure as above. Following Nakagawa et al. (2022), we executed the following steps:

1.  A uni-moderator meta-regression to detect small-study effects for both lnRR and SMDH, using the square root of the inverse effective sample size as the moderator.

2.  A subsequent uni-moderator meta-regression using the inverse effective sample size to provide a less biased adjusted overall effect.

3.  A time-lag bias analysis using publication year (mean-centered) as the moderator.

4.  A full meta-regression model incorporating small-study effects, time-lag bias, and all other moderators to assess combined heterogeneity sources.

Heterogeneity was further explored using variance decomposition and random-effects modeling, including species and phylogeny if additional effect sizes warranted phylogenetic non-independence correction (Cinar et al., 2022).

## Deviations from registration

**Sensitivity Analyses**

To ensure robustness, we conducted additional sensitivity analyses:

1.  Repeating meta-analyses and meta-regressions using only effect sizes derived from means, SDs, and sample sizes, excluding those based on inferential statistics.

2.  Implementing a bivariate multilevel meta-analytic model (Yang et al., 2024) incorporating both lnRR and SMDH as response variables.

3.  Re-running intercept-only meta-analyses with different sampling variance correlations (ρ = 0.3 and 0.7) and extending these sensitivity analyses to meta-regressions if necessary.

Non-pre-registered sensitivity analyses that emerged during data processing were reported in supplementary materials.

## Stuff that needs to go somewhere

# Results

We obtained X effect sizes from k studies including How many papers in final meta-analysis and how many effect size.

## Supplementary Appendix

<details>

<summary>Supplementary materials</summary>

{{< include supplementary_appendix.qmd >}}

</details>