### 2025-04-22 manuscript_GNM.qmd
Results: Intercept only model - overall effect

---
title: "WORKING TITLE: The adaptive value of green nest material across birds: A systematic review and meta-analysis"
authors:
  - name: Shreya Dimri
    affiliation: Department of Evolutionary Biology, Bielefeld University, Bielefeld, Germany
    corresponding: true
  - name: Tuba Rizvi
    affiliation: Department of Evolutionary Biology, Bielefeld University, Bielefeld, Germany
  - name: Júlio Segovia
    affiliation: Department of Evolutionary Biology, Bielefeld University, Bielefeld, Germany
  - name: Alfredo Sánchez-Tójar
    affiliation: Department of Evolutionary Biology, Bielefeld University, Bielefeld, Germany

bibliography: references.bib
---

```{r Setup, include=FALSE, echo=FALSE}
devtools::install_github("daniel1noble/orchaRd", ref = "main", force = TRUE)
pacman::p_load(readr,tidyverse,metafor,orchaRd)

```

```{r load search data, include=FALSE, echo=FALSE, message= FALSE}

# Own reference list
own_library <- read_csv("data/01_systematic_search/01_search_strategy/own_library_references.csv")

# Systematic search and abstract screening
unique_reference_list <- read_csv("data/unique_reference_list.csv")
full_reference_before_deduplication<-read_csv("data/full_reference_before_deduplication.csv")
selected_abstract_screening<-read_csv("data/selected_abstract_screening.csv")

# For updated search
unique_reference_list_repeat <- read_csv("data/unique_reference_repeat_12082024.csv")
full_reference_before_deduplication_repeat<-read_csv("data/full_reference_repeat_12082024.csv")
selected_abstract_screening_repeat<-read_csv("data/selected_abstract_screening_repeat.csv")

# For fulltext screening
excluded_fulltext <- read_csv("data/excluded_fulltext.csv")
```

# Methods

## Registration and Reporting

We pre-registered our study protocol on Open Science Framework [@dimri2024] prior to data exploration and analysis. The pre-registration details the study's objectives, a priori hypotheses and predictions, methods and planned data analyses. Unless stated otherwise, we adhered to the registered protocol. We followed the PRISMA-EcoEvo guidelines [@odea2021] for reporting this study (@figsupp-checklist). All resources associated with this study, including data and code, are available at \[<https://github.com/shreyadimri/Green_Nest_Material>\] (insert Zenodo citation later).

## Information source and search

We conducted a systematic search on Web of Science Core Collection and Scopus on 07.09.2022. The keyword search query was designed in three blocks to included terms for predictor/treatment ("green\*" OR "herb\*" OR "aromatic\*") AND "nest\*"), population of interest ("bird\*" OR "aves" OR "avian" OR "ornithol\*" OR "passerine\*" OR "passeriform\*" OR "songbird\*" OR list of all bird genera) and only experimental studies ("experiment\*" OR "manipulat\*") (for a complete search string tailored to specific database see @figsupp-string). We searched the title, abstract and keywords in English-language only, with no restrictions on publication date or study type. We found `{r} nrow(unique_reference_list)` unique articles after removing `{r} nrow(full_reference_before_deduplication)-nrow(unique_reference_list)` duplicates using R package 'revtools' v0.4.1 (@westgate2019). We validated our search using an initial library of `{r} nrow(own_library)` relevant articles (@figsupp-own-library), ensuring all were retrieved. Additionally, we used R package 'litsearchr' v1.0.0 (@grames2019) to confirm our keyword selection, which did not yield any new relevant terms.

## Eligibility criteria

All title and abstract screening were screened independently by atleast two screeners (\~66% each by SD, TR, JMGS) following a pre-designed decision tree (@figsupp-decision-tree) in `revtools`. We included all studies on birds that experimentally manipulated the green nest material (including aromatics or herbs). `{r} nrow(selected_abstract_screening)` articles passed title and abstract screening and further evaluated in a fulltext screening by a single screener (\~33% each by SD, TR, JMGS) and double checked by another. We used the same criteria as above with an additional evaluation of whether the variable of interest in the article was a fitness proxy. `{r} nrow(excluded_fulltext)` articles were excluded and all conflicts in screening decisions (\~1% disagreement for title and abstract screening and none for fulltext screening) were resolved through discussion.

We repeated our search using the same search string on 12.08.2024 and found `{r} nrow(full_reference_before_deduplication_repeat)` additional articles. After de-duplicating, SD screened the title and abstract for the `{r} nrow(unique_reference_list_repeat)` unique articles using the same decision tree. We found `{r} sum(selected_abstract_screening_repeat$screened_abstracts=="selected", na.rm=T)` article by updating our search. In addition, we identified `{r} nrow(selected_abstract_screening_repeat)-(sum(selected_abstract_screening_repeat$screened_abstracts=="selected", na.rm=T))` articles through other sources that were not a part of our search. These studies met our inclusion criteria and were added to our meta-analysis. The exact number of screened and included studies along with reasons of exclusion are shown in @figsupp-prisma-flowchart.

## Data Extraction

```{r load data extraction, include=FALSE, echo=FALSE}

merged_data_extraction <- read_csv("data/03_data_cleaning/merged_data_extraction.csv", show_col_types = F)

dataset_after_cleaning <- merged_data_extraction %>% filter(fulltext_screening!="excluded")
included_papers_no<-length(unique(dataset_after_cleaning$paper_ID))

# Numbers for extraction ID
data_extractors <- merged_data_extraction %>%
     distinct(paper_ID, .keep_all = TRUE) %>%
     select(paper_ID, extractor_ID) %>%
     mutate(extractor_ID = case_when(
    extractor_ID == "JS/SD/TR" ~ "TR/JMGS/SD",
    TRUE ~ extractor_ID)) %>%
     group_by(extractor_ID) %>%
     summarise(total_papers = n(), .groups = "drop")

excluded_proxies<- dataset_after_cleaning %>% filter(proxy_decision=="exclude")
SAonly_proxies<- dataset_after_cleaning %>% filter(proxy_decision=="senstivity-analysis" | proxy_decision=="sensitivity-analysis")
```

Data was extracted evenly by one of the three authors (\~33% each SD, TR and JMGS) and was checked by another to ensure correctness of the data. We extracted the primary data (i.e., measures of central tendency and variance) from text whenever possible. In cases where primary data was only available in figures (such as boxplots), we used `metaDigitise` v.1.0.1 (@pick2018). If primary data was not available, we extracted any inferential statistics (t-test, F-statistics) that could be used to calculate an effect size. In cases where the values could not be extracted from the article but datasets were provided, we calculated the means, SD and sample sizes from the raw data. When data was presented in different sources, we prioritized the source having the maximum resolution.

To obtain any missing data and to locate any grey literature, we wrote to the corresponding author of the articles that passed fulltext screening using a standardized email template (\@figsupp-email). We contacted 4 authors for missing data, while the rest were asked for any potential unpublished data. All four authors responded to our email- 2 provided the data as requested, one did not have the data any more and one did not give a conclusive response. One author provided unpublished data. Another unpublished data was provided by the author of this study (MO). We extracted data for `{r} included_papers_no - 2` published articles and 2 unpublished studies.

We extracted mean, variance and sample size for a control-treatment pairwise comparison. Controls included non-aromatic materials (e.g., grass, moss), lesser amounts of aromatic material, or no material. Treatments consisted of aromatic materials added by authors, birds, or both. In cases comparing non-aromatic material to a blank nest, the non-aromatic material was considered the treatment. We recorded the type of comparison being made as a moderator.

For each study we also extracted information related to the study subjects (species of birds, location of the study population), the methodology of the study (plant species used as treatment and control, trait measured) and information about the extracted data (location or source of the data recorded). We categorized the trait measured into 6 broad categories - physiology, morphology, reproduction, behaviour, parasite and pathogenic load, phenology. In cases where the trait measured was a type of parasite or pathogen, we categorised the parasite studied into arthropods or micro-organisms. For a complete list of variables extracted see @figsupp-variables . We excluded some traits (k=`{r} nrow(excluded_proxies)` , n=`{r} length(unique(excluded_proxies$paper_ID))`) measured from all our analysis because it was unclear how they related to fitness (see @figsupp-excluded-proxies for the list and reason of exclusion). We also had to exclude some fitness proxies (n=?, k=?) because of the low sample size per group (less than x).

## Effect size calculation

```{r load effect size, include=FALSE, echo=FALSE}

inferential_stats<-dataset_after_cleaning%>%filter(statistics_type=="t-test"|statistics_type=="ANCOVA")

contingency_table<-dataset_after_cleaning%>%filter(proxy_decision=="contingency table")

no_ES<-dataset_after_cleaning%>%filter(proxy_comment == "use 0 as ES")

negative_lnRR<-dataset_after_cleaning%>%
  filter(measure_central_tendency_experiment<=0 | measure_central_tendency_control<=0)

```

We calculated two effect sizes for each pairwise comparison-log response ratio (lnRR) and standardised mean difference with heteroskedasticity correction (SMDH), using `escalc` function (`measure= ROM` and `SMDH` respectively, `vtype= "LS"`) in `metafor` package (@metafor). We specified the treatment group as the numerator and the control group as the denominator for lnRR and treatment minus control for SMDH, such that the positive values indicate the trait value increased in the treatment group. In cases such as measures of parasite load or prevalence, laying or hatching dates, courtship time, scab score etc., the sign was inverted to ensure comparability across all fitness proxies. In all cases, the effect sizes were coded such that the fitness proxies were positively correlated with the overall fitness.

For studies reporting only inferential statistics (e.g., t-tests, ANCOVA), we estimated SMD values (k=`{r} nrow(inferential_stats)`, n=`{r} length(unique(inferential_stats$paper_ID))`) using the equations provided by @thehand2019 . In cases of dichotomous variables (e.g., successful nests, k=`{r} nrow(contingency_table)`, n=`{r} length(unique(contingency_table$paper_ID))`) we used the data to calculate log odds ratio using `escalc` function (`measure=OR`) and convert it to SMD using equations from @thehand2019. Since lnRR cannot be calculated from inferential statistics or dichotomous variables, only SMD under the assumption of homoscedasticity was derived from such data. We excluded these effect sizes along with any zero or negative values (k=`{r} nrow(negative_lnRR)`, n=`{r} length(unique(negative_lnRR$paper_ID))`) when calculating lnRR.

The sample sizes used were the number of nests in each group, and not the number of individuals, whenever possible. In cases where the control or treatment group was used repeatedly across multiple comparisons, sample sizes were adjusted by dividing them by the number of times each group was used in the comparison, to account for shared group non-independence. For the cases where the authors reported no significant difference between the two groups (n=`{r} nrow(no_ES)`, k=`{r} length(unique(no_ES$paper_ID))`) but did not provide the values, we used zero as the effect size. To impute the SD for these cases, we used the missing-case approach following @nakagawa2022.

## Data Analysis

We performed analysis in R version 4.3.0 (@positteam2023). All multilevel meta-analytical models and meta-regressions were fit using `metafor v.4.8-0` (@metafor). Sampling variances were modelled as variance-covariance matrices assuming a 0.5 correlation (ρ) between sampling variances from the same study [@noble2017] using `vcalc` function. (See Supplementary Appendix @figsupp-correlation for sensitivity analysis using different values of correlation (ρ)).

Our data had various sources of non-independence which we accounted for using random effects. We considered the following level of non-independence- (a) multiple effect sizes were collected from same study (Paper ID), (b) some bird species (species ID) and (c) some geographical locations (population ID) were studied multiple times, (d) some sub-populations (Experiment ID e.g. birds studied at the same location for multiple years or birds studied at the same location separated by distance) and (e) group of individuals (Group ID, eg. males and females in a population, first clutch and second clutch) were represented in our data several times and (f) repeated traits collected on the same group of animals (repeated trait ID e.g. chick mass on day 7, 14 and 21).

We fitted the following random effects in all our models unless stated otherwise- Paper ID, Group ID, Species ID, Population ID and an observation-level random effect (Observation ID). Initially we had pre-registered the inclusion of Experiment ID and repeated trait ID and not population ID. Experiment ID and group ID were almost indistinguishable, as such we decided to leave out the one with lesser resolution i.e. Experiment ID. Additionally, we observed that not many effect sizes were measured repeatedly, making the repeated trait ID very similar to the observation ID. After exploring the overlap (@figsupp-random-effect), we decided to only keep observation ID. We added population ID, which accounts for many effect sizes from different studies coming from the same geographical location.

### Meta-analytical models for biological predictions

To estimate the overall effect of green nest material addition to the nest on fitness across bird species that perform this behaviour, we fitted two separate multilevel intercept-only meta-analytical models using lnRR or SMDH as dependent variables. We used the `rma.mv` function from the `metafor` package using the 'REML' method. Test statistics and confidence intervals for the fixed effects were computed using t distribution. We used the entire dataset for this analysis except some traits (k=`{r} nrow(SAonly_proxies)` , n=`{r} length(unique(SAonly_proxies$paper_ID))`; see @figsupp-SAonly_proxies for details) that we decided to exclude from all our analysis due to lack of evidence supporting their relationship with fitness. However, we also conducted a sensitivity analysis (@figsupp-MLMA-excluded) including these proxies to ensure robustness of our results. In the model using SMDH, We estimated total amount of heterogeneity ($𝜎^2$; the variation not explained by the sampling error), its sources (or relative heterogeneity; $I^2$), mean-standardised metric ($CVH2$) and variance-mean-standardised metric of heterogeneity ($M2$) (@yang2025).

To test the effect of green nest material on reproductive success (courtship hypothesis) and on protective effects and health benefits for the nestling (Parental care hypothesis), we ran a multilevel meta-regression with 'hypothesis' (Courtship hypothesis, Parental care hypothesis, Both) as a moderator. using the entire dataset. For this and all meta-regressions, we estimated $R_{marginal}^2$, the percentage of heterogeneity explained by the moderators (@nakagawa2012) using the entire dataset.

### Meta-regressions for exploratory analysis

We performed additional pre-registered multi-level meta-regressions for which we did not have clear or strong directional predictions on specific subsets of data. To understand the effect of green nest material on the parasite load depending on the type of parasites, we used effect sizes on parasites and pathogens (n=?, k=?). We ran a meta-regression with a type of parasite as a moderator (levels: arthropod, micro-organism). We also ran a meta-regression with with the entire dataset set using bird species as a moderator (levels: Cyanistes caeruleus, Sturnus unicolor, Tachycineta bicolor, Sturnus vulgaris, Parus major, Passer cinnamomeus, Buteo buteo) to understand if the effect of green nest material on fitness depended on the system being studied. To test the effect of time of addition of green nest material on the fitness outcome (levels: before egg hatching, after egg hatching, continuously throughout the nesting phase) we ran a meta-regression with Time of addition of green nest material as the moderator. We explored the effect of green nest material on fitness depending on the type of trait studies (levels:physiology, morphology, reproduction, behaviour, parasite and pathogenic load, phenology) using a meta-regression with the trait type as a moderator. We explored the impact of methodological differences on the relationship between green nest material and fitness. To understand the effect of type of experimental design (i.e. whether the authors compared - non-aromatic vs. aromatic, no added material vs. aromatic, no added material vs. non-aromatic), we ran a meta-regression with type of experimental design as the moderator.

### Publication Bias Test

Publication bias is the higher likelihood of a subset of research findings, such as statistically significant results, being published as compared to the rest. We tested for evidence of publication bias (following @nakagawa2021) for both lnRR and SMDH separately. To test for small study effects, we ran meta-regression , using the inverse of effective sample size as a moderator. To understand the decline effect (also known as time-lag bias) we ran another meta-regression with a mean-centered publication year as a moderator.

# Results

```{r}
dataset_analysis <- read_csv("data/03_data_cleaning/dataset_analysis.csv")

ES_per_study<-dataset_analysis%>%group_by(paper_ID)%>%count(paper_ID)

birds_info<-dataset_analysis%>%group_by(bird_species)%>%count(bird_species)
birds_info$perc<-(birds_info$n/nrow(dataset_analysis))*100
```

We obtained `{r} nrow(dataset_analysis)` effect sizes from `{r} length(unique(dataset_analysis$paper_ID))` studies, published between `{r} min((dataset_analysis$year_publication),na.rm = T)`-`{r} max((dataset_analysis$year_publication),na.rm = T)`. The mean (±SD) number of effect sizes per study was `{r} as.numeric(format(mean(ES_per_study$n), digit=3))` ± `{r} as.numeric(format(sd(ES_per_study$n), digit=3))`, with a range of `{r} min(ES_per_study$n)`--`{r} max(ES_per_study$n)` effect sizes per study. Median and mean sample sizes per effect size were `{r} median(c(dataset_analysis$effective_n_control,dataset_analysis$effective_n_experiment))` and `{r} as.numeric(format(mean(c(dataset_analysis$effective_n_control,dataset_analysis$effective_n_experiment)), digit=3))` respectively (range: `{r} min(c(dataset_analysis$effective_n_control,dataset_analysis$effective_n_experiment))`-- `{r} max(c(dataset_analysis$effective_n_control,dataset_analysis$effective_n_experiment))`; `{r} nrow(dataset_analysis%>%select(paper_ID,effective_n_control,effective_n_experiment, lnRR, SMDH)%>%filter(effective_n_experiment<2|effective_n_control<2))` effect sizes had a sample size less than 2 in each group). The studies were conducted on `{r} length(unique(dataset_analysis$bird_species))` species across `{r} length(unique(dataset_analysis$population_ID))` locations. About `{r} round(sum(birds_info$perc[birds_info$perc>15]))`% of the effect sizes come from three bird species (`{r} birds_info$bird_species[birds_info$perc > 15][1]`, k = `{r} birds_info$n[birds_info$perc > 15][1]`; `{r} birds_info$bird_species[birds_info$perc > 15][2]`, k = `{r} birds_info$n[birds_info$perc > 15][2]`; `{r} birds_info$bird_species[birds_info$perc > 15][3]`, k = `{r} birds_info$n[birds_info$perc > 15][3]`).

**Does addition of green nest material have an adaptive function?**

```{r}
intercept_lnRR<-readRDS(here::here("model/intercept_lnRR.rds"))
intercept_SMDH<-readRDS(here::here("model/intercept_SMDH.rds"))
```

The intercept only meta-analytic model using the log response ratio (lnRR) estimated a `{r} round((exp(as.double(coef(intercept_lnRR)))-1)*100,1)`% increase in fitness in nests with green nest material across bird species \[Figure ; lnRR = `{r} as.double(formatC(coef(intercept_lnRR), digits = 3, format = "f"))`, 95% CI: \[`{r} as.double(formatC((intercept_lnRR)$ci.lb, digits = 3, format = "f"))`, `{r} as.double(formatC((intercept_lnRR)$ci.ub, digits = 3, format = "f"))`\], 95% PI: \[`{r} as.double(formatC(predict(intercept_lnRR)$pi.lb, digits = 3, format = "f"))`, `{r} as.double(formatC(predict(intercept_lnRR)$pi.ub, digits = 3, format = "f"))`\], p-value = `{r} as.double(formatC(intercept_lnRR$pval, digits = 2, format = "f"))`, k = `{r} intercept_lnRR$k`, n = `{r} length(intercept_lnRR$s.levels.f[[1]])`\]. The total heterogeneity (𝜎2) was estimated at `{r} round(sum(intercept_lnRR$sigma2), 3)` (Q = `{r} round(sum(intercept_lnRR$QE),3)`, p \< 0.001). Measures of relative heterogeneity (I^2^~total~ = `{r} round(as.double(orchaRd::i2_ml(intercept_lnRR)[1]),1)`%) and standardized heterogeneity (mean-standardized: CV~total~ = `{r} round(as.double(orchaRd::cvh1_ml(intercept_lnRR)[1]),2)` and variance-mean-standardized: M~total =~ `{r} round(as.double(orchaRd::m1_ml(intercept_lnRR)[1]),2)`) suggested high variability across effect sizes. Using standardized mean difference (SMDH) as an alternate effect size, the intercept only meta-analytic model estimated a small positive but statistically significant effect of green nest material on fitness \[Figure ; SMDH = `{r} as.double(formatC(coef(intercept_SMDH), digits = 3, format = "f"))`, 95% CI: \[`{r} as.double(formatC((intercept_SMDH)$ci.lb, digits = 3, format = "f"))`, `{r} as.double(formatC((intercept_SMDH)$ci.ub, digits = 3, format = "f"))`\], 95% PI: \[`{r} as.double(formatC(predict(intercept_SMDH)$pi.lb, digits = 3, format = "f"))`, `{r} as.double(formatC(predict(intercept_SMDH)$pi.ub, digits = 3, format = "f"))`\], p-value = `{r} as.double(formatC(intercept_SMDH$pval, digits = 3, format = "f"))`, k = `{r} intercept_SMDH$k`, n = `{r} length(intercept_SMDH$s.levels.f[[1]])`\]. Total heterogeneity was higher in case of SMDH (𝜎^2^= `{r} round(sum(intercept_SMDH$sigma2), 3)` Q = `{r} round(sum(intercept_SMDH$QE),3)`, p \< 0.001). and the relative heterogeneity was substantially reduced (I^2^~total~ = `{r} round(as.double(orchaRd::i2_ml(intercept_SMDH)[1]),1)`%). The standardized heterogeneity (CV~total~ = `{r} round(as.double(orchaRd::cvh1_ml(intercept_SMDH)[1]),2)` and M~total =~ `{r} round(as.double(orchaRd::m1_ml(intercept_SMDH)[1]),2)`) remained high. While lnRR and SMDH differ in their scale, both effect sizes suggest a small but positive effect of green nest material on fitness. These differences in overall effect can stem from the different underlying biological and statistical assumptions. lnRR incorporates the mean difference at a multiplicative scale whereas SMDH expresses the difference on an additive scale. The difference in relative heterogeneity between the two effect sizes aligns with broader trends observed in meta-analysis in ecology and evolution (see @yang2024 for more details).

```{r}
orchaRd::caterpillars(intercept_lnRR, mod = "1", group="paper_ID",xlab = "Effect size (Log Response Ratio - lnRR)")

orchaRd::caterpillars(intercept_SMDH, mod = "1", group="paper_ID",xlab = "Effect size (Standardized Mean Difference - SMDH)")

intercept_lnRR_plot <- orchard_plot(intercept_lnRR,
                                         mod = "1",
                                         group = "paper_ID",
                                         trunk.size = 1,
                                         branch.size = 2,
                                         alpha = 0.3,
                                         twig.size = 1,
                                         xlab = "Effect size (Log Response Ratio - lnRR)",
                                         colour = "#218833",
                                         fill = T) +
  scale_fill_manual(values = "#218833") +
  scale_colour_manual(values = "#218833") +  
  theme(legend.direction="horizontal", legend.title = element_text(size =8),
        legend.text = element_text(size = 10), 
        axis.title = element_text(size = 15),
        axis.text.x = element_text(size = 15),
        axis.text.y = element_blank()) +
  labs(x = "Intercept")

print(intercept_lnRR_plot)

intercept_SMDH_plot <- orchard_plot(intercept_SMDH,
                                         mod = "1",
                                         group = "paper_ID",
                                         trunk.size = 1,
                                         branch.size = 2,
                                         alpha = 0.3,
                                         twig.size = 1,
                                         xlab = "Effect size (Standardized Mean Difference - SMDH)",
                                         colour = "#895129",
                                         fill = T) +
  scale_fill_manual(values = "#895129") +
  scale_colour_manual(values = "#895129") +  
  theme(legend.direction="horizontal", legend.title = element_text(size =8),
        legend.text = element_text(size = 10), 
        axis.title = element_text(size = 15),
        axis.text.x = element_text(size = 15),
        axis.text.y = element_blank()) +
  labs(x = "Intercept")

print(intercept_SMDH_plot)
```

\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\--

## Supplementary Appendix

<details>

<summary>Supplementary materials</summary>

{{< include supplementary_appendix.qmd >}}

</details>