### 2025-05-08 docs/manuscript_GNM.qmd
Incorporated changes for methods

---
title: "WORKING TITLE: The adaptive value of green nest material across birds: A systematic review and meta-analysis"
authors:
  - name: Shreya Dimri
    affiliation: Department of Evolutionary Biology, Bielefeld University, Bielefeld, Germany
    corresponding: true
  - name: Tuba Rizvi
    affiliation: Department of Evolutionary Biology, Bielefeld University, Bielefeld, Germany
  - name: Júlio Segovia
    affiliation: Department of Evolutionary Biology, Bielefeld University, Bielefeld, Germany
  - name: Alfredo Sánchez-Tójar
    affiliation: Department of Evolutionary Biology, Bielefeld University, Bielefeld, Germany

bibliography: references.bib
---

```{r Setup, include=FALSE, echo=FALSE}
devtools::install_github("daniel1noble/orchaRd", ref = "main", force = TRUE)
pacman::p_load(readr,tidyverse,metafor,orchaRd,patchwork)

```

```{r load search data, include=FALSE, echo=FALSE, message= FALSE}

# Own reference list
own_library <- read_csv("data/01_systematic_search/01_search_strategy/own_library_references.csv")

# Systematic search and abstract screening
unique_reference_list <- read_csv("data/unique_reference_list.csv")
full_reference_before_deduplication<-read_csv("data/full_reference_before_deduplication.csv")
selected_abstract_screening<-read_csv("data/selected_abstract_screening.csv")

# For updated search
unique_reference_list_repeat <- read_csv("data/unique_reference_repeat_12082024.csv")
full_reference_before_deduplication_repeat<-read_csv("data/full_reference_repeat_12082024.csv")
selected_abstract_screening_repeat<-read_csv("data/selected_abstract_screening_repeat.csv")

# For fulltext screening
excluded_fulltext <- read_csv("data/excluded_fulltext.csv")
```

# Methods

## Registration and Reporting

We pre-registered our study protocol on Open Science Framework [@dimri2024] prior to data exploration and analysis. The pre-registration details the study's objectives, a priori hypotheses and predictions, methods and planned data analyses. Unless stated otherwise, we adhered to the registered protocol. We followed the PRISMA-EcoEvo guidelines [@odea2021] for reporting this study (@figsupp-checklist). All resources associated with this study, including data and code, are available at \[<https://github.com/shreyadimri/Green_Nest_Material>\] (insert Zenodo citation later).

## Information source and search

We conducted a systematic search on Web of Science Core Collection and Scopus on 07.09.2022. The keyword search query was designed as a three-block query to included terms related to the predictor/treatment (("green\*" OR "herb\*" OR "aromatic\*") AND "nest\*"), our population of interest ("bird\*" OR "aves" OR "avian" OR "ornithol\*" OR "passerine\*" OR "passeriform\*" OR "songbird\*" OR list of all bird genera) and to limit it to only experimental studies only ("experiment\*" OR "manipulat\*"). The complete search strings tailored to each specific platform are provided in @figsupp-string. We searched through titles, abstracts and keywords in English-language only with no restrictions on publication date or study type. We validated our search using an initial library of `{r} nrow(own_library)` relevant articles that we had previously identified (@figsupp-own-library), ensuring all were retrieved. Additionally, we used the R package 'litsearchr' v1.0.0 (@grames2019) to confirm our keyword selection, which did not yield any new relevant terms. We found `{r} nrow(unique_reference_list)` unique articles after removing `{r} nrow(full_reference_before_deduplication)-nrow(unique_reference_list)` duplicates using R package 'revtools' v0.4.1 (@westgate2019).

## Eligibility criteria

All titles and abstracts were equally allocated and screened independently by atleast two screeners (SD, TR, JMGS) using the R package `revtools` following a pre-designed decision tree (@figsupp-decision-tree). We included all studies on birds that experimentally manipulated the green nest material (both aromatics or herbs). `{r} nrow(selected_abstract_screening)` articles passed the title-and-abstract screening and were further evaluated during the full-text screening. Each article was screened by a single screener (SD, TR, JMGS), with another screener verifying the procedure. We used the same criteria as above (see @figsupp-decision-tree) with an additional evaluation of whether the variable of interest in the article was in our list of fitness proxies of interest (see @figsupp-fitness-proxies). `{r} nrow(excluded_fulltext)` articles were excluded. All conflicts in screening decisions (\~1% disagreement for title-and-abstract screening, and none for full-text screening) were resolved through discussion.

To update our systematic lietrature search, we repeated our search on 12.08.2024. This updated search found `{r} nrow(unique_reference_list_repeat)` unique articles after removing ``` {r} nrow(full_reference_before_deduplication_repeat)-``nrow(unique_reference_list_repeat) ``` duplicates. Only one observer (SD) performed the title-and-abstract and full-text screening using the same decision trees as before and found `{r} sum(selected_abstract_screening_repeat$screened_abstracts=="selected", na.rm=T)` additional eligible article. Lastly, we identified `{r} nrow(selected_abstract_screening_repeat)-(sum(selected_abstract_screening_repeat$screened_abstracts=="selected", na.rm=T))` additional eligible articles. In all, X <!--# Fill the no. of articles that are included --> articles were eligible for our meta-analysis (more details in @figsupp-prisma-flowchart).

## Data Extraction

```{r load data extraction, include=FALSE, echo=FALSE}

merged_data_extraction <- read_csv("data/03_data_cleaning/merged_data_extraction.csv", show_col_types = F)

dataset_after_cleaning <- merged_data_extraction %>% filter(fulltext_screening!="excluded")
included_papers_no<-length(unique(dataset_after_cleaning$paper_ID))

# Numbers for extraction ID
data_extractors <- merged_data_extraction %>%
     distinct(paper_ID, .keep_all = TRUE) %>%
     select(paper_ID, extractor_ID) %>%
     mutate(extractor_ID = case_when(
    extractor_ID == "JS/SD/TR" ~ "TR/JMGS/SD",
    TRUE ~ extractor_ID)) %>%
     group_by(extractor_ID) %>%
     summarise(total_papers = n(), .groups = "drop")

excluded_proxies<- dataset_after_cleaning %>% filter(proxy_decision=="exclude")
SAonly_proxies<- dataset_after_cleaning %>% filter(proxy_decision=="senstivity-analysis" | proxy_decision=="sensitivity-analysis")
```

The data of each article was extracted evenly by one of the three authors (SD, TR, JMGS), with another screener verifying the procedure. We extracted the primary data (i.e., measures of central tendency, variance and sample size) from text whenever possible. In cases where primary data was only available in figures (e.g., boxplots), we used `metaDigitise` v.1.0.1 (@pick2018). If primary data was not available, we extracted any inferential statistics (t value, F value) that could be transformed into an effect size. In cases where the values could not be extracted from the article but datasets were provided, we calculated the means, SD and sample sizes from the provided raw data. When data was presented in different sources (e.g. in a boxplot and as an inferential statistics), we prioritized the source having the maximum resolution (the boxplot, in this example).

To obtain missing data (e.g., non-reported results, missing sample sizes, etc.) and to locate grey literature (i.e. unpublished data), we wrote to the corresponding author of the articles that passed full-text screening using a standardized email template (\@figsupp-email). We contacted four authors for missing data, while all corresponding authors were asked for any potential unpublished data that could be included in our meta-analysis. All four authors responded to our email but only two could provide the data as requested; one did not have the data any more and one did not give a conclusive response. In addition, one author provided data for an unpublished study (see description in @figsupp-unpublished-study) and one of the authors of this study (MO) provided data for an unpublished study. In all, we extracted data for `{r} included_papers_no - 2` published articles and 2 unpublished studies.

We extracted the means, variances and sample size for each control-treatment pairwise comparison. Controls included non-aromatic materials such as grass or moss, lesser amounts of aromatic material, or no material at all. Treatments consisted of aromatic materials added by authors, birds, or both. In cases comparing non-aromatic material to a blank nest, the non-aromatic material was considered the treatment. We recorded the type of comparison being made as a moderator.

For each study we also extracted information related to the study subjects (species of birds, location of the study population), the methodology of the study (plant species used as treatment and control, fitness proxy measured) and information about the extracted data (location or source of the data recorded). We categorized the fitness proxy measured into 6 broad categories - physiology, morphology, reproduction, behaviour, parasite and pathogenic load and phenology. For parasites or pathogens, we categorised them into arthropods or micro-organisms. For a complete list of variables extracted see @figsupp-variables . We did not include some traits in all our analyses becaause it was unclear how they related to fitness (k=`{r} nrow(excluded_proxies)` , n=`{r} length(unique(excluded_proxies$paper_ID))`;see @figsupp-excluded-proxies for the list and reason of exclusion). We also had to exclude some effect sizes (n=?, k=?) because of the low sample size per group (less than x).

## Effect size calculation

```{r load effect size, include=FALSE, echo=FALSE}

inferential_stats<-dataset_after_cleaning%>%filter(statistics_type=="t-test"|statistics_type=="ANCOVA")

contingency_table<-dataset_after_cleaning%>%filter(proxy_decision=="contingency table")

no_ES<-dataset_after_cleaning%>%filter(proxy_comment == "use 0 as ES")

negative_lnRR<-dataset_after_cleaning%>%
  filter(measure_central_tendency_experiment<=0 | measure_central_tendency_control<=0)

```

We calculated two effect sizes for each pairwise comparison: a log response ratio (lnRR) and a standardised mean difference with heteroskedasticity correction (SMDH), using the `escalc` function (`measure= ROM` and `SMDH` respectively, `vtype= "LS"`) in the R package `metafor` (@metafor). We specified the treatment group as the numerator and the control group as the denominator for lnRR and treatment minus control for SMDH, such that positive values indicate the 'fitness' value increased in the treatment group. In cases such as measures of parasite load or prevalence, laying or hatching dates, courtship time, scab score etc., the effect size sign was inverted to ensure comparability across all fitness proxies. In all cases, the effect sizes were coded such that the fitness proxies were positively correlated with the overall fitness.

For inferential statistics (e.g., *t* value, *F* value), we estimated SMD values (k=`{r} nrow(inferential_stats)`, n=`{r} length(unique(inferential_stats$paper_ID))`) using the equations provided by @thehand2019. For dichotomous variables (e.g., number of successful nests, k= `{r} nrow(contingency_table)`, n= `{r} length(unique(contingency_table$paper_ID))`), we calculated the log odds ratio (lnOR) using the `escalc` function (`measure=OR`) from `metafor` and subsequently converted it to SMD using equations from @thehand2019. Since lnRR cannot be calculated from inferential statistics, or dichotomous variables, or for comparisons including zero or negative values (k=`{r} nrow(negative_lnRR)`, n=`{r} length(unique(negative_lnRR$paper_ID))`), the number of effect sizes included in the lnRR (k=?, n=?) subset is smaller than that of SMDH (k=?,n=?).

When calculating each effect size's associated sampling variance, we prioritised the number of nests in each group over the number of individuals as the sample size, whenever possible to reduce the effect of genetic relatedness non-independence. In cases where the control or treatment group was used repeatedly across multiple comparisons, sample sizes were adjusted by dividing them by the number of times each group was used in a comparison to account for shared group non-independence. For the cases where the authors of an original article reported no statistically significant difference between two groups but did not provide numerical values (k=`{r} nrow(no_ES)`, n=`{r} length(unique(no_ES$paper_ID))`), we assigned zero as the effect size for that comparision and imputed its SD using the missing-case approach described in @nakagawa2022.

## Data Analysis

We performed all analyses in R version 4.3.0 (@positteam2023). All multilevel meta-analytical models and meta-regressions were fit using `metafor v.4.8-0` (@metafor). Sampling variances were modelled as variance-covariance matrices assuming a 0.5 correlation (*ρ*) between sampling variances from the same study [@noble2017] using the `vcalc` function from `metafor`. (See Supplementary Appendix @figsupp-correlation for sensitivity analyses using different values of *ρ*).

We considered the following levels of non-independence for our analyses: (a) multiple effect sizes collected from same study (Paper ID), (b) bird species (Species ID) and (c) geographical locations (population ID), (d) some sub-populations (Experiment ID; e.g. birds studied at the same location for multiple years or birds studied at the same location separated by distance) and (e) group of individuals (Group ID; eg. males and females in a population, first clutch and second clutch) were represented in our data several times, and (f) some traits were collected on the same group of animals repeatedly (Repeated trait ID; e.g. chick mass on day 7, 14 and 21).

Unless stated otherwise, we fitted the following random effects in all our models to account for non-independence: Paper ID, Group ID, Species ID, Population ID and an observation-level random effect (Observation ID). Our random effect structure deviated from our pre-registration since, initially we had pre-registered the inclusion of Experiment ID and repeated trait ID and not population ID. However, Experiment ID and Group ID were almost indistinguishable, so we decided to leave out the one with lesser resolution i.e. Experiment ID. Additionally, few effect sizes were measured repeatedly, making the Repeated trait ID almost identical to the observation-level random effect (Observation ID), so we excluded it (@figsupp-random-effect). Last, we decided to add Population ID, since we noticed that several studies were performed in the same geographical locations.

### Meta-analytic models: biological predictions

To estimate the effect of green nest material addition on fitness across birds, we used two effect sizes: lnRR and SMDH. Thus, throughout, we fitted two separate sets of model for each effect size. First, we fitted multilevel intercept-only meta-analytical model to test the overall effect, its uncertainty and heterogeneity. We used the `rma.mv` function from the `metafor` package using the 'REML' method and a *t* distribution for calculating test statistics and confidence intervals for the fixed effects. We also conducted a sensitivity analysis including some traits for which the relationship with fitness was unclear, to test the robustness of our results (@figsupp-MLMA-excluded; k=`{r} nrow(SAonly_proxies)` , n=`{r} length(unique(SAonly_proxies$paper_ID))`; see @figsupp-SAonly_proxies). For all intercept-only models, we estimated the total amount of absolute heterogeneity ($𝜎^2$; the variation not explained by the sampling error), its sources (or relative heterogeneity; $I^2$), mean-standardised metric ($CVH2$) and variance-mean-standardised metric of heterogeneity ($M2$) following the pluralistic approach suggested by @yang2025.

<!--# You have to correct the numbers for contingency table. It does not include the one data point from unpublished results -->

<!--# Check what is MLMA-excluded and figsupp-SAonly-proxies for figures and decide which one makes more sense to include/exclude?? -->

To test the relative importance of the two main hypotheses suggested for green nest material: Courtship hypothesis and Parental care hypothesis (combining nest protection and drug hypothesis), we ran a multilevel meta-regression with 'hypothesis' (Courtship hypothesis, Parental care hypothesis, Both) as a moderator. For this and all meta-regressions, we estimated $R_{marginal}^2$, which corresponds to the percentage of heterogeneity explained by the moderators (@nakagawa2012).

### Meta-analytic models: exploratory analyses

We performed additional pre-registered multi-level meta-regressions for which we did not have clear or strong directional predictions on specific subsets of data. To understand the effect of green nest material on the parasite load depending on the type of parasites, we used effect sizes on parasites and pathogens (n=?, k=?). We ran a meta-regression with a type of parasite as a moderator (levels: arthropod, micro-organism). We also ran a meta-regression with with the entire dataset set using bird species as a moderator (levels: Cyanistes caeruleus, Sturnus unicolor, Tachycineta bicolor, Sturnus vulgaris, Parus major, Passer cinnamomeus, Buteo buteo) to understand if the effect of green nest material on fitness depended on the system being studied. To test the effect of time of addition of green nest material on the fitness outcome (levels: before egg hatching, after egg hatching, continuously throughout the nesting phase) we ran a meta-regression with Time of addition of green nest material as the moderator. We explored the effect of green nest material on fitness depending on the type of trait studies (levels:physiology, morphology, reproduction, behaviour, parasite and pathogenic load, phenology) using a meta-regression with the trait type as a moderator. We explored the impact of methodological differences on the relationship between green nest material and fitness. To understand the effect of type of experimental design (i.e. whether the authors compared - non-aromatic vs. aromatic, no added material vs. aromatic, no added material vs. non-aromatic), we ran a meta-regression with the type of experimental design as a moderator.

### Publication Bias Test

Publication bias occurs when a subset of research findings, most often statistically non-significant ones, are less likely to be published or being published as compared to the rest. We tested for evidence of publication bias, both small-study and decline effects following @nakagawa2021. Publication bias tests were performed for both lnRR and SMDH separately, and using the same random effect structure as reported above. To test for small-study effects, we ran a multilevel meta-regression, using the inverse of the effective sample size as a moderator. To test for decline effects (also known as time-lag bias; @sanchez-tojar2018), we ran multilevel meta-regression year of publication (mean-centered) as a moderator.

# Results

```{r}
dataset_analysis <- read_csv("data/03_data_cleaning/dataset_analysis.csv")

ES_per_study<-dataset_analysis%>%group_by(paper_ID)%>%count(paper_ID)

birds_info<-dataset_analysis%>%group_by(bird_species)%>%count(bird_species)
birds_info$perc<-(birds_info$n/nrow(dataset_analysis))*100
```

We obtained `{r} nrow(dataset_analysis)` effect sizes from `{r} length(unique(dataset_analysis$paper_ID))` studies conducted on `{r} length(unique(dataset_analysis$bird_species))` species across `{r} length(unique(dataset_analysis$population_ID))` geographical locations. About `{r} round(sum(birds_info$perc[birds_info$perc>15]))`% of the effect sizes came from three bird species (`{r} birds_info$bird_species[birds_info$perc > 15][1]`, k = `{r} birds_info$n[birds_info$perc > 15][1]`; `{r} birds_info$bird_species[birds_info$perc > 15][2]`, k = `{r} birds_info$n[birds_info$perc > 15][2]`; `{r} birds_info$bird_species[birds_info$perc > 15][3]`, k = `{r} birds_info$n[birds_info$perc > 15][3]`). Our dataset included two unpublished studies (k=XX effect sizes), with the remaining studies being published between `{r} min((dataset_analysis$year_publication),na.rm = T)` and `{r} max((dataset_analysis$year_publication),na.rm = T)`. The mean (±SD) number of effect sizes per study was `{r} as.numeric(format(mean(ES_per_study$n), digit=3))` ± `{r} as.numeric(format(sd(ES_per_study$n), digit=2))`(range: `{r} min(ES_per_study$n)`--`{r} max(ES_per_study$n)`) and the median and mean sample size per effect size were `{r} median(c(dataset_analysis$effective_n_control,dataset_analysis$effective_n_experiment))` and `{r} as.numeric(format(mean(c(dataset_analysis$effective_n_control,dataset_analysis$effective_n_experiment)), digit=3))` respectively (range: `{r} min(c(dataset_analysis$effective_n_control,dataset_analysis$effective_n_experiment))`--`{r} max(c(dataset_analysis$effective_n_control,dataset_analysis$effective_n_experiment))`; `{r} nrow(dataset_analysis%>%select(paper_ID,effective_n_control,effective_n_experiment, lnRR, SMDH)%>%filter(effective_n_experiment<2|effective_n_control<2))` effect sizes had a sample size of less than 2 in each group).

**Overall effect of green nest material on fitness proxies**

```{r}
intercept_lnRR<-readRDS(here::here("model/intercept_lnRR.rds"))
intercept_SMDH<-readRDS(here::here("model/intercept_SMDH.rds"))
bivariate_model<-readRDS(here::here("sensitivity_analysis/bivariate_model.rds"))
```

Both lnRR and SMDH intercept-only multilevel meta-analyses showed an increase in fitness in the nests with an increase in green nest material compared to control nests; however, the overall effect was only statistically significant for SMDH.

The model using lnRR showed an average increase in fitness of `{r} round((exp(as.double(coef(intercept_lnRR)))-1)*100,1)`% in nests with an experimental increase in green nest material \[@fig-mean-effect (A); lnRR = `{r} as.double(formatC(coef(intercept_lnRR), digits = 3, format = "f"))`, 95% CI: \[`{r} as.double(formatC((intercept_lnRR)$ci.lb, digits = 3, format = "f"))` to `{r} as.double(formatC((intercept_lnRR)$ci.ub, digits = 3, format = "f"))`\], 95% PI: \[`{r} as.double(formatC(predict(intercept_lnRR)$pi.lb, digits = 3, format = "f"))` to `{r} as.double(formatC(predict(intercept_lnRR)$pi.ub, digits = 3, format = "f"))`\], p-value = `{r} as.double(formatC(intercept_lnRR$pval, digits = 3, format = "f"))`, k = `{r} intercept_lnRR$k`, n = `{r} length(intercept_lnRR$s.levels.f[[1]])`\] but heterogeneity among effect sizes was high. Total raw heterogeneity (𝜎^2^) was `{r} round(sum(intercept_lnRR$sigma2), 3)` (Q = `{r} round(sum(intercept_lnRR$QE),1)`, p \< 0.001), the measures of relative heterogeneity (I^2^~total~ = `{r} round(as.double(orchaRd::i2_ml(intercept_lnRR)[1]),1)`%) and standardized heterogeneity metrics (CVH1~total~ = `{r} round(as.double(orchaRd::cvh1_ml(intercept_lnRR)[1]),2)` and M1~total =~ `{r} round(as.double(orchaRd::m1_ml(intercept_lnRR)[1]),2)`) suggested high variability across effect sizes. The model using SMDH showed a small positive and statistically significant effect of green nest material on fitness \[@fig-mean-effect (B) ; SMDH = `{r} as.double(formatC(coef(intercept_SMDH), digits = 3, format = "f"))`, 95% CI: \[`{r} as.double(formatC((intercept_SMDH)$ci.lb, digits = 3, format = "f"))` to `{r} as.double(formatC((intercept_SMDH)$ci.ub, digits = 3, format = "f"))`\], 95% PI: \[`{r} as.double(formatC(predict(intercept_SMDH)$pi.lb, digits = 3, format = "f"))` to `{r} as.double(formatC(predict(intercept_SMDH)$pi.ub, digits = 3, format = "f"))`\], p-value = `{r} as.double(formatC(intercept_SMDH$pval, digits = 3, format = "f"))`, k = `{r} intercept_SMDH$k`, n = `{r} length(intercept_SMDH$s.levels.f[[1]])`\], and heterogeneity among effect sizes was moderate to high. Raw heterogeneity (𝜎^2^) was `{r} round(sum(intercept_SMDH$sigma2), 3)` (Q = `{r} round(sum(intercept_SMDH$QE),1)`, p \< 0.001), relative heterogeneity was substantially smaller compared to that of lnRR (I^2^~total~ = `{r} round(as.double(orchaRd::i2_ml(intercept_SMDH)[1]),1)`%) and the standardized heterogeneity metrics (CV~total~ = `{r} round(as.double(orchaRd::cvh1_ml(intercept_SMDH)[1]),2)` and M~total =~ `{r} round(as.double(orchaRd::m1_ml(intercept_SMDH)[1]),2)`) were moderately high (50th to 75th percentile of empirically derived estimates from meta-analyses in ecology and evolution; @yang2023; see @supp-tab-heterogeneity for detailed estimates of heterogeneity)

While lnRR and SMDH differ in their measurement scale, both effect sizes suggest a small positive but non-conclusive effect of green nest material on fitness. These differences in overall effect size magnitude likely reflect the underlying biological and statistical assumptions - lnRR incorporates the mean difference at a multiplicative scale whereas SMDH expresses the difference on an additive scale. The difference in relative heterogeneity (I^2^) between the two effect sizes aligns with broader trends observed in meta-analysis in ecology and evolution (@yang2024). As a pre-registered sensitivity analysis, we fitted a bivariate multilevel meta-analytic model including both the effect sizes. While neither intercept was statistically significant, the results strengthens the interpretation that any effect of green nest material on fitness outcomes in likely to be very small at best \[lnRR = `{r} as.double(formatC(coef(bivariate_model)[1], digit = 3, format="f"))`, p-value = `{r} as.double(formatC(bivariate_model$pval[1], digits = 3, format = "f"))` and SMDH = `{r} as.double(formatC(coef(bivariate_model)[2], digit = 3, format="f"))` , p-value = `{r} as.double(formatC(bivariate_model$pval[2], digits = 3, format = "f"))`\].

![Overall effect of green nest material on fitness: Mean meta-analytic estimate (large circle) with 95% confidence intervals (thicker bars) and prediction intervals (thinner bars) are shown along with the data points (coloured circles) scaled by precision (inverse of standard error). Results are shown as an orchard plot for the intercept-only model (A) using log response ratio and (B) standardized mean difference as effect sizes. k: number of effect sizes (number of studies is shown in the brackets)](figures/Figure_overall_effect.png){#fig-mean-effect width="1059"}

**Mechanisms behind green nest material effects**

```{r}
lnRR_hypothesis<-readRDS(here::here("model/lnRR_hypothesis.rds"))
SMDH_hypothesis<-readRDS(here::here("model/SMDH_hypothesis.rds"))
```

Using the multilevel meta-regressions with hypothesis type as a moderator, we found no clear support for either of the proposed mechanism [@fig-hypothesis]. There was a small positive but non-significant effect of green nest material on reproductive success in parents (courtship hypothesis) \[lnRR~courtship~ = `{r} as.double(formatC(coef(lnRR_hypothesis)[2], digit = 3, format="f"))` , p-value = `{r} as.double(formatC(lnRR_hypothesis$pval[2], digits = 3, format = "f"))` and SMDH~courtship~ = `{r} as.double(formatC(coef(SMDH_hypothesis)[2], digit = 3, format="f"))` , p-value = `{r} as.double(formatC(SMDH_hypothesis$pval[2], digits = 3, format = "f"))`\]. We found even smaller, non-significant effect of green nest material on protective effect or overall health benefits for the nestling as suggested by nest protection and drug hypothesis \[lnRR~parental_care~ = `{r} as.double(formatC(coef(lnRR_hypothesis)[3], digit = 3, format="f"))` , p-value = `{r} as.double(formatC(lnRR_hypothesis$pval[3], digits = 3, format = "f"))` and SMDH~parental_care~ = `{r} as.double(formatC(coef(SMDH_hypothesis)[3], digit = 3, format="f"))` , p-value = `{r} as.double(formatC(SMDH_hypothesis$pval[3], digits = 3, format = "f"))`\]. In case of the third category, both, that included the fitness proxies that could not be clearly assigned to either mechanisms, we found a small positive effect \[lnRR~both~ = `{r} as.double(formatC(coef(lnRR_hypothesis)[1], digit = 3, format="f"))` , p-value = `{r} as.double(formatC(lnRR_hypothesis$pval[1], digits = 3, format = "f"))` and SMDH~both~ = `{r} as.double(formatC(coef(SMDH_hypothesis)[1], digit = 3, format="f"))` , p-value = `{r} as.double(formatC(SMDH_hypothesis$pval[1], digits = 3, format = "f"))`\]. The moderator explained very little heterogeneity ($R^2~marginal$ for lnRR = `{r} as.double(round(orchaRd::r2_ml(lnRR_hypothesis)*100,1))[1]`% and ($R^2~marginal$ for SMDH = `{r} as.double(round(orchaRd::r2_ml(SMDH_hypothesis)*100,1))[1]`%).

![There is no clear evidence to favour either the courtship hypothesis or the parental care hypothesis as drivers of fitness benefits from green nest material in nests. Orchard plot of the multilevel meta-regression for A. log response ratio (lnRR) and B. Standardised mean difference (SMDH). Plots showing the mean estimates (large circles), 95% confidence intervals (thicker bars) and prediction intervals (thinner bars) and the individual data points (coloured circles) scaled by precision (inverse of standard error). k: number of effect sizes (number of studies is shown in the brackets).](figures/Figure_hypothesis.png){#fig-hypothesis width="1035"}

## Pre-registered exploratory analyses

```{r}
# Loading models for experimental design
lnRR_design<-readRDS(here::here("model/lnRR_design.rds"))
SMDH_design<-readRDS(here::here("model/SMDH_design.rds"))

# Loading models for parasite type
lnRR_parasite<-readRDS(here::here("model/lnRR_parasite.rds"))
SMDH_parasite<-readRDS(here::here("model/SMDH_parasite.rds"))

# Loading models for time of green nest material addition
lnRR_time_gnm_addition<-readRDS(here::here("model/lnRR_time_gnm_addition.rds"))
SMDH_time_gnm_addition<-readRDS(here::here("model/SMDH_time_gnm_addition.rds"))

# Loading models for the type of trait studied
lnRR_trait<-readRDS(here::here("model/lnRR_trait.rds"))
SMDH_trait<-readRDS(here::here("model/SMDH_trait.rds"))

# Loading models for the bird species studied
lnRR_birds<-readRDS(here::here("model/lnRR_birds.rds"))
SMDH_birds<-readRDS(here::here("model/SMDH_birds.rds"))

```

There was no evidence that the effect of green nest material on parasite or pathogen load differed between arthropods and micro-organisms \[[@supp-fig-MR-non-sig]A-B; lnRR~arthropods~ = `{r} as.double(formatC(coef(lnRR_parasite)[1], digit = 3, format="f"))` , p-value = `{r} as.double(formatC(lnRR_parasite$pval[1], digits = 3, format = "f"))` and SMDH~arthropods~ = `{r} as.double(formatC(coef(SMDH_parasite)[1], digit = 3, format="f"))` , p-value = `{r} as.double(formatC(SMDH_parasite$pval[1], digits = 3, format = "f"))`; lnRR~micro-organisms~ = `{r} as.double(formatC(coef(lnRR_parasite)[2], digit = 3, format="f"))` , p-value = `{r} as.double(formatC(lnRR_parasite$pval[2], digits = 3, format = "f"))` and SMDH~micro-organisms~ = `{r} as.double(formatC(coef(SMDH_parasite)[2], digit = 3, format="f"))` , p-value = `{r} as.double(formatC(SMDH_parasite$pval[2], digits = 3, format = "f"))`\]. The type of parasite or pathogen explained very little heterogeneity ($R^2~marginal$ for lnRR = `{r} as.double(round(orchaRd::r2_ml(lnRR_parasite)*100,1))[1]`% and $R^2~marginal$ for SMDH = `{r} as.double(round(orchaRd::r2_ml(SMDH_parasite)*100,1))[1]`%).

The meta-regression using bird species as a moderator showed no significantly different effect of green nest material on fitness depending on the bird being studied \[[@supp-fig-MR-non-sig]C-D\]; . However, it explained some heterogeneity in the effect sizes ($R^2~marginal$ for lnRR = `{r} as.double(round(orchaRd::r2_ml(lnRR_birds)*100,1))[1]`% and $R^2~marginal$ for SMDH = `{r} as.double(round(orchaRd::r2_ml(SMDH_birds)*100,1))[1]`%).

We explored the effect of green nest material on fitness depending on the type of trait studies (levels:physiology, morphology, reproduction, behaviour, parasite and pathogenic load, phenology) using a meta-regression with the trait type as a moderator.

To test the effect of time of addition of green nest material on the fitness outcome (levels: before egg hatching, after egg hatching, continuously throughout the nesting phase) we ran a meta-regression with Time of addition of green nest material as the moderator.

We explored the impact of methodological differences on the relationship between green nest material and fitness. To understand the effect of type of experimental design (i.e. whether the authors compared - non-aromatic vs. aromatic, no added material vs. aromatic, no added material vs. non-aromatic), we ran a meta-regression with type of experimental design as the moderator.

\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\--

## Supplementary Appendix

```{r results='asis'}
cat(readLines("tables/SA_intercept_only.html"), sep ="\n")
```

![](figures/Figure_parasite_birds_trait.png){#supp-fig-MR-non-sig}

<details>

<summary>Supplementary materials</summary>

{{< include supplementary_appendix.qmd >}}

</details>