### 2025-05-08 notebook/figures.qmd

---
title: "Figures and Plots"
---

# Intercept only model

Orchard plot for overall effect of green nest material on fitness using lnRR and SMDH

```{r orchard plot for intercept only models}
devtools::install_github("daniel1noble/orchaRd", ref = "main", force = TRUE)
pacman::p_load(here,tidyverse,metafor,orchaRd,patchwork,cowplot)

# Loading models for SMDH and lnRR

intercept_lnRR<-readRDS(here::here("model/intercept_lnRR.rds"))
intercept_SMDH<-readRDS(here::here("model/intercept_SMDH.rds"))

# Plot for lnRR
intercept_lnRR_plot <- orchard_plot(intercept_lnRR,
                                    mod = "1",
                                    group = "paper_ID",
                                    trunk.size = 0.5,
                                    branch.size = 2,
                                    twig.size = 0.5,
                                    alpha = 0.3,
                                    xlab = "Effect size (Log Response Ratio- lnRR)",
                                    colour = "#218833",
                                    fill = TRUE) +
  scale_fill_manual(values = "#218833") +
  scale_colour_manual(values = "#218833") +
  theme_minimal(base_size = 14) +
  theme(
    panel.border = element_rect(colour = "black", fill = NA, size = 0.8),
    legend.position = c(0.98, 0.05),
    legend.justification = c(1, 0),
    legend.direction = "horizontal",
    legend.title = element_text(size = 7),
    legend.text = element_text(size = 7),
    axis.title = element_text(size = 12),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_blank()
  ) +
  labs(x = "Intercept")

# Plot for SMDH
intercept_SMDH_plot <- orchard_plot(intercept_SMDH,
                                    mod = "1",
                                    group = "paper_ID",
                                    trunk.size = 0.5,
                                    branch.size = 2,
                                    twig.size = 0.5,
                                    alpha = 0.3,
                                    xlab = "Effect size (Standardized Mean Difference- SMDH)",
                                    colour = "#895129",
                                    fill = TRUE) +
  scale_fill_manual(values = "#895129") +
  scale_colour_manual(values = "#895129") +
  theme_minimal(base_size = 14) +
  theme(
    panel.border = element_rect(colour = "black", fill = NA, size = 0.8),
    legend.position = c(0.98, 0.05),
    legend.justification = c(1, 0), 
    legend.direction = "horizontal",
    legend.title = element_text(size = 7),
    legend.text = element_text(size = 7),
    axis.title = element_text(size = 12),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_blank()
  ) +
  labs(x = "Intercept")

# Combine using patchwork
Figure_intercept <- intercept_lnRR_plot / intercept_SMDH_plot +
  plot_annotation(tag_level = c("A","B"))

ggsave(here::here("figures/Figure_overall_effect.png"), plot = Figure_intercept, height = 250, width = 400, units = "mm", dpi = 600)
```

# Hypothesis meta-regression

```{r orchard plot hypothesis meta-regression}

# Loading models for SMDH and lnRR

lnRR_hypothesis<-readRDS(here::here("model/lnRR_hypothesis.rds"))
SMDH_hypothesis<-readRDS(here::here("model/SMDH_hypothesis.rds"))

# Plot lnRR meta-regression on hypothesis

lnRR_hypothesis_plot <-
  orchard_plot(mod_results(
      lnRR_hypothesis,
      mod = "Hypothesis",
      group = "paper_ID"
    ), xlab ="") +
  scale_colour_manual(values = c("#895129", "mediumpurple4", "#218833")) +
  scale_fill_manual(values = c("#895129", "mediumpurple4", "#218833")) +
  labs(x = "",
       y = "lnRR")+
    theme(
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.8),
    legend.position.inside = c(0.98, 0.05),
    legend.justification = c(1, 0), 
    legend.direction = "horizontal",
    legend.title = element_text(size = 7),
    legend.text = element_text(size = 7),
    axis.title = element_text(size = 12),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(angle = 90, vjust = 0.5, hjust = 1)
  )

# Plot SMDH meta-regression on hypothesis

SMDH_hypothesis_plot <-
  orchard_plot(
    mod_results(
      SMDH_hypothesis,
      mod = "Hypothesis",
      group = "paper_ID"
    ), xlab ="") +
  scale_colour_manual(values = c("#895129", "mediumpurple4", "#218833")) +
  scale_fill_manual(values = c("#895129", "mediumpurple4", "#218833")) +
  labs(x = "",
       y = "SMDH")+
    theme(
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.8),
    legend.position.inside = c(0.98, 0.05),
    legend.justification = c(1, 0), 
    legend.direction = "horizontal",
    legend.title = element_text(size = 7),
    legend.text = element_text(size = 7),
    axis.title = element_text(size = 12),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(angle = 90, vjust = 0.5, hjust = 1)
  )

# Combine using patchwork
Figure_hypothesis <- lnRR_hypothesis_plot + SMDH_hypothesis_plot +
  plot_annotation(tag_level = c("A","B"))

ggsave(here::here("figures/Figure_hypothesis.png"), plot = Figure_hypothesis, height = 150, width = 400, units = "mm", dpi = 600)
```

# Exploratory Analysis

## Parasite Type

```{r orchard plot parasite type meta-regression}

# Loading models for SMDH and lnRR

lnRR_parasite<-readRDS(here::here("model/lnRR_parasite.rds"))
SMDH_parasite<-readRDS(here::here("model/SMDH_parasite.rds"))

lnRR_parasite_plot <-
  orchard_plot(mod_results(
      lnRR_parasite,
      mod = "parasite_type",
      group = "paper_ID"
    ), xlab ="") +
  scale_colour_manual(values = c("#895129", "#218833")) +
  scale_fill_manual(values = c("#895129",  "#218833")) +
  labs(x = "",
       y = "lnRR")+
    theme(
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.8),
    legend.position = "none",
    # legend.position.inside = c(0.98, 0.05),
    # legend.justification = c(1, 0), 
    # legend.direction = "horizontal",
    # legend.title = element_text(size = 7),
    # legend.text = element_text(size = 7),
    axis.title = element_text(size = 12),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(angle = 0, vjust = 0.5, hjust = 1)
  )

SMDH_parasite_plot <-
  orchard_plot(mod_results(
      SMDH_parasite,
      mod = "parasite_type",
      group = "paper_ID"
    ), xlab ="") +
  scale_colour_manual(values = c("#895129", "#218833")) +
  scale_fill_manual(values = c("#895129",  "#218833")) +
  labs(x = "",
       y = "SMDH")+
    theme(
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.8),
    legend.position = "none",
    # legend.position.inside = c(0.98, 0.05),
    # legend.justification = c(1, 0), 
    # legend.direction = "horizontal",
    # legend.title = element_text(size = 7),
    # legend.text = element_text(size = 7),
    axis.title = element_text(size = 12),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(angle = 0, vjust = 0.5, hjust = 1)
  )

```

## Bird Species

```{r orchard plot bird species meta-regression}

# Loading models for SMDH and lnRR

lnRR_birds<-readRDS(here::here("model/lnRR_birds.rds"))
SMDH_birds<-readRDS(here::here("model/SMDH_birds.rds"))

lnRR_birds_plot <-
  orchard_plot(mod_results(
      lnRR_birds,
      mod = "bird_species",
      group = "paper_ID"
    ), xlab ="") +
  scale_colour_manual(values = c("darkorange2", "lightgoldenrod3", "mediumpurple4", "#218833","#895129","lightcoral","grey5")) +
  scale_fill_manual(values = c("darkorange2", "lightgoldenrod3", "mediumpurple4", "#218833","#895129","lightcoral","grey5")) +
  labs(x = "",
       y = "lnRR")+
    theme(
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.8),
    legend.position = "none",
    # legend.position.inside = c(0.98, 0.05),
    # legend.justification = c(1, 0), 
    # legend.direction = "horizontal",
    # legend.title = element_text(size = 7),
    # legend.text = element_text(size = 7),
    axis.title = element_text(size = 12),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(angle = 0, vjust = 0.5, hjust = 1)
  )

SMDH_birds_plot <-
  orchard_plot(mod_results(
      SMDH_birds,
      mod = "bird_species",
      group = "paper_ID"
    ), xlab ="") +
  scale_colour_manual(values =  c("darkorange2", "lightgoldenrod3", "mediumpurple4", "#218833","#895129","lightcoral","grey5")) +
  scale_fill_manual(values =  c("darkorange2", "lightgoldenrod3", "mediumpurple4", "#218833","#895129","lightcoral","grey5")) +
  labs(x = "",
       y = "SMDH")+
    theme(
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.8),
    legend.position = "none",
    # legend.position.inside = c(0.98, 0.05),
    # legend.justification = c(1, 0), 
    # legend.direction = "horizontal",
    # legend.title = element_text(size = 7),
    # legend.text = element_text(size = 7),
    axis.title = element_text(size = 12),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(angle =0, vjust = 0.5, hjust = 1)
  )

```

## Trait studied

```{r orchard plot trait-studied meta-regression}

# Loading models for SMDH and lnRR

lnRR_trait<-readRDS(here::here("model/lnRR_trait.rds"))
SMDH_trait<-readRDS(here::here("model/SMDH_trait.rds"))

lnRR_trait_plot <-
  orchard_plot(mod_results(
      lnRR_trait,
      mod = "trait_type",
      group = "paper_ID"
    ), xlab ="") +
  scale_colour_manual(values = c("darkorange2", "lightgoldenrod3", "mediumpurple4", "#218833","#895129","lightcoral","grey5")) +
  scale_fill_manual(values = c("darkorange2", "lightgoldenrod3", "mediumpurple4", "#218833","#895129","lightcoral","grey5")) +
  labs(x = "",
       y = "lnRR")+
    theme(
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.8),
    legend.position = "bottom",
    # legend.position.inside = c(0.98, 0.05),
    # legend.justification = c(1, 0), 
    # legend.direction = "horizontal",
    legend.title = element_text(size = 7),
    legend.text = element_text(size = 7),
    axis.title = element_text(size = 12),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(angle = 0, vjust = 0.5, hjust = 1)
  )

SMDH_trait_plot <-
  orchard_plot(mod_results(
      SMDH_trait,
      mod = "trait_type",
      group = "paper_ID"
    ), xlab ="") +
  scale_colour_manual(values =  c("darkorange2", "lightgoldenrod3", "mediumpurple4", "#218833","#895129","lightcoral","grey5")) +
  scale_fill_manual(values =  c("darkorange2", "lightgoldenrod3", "mediumpurple4", "#218833","#895129","lightcoral","grey5")) +
  labs(x = "",
       y = "SMDH")+
    theme(
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.8),
    legend.position = "bottom",
    # legend.position.inside = c(0.98, 0.05),
    # legend.justification = c(1, 0), 
    # legend.direction = "horizontal",
    legend.title = element_text(size = 7),
    legend.text = element_text(size = 7),
    axis.title = element_text(size = 12),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(angle = 0, vjust = 0.5, hjust = 1)
  )

```

### Combined plot

```{r combining meta-regression plots, no effect}

## I am going to remove the y ticks from the SMDH plots to make it smaller..
SMDH_parasite_plot <- SMDH_parasite_plot +
  theme(axis.text.y  = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank())

SMDH_birds_plot <- SMDH_birds_plot +
  theme(axis.text.y  = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank())

SMDH_trait_plot <- SMDH_trait_plot +
  theme(axis.text.y  = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank())

# Combine using patchwork
Figure_parasite_birds_trait <- ((lnRR_parasite_plot + SMDH_parasite_plot) / 
                                  (lnRR_birds_plot + SMDH_birds_plot) / 
                                  (lnRR_trait_plot + SMDH_trait_plot)) +
  plot_annotation(tag_levels = "A")

ggsave(here::here("figures/Figure_parasite_birds_trait.png"), plot = Figure_parasite_birds_trait, height = 250, width = 400, units = "mm", dpi = 600)

```

## Time of green nest material addition

```{r orchard plot time of gnm addition meta-regression}

# Loading models for SMDH and lnRR

lnRR_time_gnm_addition<-readRDS(here::here("model/lnRR_time_gnm_addition.rds"))
SMDH_time_gnm_addition<-readRDS(here::here("model/SMDH_time_gnm_addition.rds"))

lnRR_time_gnm_addition_plot <-
  orchard_plot(mod_results(
      lnRR_time_gnm_addition,
      mod = "time_of_gnm_addition",
      group = "paper_ID"
    ), xlab ="") +
  scale_colour_manual(values = c( "#218833","mediumpurple4","#895129")) +
  scale_fill_manual(values = c( "#218833","mediumpurple4","#895129")) +
  labs(x = "",
       y = "lnRR")+
    theme(
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.8),
    legend.position = "bottom",
    # legend.position.inside = c(0.98, 0.05),
    # legend.justification = c(1, 0), 
    # legend.direction = "horizontal",
    legend.title = element_text(size = 7),
    legend.text = element_text(size = 7),
    axis.title = element_text(size = 12),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(angle = 0, vjust = 0.5, hjust = 1)
  ) + scale_x_discrete(labels = c("After egg laying", "Before egg laying", "Continous addition"))

SMDH_time_gnm_addition_plot <-
  orchard_plot(mod_results(
      SMDH_time_gnm_addition,
      mod = "time_of_gnm_addition",
      group = "paper_ID"
    ), xlab ="") +
  scale_colour_manual(values = c( "#218833","mediumpurple4","#895129")) +
  scale_fill_manual(values = c( "#218833","mediumpurple4","#895129")) +
  labs(x = "",
       y = "SMDH")+
    theme(
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.8),
    legend.position = "bottom",
    # legend.position.inside = c(0.98, 0.05),
    # legend.justification = c(1, 0), 
    # legend.direction = "horizontal",
    legend.title = element_text(size = 7),
    legend.text = element_text(size = 7),
    axis.title = element_text(size = 12),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(angle = 0, vjust = 0.5, hjust = 1)
  )

```

## Experimental Design

```{r orchard plot experimental design meta-regression}

# Loading models for SMDH and lnRR

lnRR_design<-readRDS(here::here("model/lnRR_design.rds"))
SMDH_design<-readRDS(here::here("model/SMDH_design.rds"))

lnRR_design_plot <-
  orchard_plot(mod_results(
      lnRR_design,
      mod = "comparision_type",
      group = "paper_ID"
    ), xlab ="") +
  scale_colour_manual(values = c( "#218833","mediumpurple4","#895129")) +
  scale_fill_manual(values = c( "#218833","mediumpurple4","#895129")) +
  labs(x = "",
       y = "lnRR")+
    theme(
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.8),
    legend.position = "none",
    # legend.position.inside = c(0.98, 0.05),
    # legend.justification = c(1, 0), 
    # legend.direction = "horizontal",
    legend.title = element_text(size = 7),
    legend.text = element_text(size = 7),
    axis.title = element_text(size = 12),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(angle = 0, vjust = 0.5, hjust = 1)
  ) + scale_x_discrete(labels = c("Non-aromatic Vs. Aromatic","Aromatic Vs. Blank", "Non-Aromatic Vs. Blank"))

SMDH_design_plot <-
  orchard_plot(mod_results(
      SMDH_design,
      mod = "comparision_type",
      group = "paper_ID"
    ), xlab ="") +
  scale_colour_manual(values = c( "#218833","mediumpurple4","#895129")) +
  scale_fill_manual(values = c( "#218833","mediumpurple4","#895129")) +
  labs(x = "",
       y = "SMDH")+
    theme(
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.8),
    legend.position = "none",
    # legend.position.inside = c(0.98, 0.05),
    # legend.justification = c(1, 0), 
    # legend.direction = "horizontal",
    legend.title = element_text(size = 7),
    legend.text = element_text(size = 7),
    axis.title = element_text(size = 12),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(angle = 0, vjust = 0.5, hjust = 1)
  )

```

### Combined plot

```{r combining meta-regression plots, effect}

## I am going to remove the y ticks from the SMDH plots to make it smaller..
SMDH_design_plot <- SMDH_design_plot +
  theme(axis.text.y  = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank())

SMDH_time_gnm_addition_plot <- SMDH_time_gnm_addition_plot +
  theme(axis.text.y  = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank())

# Combine using patchwork
Figure_design_time_gnm_addition <- ((lnRR_design_plot + SMDH_design_plot) /
                               (lnRR_time_gnm_addition_plot + SMDH_time_gnm_addition_plot)) +
  plot_annotation(tag_levels = "A")

ggsave(here::here("figures/Figure_design_time_gnm_addition.png"), plot = Figure_design_time_gnm_addition, height = 300, width = 400, units = "mm", dpi = 600)

```

##