### 2025-06-17 docs/notebook/figures.qmd
Some sensitivity analysis added

---
title: "Figures and Plots"
---

# Visualising location and bird species

I am using the figure code from here and adapting it for my needs..

Mentesana, L., Hau, M., D'Amelio, P.B., Adreani, N.M. and Sánchez-Tójar, A. (2025), Do Egg Hormones Have Fitness Consequences in Wild Birds? A Systematic Review and Meta-Analysis. Ecology Letters, 28: e70100. <https://doi.org/10.1111/ele.70100> \[[Code](https://github.com/ASanchez-Tojar/meta-analysis_egg_hormones_and_fitness/blob/main/code/003_figures.R)\]

```{r}
devtools::install_github("daniel1noble/orchaRd", ref = "main", force = TRUE)
pacman::p_load(ggplot2,dplyr,stringr,readr,cowplot)

# Loading cleaned dataset
dataset_analysis<- read_csv(here::here("data/03_data_cleaning/dataset_analysis.csv"))

# To split population_ID into location and country
dataset_analysis[c('Location','Country')] <- str_split_fixed(dataset_analysis$population_ID, ',', 2)

### For the purpose of this map, I am adding the the longitude and latitude of the study area manually. They exist in the dataset_analysis$population_location as they were extracted from the paper. I have also added some by just searching the general location since the coordinates were not provided for these..

data_coordinates<-data.frame(Location = unique(dataset_analysis$Location),
                             Lat_Location=c(51.50,38.41,40.45,40.75,45.40,52.39,37.18,
                                             40.53,54,39.05,
                                             51.57,48,42.54,39.95,
                                             40.13,47.96,28.10,52.06,60.39),
                             Long_Location=c(19.29,-9.02,-3.50,-3.78,-64.10,6.05,-3.11,
                                              -4.0,-124,-76.81,
                                              -2.26,11,8.91,-75.16,
                                              -8.27,11.18,107.10,8.25,5.32))

## Let's now merge the datasets by the location names

dataset_analysis <- dataset_analysis %>%
  left_join(data_coordinates, by = "Location")

## Count how many times each Location appears in the full dataset
location_counts <- dataset_analysis %>%
  group_by(Location) %>%
  summarise(size.freq = n_distinct(paper_ID), .groups = 'drop')

# I will subset the data base so that I get one row per Location AND bird_species.
# Like this, I get one point per location independently of the number of effect sizes 
# we have per place.
dataset_analysis_unique_rows <- dataset_analysis %>%
  distinct(Location, bird_species, .keep_all = TRUE)

# Merge the counts into the unique rows dataset
dataset_analysis_unique_rows <- dataset_analysis_unique_rows %>%
  left_join(location_counts, by = "Location")

# create data for world coordinates using  
# map_data() function 
world_coordinates <- map_data("world")

# create world map using ggplot() function 
# geom_map() function takes world coordinates as input to plot world map 
fig_map <- ggplot() + 
  geom_map(data = world_coordinates, map = world_coordinates, 
    aes(long, lat, map_id = region), 
    color = "white", fill= "#D9EAF7") + 
  geom_point(data = dataset_analysis_unique_rows, 
             aes(x = Long_Location, y = Lat_Location,
                 size = size.freq, colour=bird_species), alpha = 0.6) +
scale_size(range = c(3, 10)) +
  scale_fill_manual(values = c("#11a3d9","#f6c800",  "#fb5158", 
                               "#ea60e9", "#40ad28","#ef8c27","#a083e2")) +

  scale_colour_manual(values = c("#11a3d9","#f6c800",  "#fb5158", 
                               "#ea60e9", "#40ad28","#ef8c27","#a083e2")) +
  guides(size = "none", alpha = "none") +
  labs(x = "Longitude",
       y = "Latitude")+
  theme_classic() +
  theme(legend.position = c(0.88, 0.10),
        legend.justification = c("right", "bottom"),
        legend.background = element_rect(fill = "white", color = "black"),
        axis.title = element_text(size = 12),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_blank())

# To add the zoomed in map.. to show the points more clearly
zoom_map <- ggplot() + 
  geom_map(data = world_coordinates, map = world_coordinates,
           aes(long, lat, map_id = region),
           color = "white", fill = "lightgrey") +
  geom_point(data = dataset_analysis_unique_rows,
             aes(x = Long_Location, y = Lat_Location,
                 size = size.freq, colour = bird_species),
             alpha = 0.6) +
  scale_size(range = c(3, 10)) +
  scale_colour_manual(values = c("#11a3d9","#f6c800","#fb5158",
                                 "#ea60e9","#40ad28","#ef8c27","#a083e2")) +
  coord_cartesian(xlim = c(-120, 120), ylim = c(25, 65)) +  # Northern Hemisphere
  theme_void() +
  theme(
    legend.position = "none",
    plot.background = element_rect(color = "white", fill="white"),
    panel.border = element_rect(colour = "black", fill = NA, size = 1)
  )

geographic_distribution <- ggdraw() +
  draw_plot(fig_map) +
  draw_plot(zoom_map, x = 0.10, y = 0.12, width = 0.62, height = 0.22)

ggsave(here::here("figures/Figure_Distribution.png"), plot = geographic_distribution, height = 250, width = 400, units = "mm", dpi = 600)
```

# Visualizing author network

```{r}
# # Loading cleaned dataset
# dataset_analysis<- read_csv(here::here("data/03_data_cleaning/dataset_analysis.csv"))
# 
# library(tidyverse)
# library(tidyr)
# library(igraph)
# installed.packages("ggraph")
# library(ggraph)
# 
# clean_author <- function(name) {
#   name %>%
#     str_replace_all(",", "") %>%   # remove commas
#     str_squish() %>%              # remove excess whitespace
#     str_to_title()                # standardize case (optional, if needed)
# }
# 
# # Apply cleaning to first and last author fields
# author_roles <- dataset_analysis %>%
#   distinct(paper_ID, .keep_all = TRUE) %>%
#   select(paper_ID, authors, bird_species) %>%
#   mutate(
#     author_list = str_split(authors, ";\\s*"),
#     first_author = map_chr(author_list, 1),
#     last_author = map_chr(author_list, ~ tail(.x, 1))
#   )%>%
#   mutate(
#     first_author=case_when(
#       first_author == "Polo, V" | first_author == "Polo V" ~ "Polo V.",
#       first_author == "Soler J J" ~ "Soler J. J.",
#       first_author == "Glądalski, M" ~ "Gladalski M.",
#       first_author == "Gwinner, H" | first_author == "Gwinner H" ~ "Gwinner H.",
#       TRUE ~ as.character(first_author)
#     ))%>%
#   mutate(
#     last_author=case_when(
#       last_author == "Veiga J P" ~ "Veiga J. P.",
#       TRUE ~ as.character(last_author)
#     )
#   )
# 
# # 2. Create Core Author Pairs (First → Last)
# author_pairs_core <- author_roles %>%
#   select(paper_ID, first_author, last_author) %>%
#   rename(from = first_author, to = last_author) %>%
#   filter(from != to) %>%
#   count(from, to, name = "weight")
# 
# # 3. Build Author Graph
# author_graph <- graph_from_data_frame(author_pairs_core, directed = TRUE)
# 
# # 4. Assign Dominant Bird Species to PIs (last authors)
# author_species <- author_roles %>%
#   select(last_author, bird_species) %>%
#   group_by(last_author, bird_species) %>%
#   count() %>%
#   slice_max(n, n = 1) %>%
#   distinct(last_author, .keep_all = TRUE)
# 
# # Add species as vertex attribute
# V(author_graph)$bird_species <- author_species$bird_species[
#   match(V(author_graph)$name, author_species$last_author)
# ]
# 
# # 5. Optional: Scale node size by number of papers per author
# author_paper_counts <- author_roles %>%
#   pivot_longer(cols = c(first_author, last_author), names_to = "role", values_to = "author") %>%
#   count(author, name = "paper_count")
# 
# # Add paper count attribute
# V(author_graph)$paper_count <- author_paper_counts$paper_count[
#   match(V(author_graph)$name, author_paper_counts$author)
# ]
# 
# # Replace NAs with small size for isolated authors
# V(author_graph)$paper_count[is.na(V(author_graph)$paper_count)] <- 1
# 
# # 6. Plot with ggraph
# ggraph(author_graph, layout = "fr") +
#   geom_edge_link(aes(width = weight), alpha = 0.4, color = "grey50") +
#   geom_node_point(aes(color = bird_species, size = paper_count)) +
#   geom_node_text(aes(label = name), repel = TRUE, size = 3) +
#   scale_edge_width(range = c(0.3, 2)) +
#   scale_size(range = c(3, 10)) +
#   theme_void() +
#   labs(title = "Author Collaboration Network by Bird Species",
#        size = "Paper Count") +
#   theme(legend.position = "right")
#

```

#

# Intercept only model

Orchard plot for overall effect of green nest material on fitness using lnRR and SMDH

```{r orchard plot for intercept only models}
pacman::p_load(here,tidyverse,metafor,orchaRd,patchwork,cowplot)

# Loading models for SMDH and lnRR

intercept_lnRR<-readRDS(here::here("model/intercept_lnRR.rds"))
intercept_SMDH<-readRDS(here::here("model/intercept_SMDH.rds"))

# Plot for lnRR
intercept_lnRR_plot <- orchaRd::orchard_plot(intercept_lnRR,
                                    mod = "1",
                                    group = "paper_ID",
                                    trunk.size = 0.5,
                                    branch.size = 2,
                                    twig.size = 0.5,
                                    alpha = 0.3,
                                    xlab = "Effect size (Log Response Ratio- lnRR)",
                                    fill = T) +
   scale_fill_manual(values = "#218833") +
   scale_colour_manual(values = "#218833" ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.border = element_rect(colour = "black", fill = NA, size = 0.8),
    legend.position = c(0.98, 0.05),
    legend.justification = c(1, 0),
    legend.direction = "horizontal",
    legend.title = element_text(size = 7),
    legend.text = element_text(size = 7),
    axis.title = element_text(size = 12),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_blank()
  ) +
  labs(x = "Intercept")

# Plot for SMDH
intercept_SMDH_plot <- orchaRd::orchard_plot(intercept_SMDH,
                                    mod = "1",
                                    group = "paper_ID",
                                    trunk.size = 0.5,
                                    branch.size = 2,
                                    twig.size = 0.5,
                                    alpha = 0.3,
                                    xlab = "Effect size (Standardized Mean Difference- SMDH)",
                                    fill = TRUE) +
  scale_fill_manual(values = "#895129") +
  scale_colour_manual(values = "#895129") +
  theme_minimal(base_size = 14) +
  theme(
    panel.border = element_rect(colour = "black", fill = NA, size = 0.8),
    legend.position = c(0.98, 0.05),
    legend.justification = c(1, 0), 
    legend.direction = "horizontal",
    legend.title = element_text(size = 7),
    legend.text = element_text(size = 7),
    axis.title = element_text(size = 12),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_blank()
  ) +
  labs(x = "Intercept")

# Combine using patchwork
Figure_intercept <- intercept_lnRR_plot / intercept_SMDH_plot +
  plot_annotation(tag_level = c("A","B"))

ggsave(here::here("figures/Figure_overall_effect.png"), plot = Figure_intercept, height = 250, width = 400, units = "mm", dpi = 600)
```

# Hypothesis meta-regression

```{r orchard plot hypothesis meta-regression}

# Loading models for SMDH and lnRR

lnRR_hypothesis<-readRDS(here::here("model/lnRR_hypothesis.rds"))
SMDH_hypothesis<-readRDS(here::here("model/SMDH_hypothesis.rds"))

# Plot lnRR meta-regression on hypothesis

lnRR_hypothesis_plot <-
  orchard_plot(mod_results(
      lnRR_hypothesis,
      mod = "Hypothesis",
      group = "paper_ID"
    ), xlab ="") +
  scale_colour_manual(values = c("#895129", "mediumpurple4", "#218833")) +
  scale_fill_manual(values = c("#895129", "mediumpurple4", "#218833")) +
  labs(x = "",
       y = "lnRR")+
    theme(
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.8),
    legend.position.inside = c(0.98, 0.05),
    legend.justification = c(1, 0), 
    legend.direction = "horizontal",
    legend.title = element_text(size = 7),
    legend.text = element_text(size = 7),
    axis.title = element_text(size = 12),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(angle = 90, vjust = 0.5, hjust = 1)
  )

# Plot SMDH meta-regression on hypothesis

SMDH_hypothesis_plot <-
  orchard_plot(
    mod_results(
      SMDH_hypothesis,
      mod = "Hypothesis",
      group = "paper_ID"
    ), xlab ="") +
  scale_colour_manual(values = c("#895129", "mediumpurple4", "#218833")) +
  scale_fill_manual(values = c("#895129", "mediumpurple4", "#218833")) +
  labs(x = "",
       y = "SMDH")+
    theme(
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.8),
    legend.position.inside = c(0.98, 0.05),
    legend.justification = c(1, 0), 
    legend.direction = "horizontal",
    legend.title = element_text(size = 7),
    legend.text = element_text(size = 7),
    axis.title = element_text(size = 12),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(angle = 90, vjust = 0.5, hjust = 1)
  )

# Combine using patchwork
Figure_hypothesis <- lnRR_hypothesis_plot + SMDH_hypothesis_plot +
  plot_annotation(tag_level = c("A","B"))

ggsave(here::here("figures/Figure_hypothesis.png"), plot = Figure_hypothesis, height = 150, width = 400, units = "mm", dpi = 600)
```

# Exploratory Analysis

## Parasite Type

```{r orchard plot parasite type meta-regression}

# Loading models for SMDH and lnRR

lnRR_parasite<-readRDS(here::here("model/lnRR_parasite.rds"))
SMDH_parasite<-readRDS(here::here("model/SMDH_parasite.rds"))

lnRR_parasite_plot <-
  orchard_plot(mod_results(
      lnRR_parasite,
      mod = "parasite_type",
      group = "paper_ID"
    ), xlab ="") +
  scale_colour_manual(values = c("#895129", "#218833")) +
  scale_fill_manual(values = c("#895129",  "#218833")) +
  labs(x = "",
       y = "lnRR")+
    theme(
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.8),
    legend.position = "none",
    # legend.position.inside = c(0.98, 0.05),
    # legend.justification = c(1, 0), 
    # legend.direction = "horizontal",
    # legend.title = element_text(size = 7),
    # legend.text = element_text(size = 7),
    axis.title = element_text(size = 12),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(angle = 0, vjust = 0.5, hjust = 1)
  )

SMDH_parasite_plot <-
  orchard_plot(mod_results(
      SMDH_parasite,
      mod = "parasite_type",
      group = "paper_ID"
    ), xlab ="") +
  scale_colour_manual(values = c("#895129", "#218833")) +
  scale_fill_manual(values = c("#895129",  "#218833")) +
  labs(x = "",
       y = "SMDH")+
    theme(
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.8),
    legend.position = "none",
    # legend.position.inside = c(0.98, 0.05),
    # legend.justification = c(1, 0), 
    # legend.direction = "horizontal",
    # legend.title = element_text(size = 7),
    # legend.text = element_text(size = 7),
    axis.title = element_text(size = 12),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(angle = 0, vjust = 0.5, hjust = 1)
  )

```

## Bird Species

```{r orchard plot bird species meta-regression}

# Loading models for SMDH and lnRR

lnRR_birds<-readRDS(here::here("model/lnRR_birds.rds"))
SMDH_birds<-readRDS(here::here("model/SMDH_birds.rds"))

lnRR_birds_plot <-
  orchard_plot(mod_results(
      lnRR_birds,
      mod = "bird_species",
      group = "paper_ID"
    ), xlab ="") +
  scale_colour_manual(values = c("darkorange2", "lightgoldenrod3", "mediumpurple4", "#218833","#895129","lightcoral","grey5")) +
  scale_fill_manual(values = c("darkorange2", "lightgoldenrod3", "mediumpurple4", "#218833","#895129","lightcoral","grey5")) +
  labs(x = "",
       y = "lnRR")+
    theme(
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.8),
    legend.position = "none",
    # legend.position.inside = c(0.98, 0.05),
    # legend.justification = c(1, 0), 
    # legend.direction = "horizontal",
    # legend.title = element_text(size = 7),
    # legend.text = element_text(size = 7),
    axis.title = element_text(size = 12),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(angle = 0, vjust = 0.5, hjust = 1)
  )

SMDH_birds_plot <-
  orchard_plot(mod_results(
      SMDH_birds,
      mod = "bird_species",
      group = "paper_ID"
    ), xlab ="") +
  scale_colour_manual(values =  c("darkorange2", "lightgoldenrod3", "mediumpurple4", "#218833","#895129","lightcoral","grey5")) +
  scale_fill_manual(values =  c("darkorange2", "lightgoldenrod3", "mediumpurple4", "#218833","#895129","lightcoral","grey5")) +
  labs(x = "",
       y = "SMDH")+
    theme(
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.8),
    legend.position = "none",
    # legend.position.inside = c(0.98, 0.05),
    # legend.justification = c(1, 0), 
    # legend.direction = "horizontal",
    # legend.title = element_text(size = 7),
    # legend.text = element_text(size = 7),
    axis.title = element_text(size = 12),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(angle =0, vjust = 0.5, hjust = 1)
  )

```

## Trait studied

```{r orchard plot trait-studied meta-regression}

# Loading models for SMDH and lnRR

lnRR_trait<-readRDS(here::here("model/lnRR_trait.rds"))
SMDH_trait<-readRDS(here::here("model/SMDH_trait.rds"))

lnRR_trait_plot <-
  orchard_plot(mod_results(
      lnRR_trait,
      mod = "trait_type",
      group = "paper_ID"
    ), xlab ="") +
  scale_colour_manual(values = c("darkorange2", "lightgoldenrod3", "mediumpurple4", "#218833","#895129","lightcoral","grey5")) +
  scale_fill_manual(values = c("darkorange2", "lightgoldenrod3", "mediumpurple4", "#218833","#895129","lightcoral","grey5")) +
  labs(x = "",
       y = "lnRR")+
    theme(
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.8),
    legend.position = "bottom",
    # legend.position.inside = c(0.98, 0.05),
    # legend.justification = c(1, 0), 
    # legend.direction = "horizontal",
    legend.title = element_text(size = 7),
    legend.text = element_text(size = 7),
    axis.title = element_text(size = 12),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(angle = 0, vjust = 0.5, hjust = 1)
  )

SMDH_trait_plot <-
  orchard_plot(mod_results(
      SMDH_trait,
      mod = "trait_type",
      group = "paper_ID"
    ), xlab ="") +
  scale_colour_manual(values =  c("darkorange2", "lightgoldenrod3", "mediumpurple4", "#218833","#895129","lightcoral","grey5")) +
  scale_fill_manual(values =  c("darkorange2", "lightgoldenrod3", "mediumpurple4", "#218833","#895129","lightcoral","grey5")) +
  labs(x = "",
       y = "SMDH")+
    theme(
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.8),
    legend.position = "bottom",
    # legend.position.inside = c(0.98, 0.05),
    # legend.justification = c(1, 0), 
    # legend.direction = "horizontal",
    legend.title = element_text(size = 7),
    legend.text = element_text(size = 7),
    axis.title = element_text(size = 12),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(angle = 0, vjust = 0.5, hjust = 1)
  )

```

### Combined plot

```{r combining meta-regression plots, no effect}

## I am going to remove the y ticks from the SMDH plots to make it smaller..
SMDH_parasite_plot <- SMDH_parasite_plot +
  theme(axis.text.y  = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank())

SMDH_birds_plot <- SMDH_birds_plot +
  theme(axis.text.y  = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank())

SMDH_trait_plot <- SMDH_trait_plot +
  theme(axis.text.y  = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank())

# Combine using patchwork
Figure_parasite_birds_trait <- ((lnRR_parasite_plot + SMDH_parasite_plot) / 
                                  (lnRR_birds_plot + SMDH_birds_plot) / 
                                  (lnRR_trait_plot + SMDH_trait_plot)) +
  plot_annotation(tag_levels = "A")

ggsave(here::here("figures/Figure_parasite_birds_trait.png"), plot = Figure_parasite_birds_trait, height = 250, width = 400, units = "mm", dpi = 600)

```

## Time of green nest material addition

```{r orchard plot time of gnm addition meta-regression}

# Loading models for SMDH and lnRR

lnRR_time_gnm_addition<-readRDS(here::here("model/lnRR_time_gnm_addition.rds"))
SMDH_time_gnm_addition<-readRDS(here::here("model/SMDH_time_gnm_addition.rds"))

lnRR_time_gnm_addition_plot <-
  orchard_plot(mod_results(
      lnRR_time_gnm_addition,
      mod = "time_of_gnm_addition",
      group = "paper_ID"
    ), xlab ="") +
  scale_colour_manual(values = c( "#218833","mediumpurple4","#895129")) +
  scale_fill_manual(values = c( "#218833","mediumpurple4","#895129")) +
  labs(x = "",
       y = "lnRR")+
    theme(
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.8),
    legend.position = "bottom",
    # legend.position.inside = c(0.98, 0.05),
    # legend.justification = c(1, 0), 
    # legend.direction = "horizontal",
    legend.title = element_text(size = 7),
    legend.text = element_text(size = 7),
    axis.title = element_text(size = 12),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(angle = 0, vjust = 0.5, hjust = 1)
  ) + scale_x_discrete(labels = c("After egg laying", "Before egg laying", "Continous addition"))

SMDH_time_gnm_addition_plot <-
  orchard_plot(mod_results(
      SMDH_time_gnm_addition,
      mod = "time_of_gnm_addition",
      group = "paper_ID"
    ), xlab ="") +
  scale_colour_manual(values = c( "#218833","mediumpurple4","#895129")) +
  scale_fill_manual(values = c( "#218833","mediumpurple4","#895129")) +
  labs(x = "",
       y = "SMDH")+
    theme(
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.8),
    legend.position = "bottom",
    # legend.position.inside = c(0.98, 0.05),
    # legend.justification = c(1, 0), 
    # legend.direction = "horizontal",
    legend.title = element_text(size = 7),
    legend.text = element_text(size = 7),
    axis.title = element_text(size = 12),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(angle = 0, vjust = 0.5, hjust = 1)
  )

```

## Experimental Design

```{r orchard plot experimental design meta-regression}

# Loading models for SMDH and lnRR

lnRR_design<-readRDS(here::here("model/lnRR_design.rds"))
SMDH_design<-readRDS(here::here("model/SMDH_design.rds"))

lnRR_design_plot <-
  orchard_plot(mod_results(
      lnRR_design,
      mod = "comparision_type",
      group = "paper_ID"
    ), xlab ="") +
  scale_colour_manual(values = c( "#218833","mediumpurple4","#895129")) +
  scale_fill_manual(values = c( "#218833","mediumpurple4","#895129")) +
  labs(x = "",
       y = "lnRR")+
    theme(
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.8),
    legend.position = "none",
    # legend.position.inside = c(0.98, 0.05),
    # legend.justification = c(1, 0), 
    # legend.direction = "horizontal",
    legend.title = element_text(size = 7),
    legend.text = element_text(size = 7),
    axis.title = element_text(size = 12),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(angle = 0, vjust = 0.5, hjust = 1)
  ) + scale_x_discrete(labels = c("Non-aromatic Vs. Aromatic","Aromatic Vs. Blank", "Non-Aromatic Vs. Blank"))

SMDH_design_plot <-
  orchard_plot(mod_results(
      SMDH_design,
      mod = "comparision_type",
      group = "paper_ID"
    ), xlab ="") +
  scale_colour_manual(values = c( "#218833","mediumpurple4","#895129")) +
  scale_fill_manual(values = c( "#218833","mediumpurple4","#895129")) +
  labs(x = "",
       y = "SMDH")+
    theme(
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.8),
    legend.position = "none",
    # legend.position.inside = c(0.98, 0.05),
    # legend.justification = c(1, 0), 
    # legend.direction = "horizontal",
    legend.title = element_text(size = 7),
    legend.text = element_text(size = 7),
    axis.title = element_text(size = 12),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(angle = 0, vjust = 0.5, hjust = 1)
  )

```

### Combined plot

```{r combining meta-regression plots, effect}

## I am going to remove the y ticks from the SMDH plots to make it smaller..
SMDH_design_plot <- SMDH_design_plot +
  theme(axis.text.y  = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank())

SMDH_time_gnm_addition_plot <- SMDH_time_gnm_addition_plot +
  theme(axis.text.y  = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank())

# Combine using patchwork
Figure_design_time_gnm_addition <- ((lnRR_design_plot + SMDH_design_plot) /
                               (lnRR_time_gnm_addition_plot + SMDH_time_gnm_addition_plot)) +
  plot_annotation(tag_levels = "A")

ggsave(here::here("figures/Figure_design_time_gnm_addition.png"), plot = Figure_design_time_gnm_addition, height = 300, width = 400, units = "mm", dpi = 600)

```

# Publication Bias Tests

```{r}

###---Small study Effects ----###

lnRR_small_study<-readRDS(here::here("model/lnRR_small_study.rds"))
SMDH_small_study<-readRDS(here::here("model/SMDH_small_study.rds"))

lnRR_small_study_bubble<-orchaRd::bubble_plot(mod_results(lnRR_small_study,
                     mod="sqrt_effective_n_inv",
                     group="paper_ID",
                     weights="prop"),
                     group="paper_ID") +
  scale_colour_manual(values ="mediumpurple4") +
  scale_fill_manual(values = "mediumpurple4") +
  labs(y = "Effect size (lnRR)",
       x = expression(sqrt(1 / N[effective])))+
    theme(
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.8),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 10),
    axis.title = element_text(size = 14),
    axis.text.x = element_text(size = 14),
    axis.text.y = element_text(angle = 0, vjust = 0.5, hjust = 1))

SMDH_small_study_bubble<-orchaRd::bubble_plot(mod_results(SMDH_small_study,
                     mod="sqrt_effective_n_inv",
                     group="paper_ID",
                     weights="prop"),
                     group="paper_ID") +
  scale_colour_manual(values ="mediumpurple4") +
  scale_fill_manual(values = "mediumpurple4") +
  labs(y = "Effect size (SMDH)",
       x = expression(sqrt(1 / N[effective])))+
    theme(
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.8),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 10),
    axis.title = element_text(size = 14),
    axis.text.x = element_text(size = 14),
    axis.text.y = element_text(angle = 0, vjust = 0.5, hjust = 1))

###---Decline Effects ----###

lnRR_decline_eff<-readRDS(here::here("model/lnRR_decline_eff.rds"))
SMDH_decline_eff<-readRDS(here::here("model/SMDH_decline_eff.rds"))

lnRR_decline_eff_bubble<-orchaRd::bubble_plot(mod_results(lnRR_decline_eff,
                     mod="year_MeanCenter",
                     group="paper_ID",
                     weights="prop"),
                     group="paper_ID")+
  scale_colour_manual(values ="#218833") +
  scale_fill_manual(values = "#218833") +
  labs(y = "Effect size (lnRR)",
       x = "Publication Year (Mean Cenetered)")+
    theme(
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.8),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 10),
    axis.title = element_text(size = 14),
    axis.text.x = element_text(size = 14),
    axis.text.y = element_text(angle = 0, vjust = 0.5, hjust = 1))

SMDH_decline_effect_bubble<-orchaRd::bubble_plot(mod_results(SMDH_decline_eff,
                     mod="year_MeanCenter",
                     group="paper_ID",
                     weights="prop"),
                     group="paper_ID")+
  scale_colour_manual(values ="#218833") +
  scale_fill_manual(values = "#218833") +
  labs(y = "Effect size (SMDH)",
       x = "Publication Year (Mean Cenetered)")+
    theme(
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.8),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 10),
    axis.title = element_text(size = 14),
    axis.text.x = element_text(size = 14),
    axis.text.y = element_text(angle = 0, vjust = 0.5, hjust = 1))

# Combine using patchwork
Figure_publication_bias <- ((lnRR_small_study_bubble + SMDH_small_study_bubble) /
                               (lnRR_decline_eff_bubble + SMDH_decline_effect_bubble)) +
  plot_annotation(tag_levels = "A")

ggsave(here::here("figures/Figure_publication_bias.png"), plot = Figure_publication_bias, height = 300, width = 400, units = "mm", dpi = 600)
```