### 2025-07-06 docs/code/07_data_analysis.Rmd

---
title: "Data Analysis"
author: "Shreya D."
date: "`r Sys.Date()`"
output: html_document
---

# Loading Data for analysis

```{r load_packages-data, warning=FALSE, message=FALSE}

# Setting up workspace

# For plotting trees, and although not strictly necessary, we may need the `ggtree` R package, which you should be able to install by running:
 if (!require("BiocManager", quietly = TRUE))
 install.packages("BiocManager")
 BiocManager::install("ggtree")

if (!require("orchaRd", quietly = TRUE))
devtools::install_github("daniel1noble/orchaRd", ref = "main", force = TRUE)

pacman::p_load(here,tidyverse,metafor,ggpubr,orchaRd,ggpubfigs)

#Packages needed

rm(list=ls()) ## cleaning up

set.seed(430) #set seed for reproducibility (for random number generation)

# Functions needed

# function to estimate typical sampling error variance obtained from 
# https://github.com/Yefeng0920/heterogeneity_ecoevo/tree/main/function
sigma2_v <- function(mod){
  sigma2_v <- sum(1 / mod$vi) * (mod$k - 1) /
    (sum(1 / mod$vi)^2 - sum((1 / mod$vi)^2))
  return(sigma2_v)
}

# ------------------------------------------------------------
# Loading cleaned dataset
# ------------------------------------------------------------

dataset_analysis<-read_csv(here::here("data/04_data_analysis/dataset_analysis.csv"))

dataset_lnRR<-read_csv(here::here("data/04_data_analysis/dataset_lnRR.csv"))

dataset_SMDH<-read_csv(here::here("data/04_data_analysis/dataset_SMDH.csv"))

```

### Creating Phylogeny for the plot

```{r}
# install.packages("pacman")
# load packages

pacman::p_load(rotl,
               ape,
               ggtree)

# We have 7 bird species in the dataset
bird_species <- unique(dataset_analysis$bird_species)

# extracting taxonomic information
taxa <- rotl::tnrs_match_names(names = bird_species)
taxa

# check approximate matches
taxa[taxa$approximate_match==TRUE & !(is.na(taxa$approximate_match)),]

# phylogenetic information for our species from the Open Tree of Life

tree <- rotl::tol_induced_subtree(ott_ids = taxa[["ott_id"]], label_format = "name")
ape::is.binary(tree)

# here are the species included in the tree
sort(tree$tip.label)

# removing the underscore "_" from the tree tip.label
tree$tip.label <- gsub("_"," ", tree$tip.label)
sort(tree$tip.label)

# compute branch lengths of tree
phylo_branch <- compute.brlen(tree, method = "Grafen", power = 1)

# check tree is ultrametric
is.ultrametric(phylo_branch) # TRUE

# matrix to be included in the models
bird_phylo <- vcv(phylo_branch, cor = T)
bird_phylo

plot(tree) #check: https://www.rdocumentation.org/packages/ape/versions/5.3/topics/plot.phylo
```

### Visualizing random effects structure

```{r}
# install.packages("devtools")
devtools::install_github("davidsjoberg/ggsankey")
library(ggsankey)

sankey_data <- dataset_analysis %>%
  select(Observation_ID, repeated_trait_ID_coded, paper_ID, experiment_ID_coded, group_ID_coded)%>%
  mutate(across(c(Observation_ID, repeated_trait_ID_coded, paper_ID, experiment_ID_coded, group_ID_coded), as.factor)) %>%
  make_long(Observation_ID, repeated_trait_ID_coded, paper_ID, experiment_ID_coded, group_ID_coded)

sankey_data <- dataset_analysis %>%
  select(paper_ID, bird_species, population_ID)%>%
  mutate(across(c(paper_ID, bird_species, population_ID), as.factor)) %>%
  make_long(paper_ID, bird_species, population_ID)

ggplot(sankey_data, aes(x = x, 
               next_x = next_x, 
               node = node, 
               next_node = next_node,
               fill = factor(node))) +
  geom_sankey() +
  scale_fill_discrete(drop=FALSE) +
  theme(legend.position = "none")

ggplot(sankey_data, aes(x = x, next_x = next_x, node = node, next_node = next_node, fill = factor(node))) +
  geom_alluvial(flow.alpha = .6) +
  scale_fill_viridis_d(drop = FALSE) +
  theme_alluvial(base_size = 18) +
  labs(x = NULL) +
  theme(legend.position = "none",
        plot.title = element_text(hjust = .5))

```

# Preparing datasets

```{r}
# Let's just quickly ensure all the important variables are 
# in the correct data type

dataset_analysis<- dataset_analysis%>%mutate(across(c(paper_ID, Observation_ID, repeated_trait_ID_coded, paper_ID, experiment_ID_coded, group_ID_coded, comparision_type,Hypothesis,trait_type,parasite_type,time_of_gnm_addition, blinding, random_assignment, missing_data, population_ID,bird_species), as.factor))

dataset_lnRR<- dataset_lnRR%>%mutate(across(c(paper_ID, Observation_ID, repeated_trait_ID_coded, paper_ID, experiment_ID_coded, group_ID_coded, comparision_type,Hypothesis,trait_type,parasite_type,time_of_gnm_addition, blinding, random_assignment, missing_data, population_ID, bird_species), as.factor))

dataset_SMDH<- dataset_SMDH%>%mutate(across(c(paper_ID, Observation_ID, repeated_trait_ID_coded, paper_ID, experiment_ID_coded, group_ID_coded, comparision_type,Hypothesis,trait_type,parasite_type,time_of_gnm_addition, blinding, random_assignment, missing_data, population_ID,bird_species), as.factor))

```

```{r seperating datasets for analysis}

## For lnRR (or simply load dataset as provided)
# dataset_lnRR<-dataset_analysis%>%
# filter(lnRR_sign != "NA" | lnRR_variance != "NA")

# selecting dataset which are not flagged
dataset_lnRR<-dataset_lnRR%>%
  filter(proxy_decision == "include")

# selecting only published dataset
dataset_lnRR_published<-dataset_lnRR%>%
  filter(!is.na(year_publication))

# selecting dataset which are flagged
lnRR_con_flagged<-dataset_analysis%>%
   filter(lnRR_sign != "NA" | lnRR_variance != "NA")%>%
  filter(proxy_decision != "SMDH only" | proxy_decision != "contingency table" )

# selecting dataset without those using 0 as ES because authors reported no effect
lnRR_sin_missingES <- dataset_lnRR %>% 
   filter(!proxy_comment %in% c("use 0 as ES"))

lnRR_flagged_no_missingES<- lnRR_con_flagged %>% 
   filter(!proxy_comment %in% c("use 0 as ES"))

# selecting only dataset that passes geary's test
lnRR_geary_passed<-dataset_lnRR%>%filter(geary_test=="pass")

## For SMDH
# selecting dataset which are not flagged
dataset_SMDH<-dataset_analysis%>%
   filter(proxy_decision == "include"| proxy_decision == "SMDH only"|proxy_decision =="contingency table") %>% filter(SMDH_sign != "NA")

# selecting only published dataset
dataset_SMDH_published<-dataset_SMDH%>%
  filter(!is.na(year_publication))

# selecting dataset which are flagged
SMDH_con_flagged<-dataset_analysis%>%
   filter(SMDH_sign != "NA")

# selecting dataset without those using 0 as ES because authors reported no effect

SMDH_sin_missingES <- dataset_SMDH %>% 
   filter(!proxy_comment %in% c("use 0 as ES"))

SMDH_flagged_no_missingES<- SMDH_con_flagged %>% 
   filter(!proxy_comment %in% c("use 0 as ES"))

# selecting dataset which are not flagged but without the SMDH calculated from inferential statistics

SMDH_no_inferential<-dataset_analysis%>%
  filter(proxy_decision == "include")%>% filter(SMDH_sign != "NA")
```

# Overall effect : Intercept-only meta-analytic model

*"Does adding green material to the nest have an adaptive function across the bird species that perform this behaviour?"*

## using lnRR

```{r intercept-only-model-lnRR}
VCV_lnRR<- vcalc(vi = lnRR_variance,
                  cluster = paper_ID,
                  obs = Observation_ID,
                  data = dataset_lnRR,
                  rho = 0.5 # assuming that the effect sizes within the same study are correlated with rho = 0.5
                  )

intercept_lnRR<- rma.mv(yi = lnRR_sign, # specify lnRR as the effect size measure;
                    V = VCV_lnRR, # specify lnRR's sampling variance-covariance matrix;
                    mods = ~ 1,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                    control = list(optimizer="optim"),
                    test = "t",
                    data = dataset_lnRR)
# saving the model
saveRDS(intercept_lnRR, file = here::here("model/intercept_lnRR.rds"))

# saving the dataset for this model
write.csv(dataset_lnRR,here::here("data/04_data_analysis/intercept_lnRR.csv"))
```

#### Exploring heterogeneity in intercept only model

This is the function from Yefeng's Github for interpreting heterogeneity. I have to download the data used for this function to be able to use it.

I am using the values from the table directly..

```{r function-interpret-heterogeneity}
# ## Function is from the Pluralistic approach to heterogeneity from Yefeng's Pre-print repository from GitHub
# 
# #-----------------------heterogeneity interpretation-----------------------#
# het_interpret <- function(observed_value, het_type, es_type, data) {
#   
#   if (!het_type %in% c("I2", "CVH", "M", "sigma2", "V_bar")) {
#     stop("Invalid heterogeneity type. Choose from 'I2', 'CVH', 'M', 'sigma2', or 'V_bar'.")
#   }
#   
#   if (!es_type %in% unique(data$es.type)) {
#     stop("Invalid effect size type. Check the levels of 'es.type'.")
#   }
#   
#   filtered_data <- data %>% filter(es.type == es_type)
#   
#   if (nrow(filtered_data) == 0) {
#     stop("No data available for the selected effect size type.")
#   }
#   
#   het_values <- filtered_data[[het_type]]
#   
#   percentiles <- quantile(het_values, probs = seq(0, 1, by = 0.05), na.rm = TRUE)
#   
#   lower_bound <- max(percentiles[percentiles <= observed_value], na.rm = TRUE)
#   upper_bound <- min(percentiles[percentiles >= observed_value], na.rm = TRUE)
#   lower_percentile <- names(percentiles)[percentiles == lower_bound]
#   upper_percentile <- names(percentiles)[percentiles == upper_bound]
#   
#   percentile_range <- paste0(lower_percentile, "-", upper_percentile, "th percentile")
#   
#   # return the results
#   return(tibble::tibble(
#     observed_value = observed_value,
#     het_type = het_type,
#     es_type = es_type,
#     percentile_range = percentile_range
#   ))
# }
# x<-het_interpret(observed_value= I2_main_model_vector[1], 
#                  het_type = "I2", 
#                  es_type = "lnRR_sign", 
#                  data= intercept_lnRR)

```

Let us look at the heterogeneity in lnRR intercept-only model

```{r}

# Then also:
# # Typical sampling variance (which captures the statistical noise of the data, 
# # but is rarely reported in the current meta-analytic practice):
sigma2_v(intercept_lnRR)

# To calculate heterogeneity, these are the functions from orchaRd:
# # I2, CV and M
sigma2_main_model_vector <- sum(intercept_lnRR$sigma2)

# 0.11, 0.27, and 0.57 for lnRR

I2_main_model_vector <- orchaRd::i2_ml(intercept_lnRR)

# 0.88, 0.95, and 0.99 for lnRR

CV2_main_model_vector <- orchaRd::cvh2_ml(intercept_lnRR)
CV1_main_model_vector <- orchaRd::cvh1_ml(intercept_lnRR)

# CVH2 (CVH1) - 1.36 (1.16), 3.76 (1.94), and 12.1 (3.48) for lnRR

M2_main_model_vector <- orchaRd::m2_ml(intercept_lnRR)
M1_main_model_vector <- orchaRd::m1_ml(intercept_lnRR)

# M2(M1) - 0.58 (0.54), 0.79 (0.66), and 0.78 for lnRR;

# # # Total unstandardized raw variance (i.e. total heterogeneity)
round(sum(intercept_lnRR$sigma2), 3)

```

## using SMDH

We are only running these analysis for the fitness proxy labelled "included". Some proxies have been flagged and we will run them in a sensitivity analysis only

The sampling variances will be fitted as variance--covariance matrices assuming a 0.5 correlation (ρs) between sampling variances from the same paper_ID (Noble et al. 2017).

The MLMA using SMDH as the effect size measure incorporate the imputed sampling variance

```{r intercept-only-model-SMDH}

VCV_SMDH <- vcalc(vi = SMDH_variance,
                  cluster = paper_ID,
                  obs = Observation_ID,
                  data = dataset_SMDH,
                  rho = 0.5 # assuming that the effect sizes within the same study are correlated with rho = 0.5
                  )

intercept_SMDH<- rma.mv(yi = SMDH_sign, # specify SMDH as the effect size measure;
                    V = VCV_SMDH, # specify SMDH's sampling variance-covariance matrix;
                    mods = ~ 1,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                    control = list(optimizer="optim"),
                    test = "t",
                    data = dataset_SMDH)

# saving the model
saveRDS(intercept_SMDH, file = here::here("model/intercept_SMDH.rds"))

# saving the dataset for this model
write.csv(dataset_SMDH,here::here("data/04_data_analysis/intercept_SMDH.csv"))
```

## SMDH-heterogeneity

Now looking at the intercept only model

```{r}
summary(intercept_SMDH)
predict(intercept_SMDH)

# Then also:
# # Typical sampling variance (which captures the statistical noise of the data, 
# # but is rarely reported in the current meta-analytic practice):
sigma2_v(intercept_SMDH)

# To calculate heterogeneity, these are the functions from orchaRd:
# # I2, CV and M
sigma2_main_model_vector <- sum(intercept_SMDH$sigma2)
# 0.54, 1.25, and 3.03 for SMD;
I2_main_model_vector <- orchaRd::i2_ml(intercept_SMDH)
# 0.78, 0.89, and 0.96 for SMD
# 0.88, 0.95, and 0.99 for lnRR
CV2_main_model_vector <- orchaRd::cvh2_ml(intercept_SMDH)
CV1_main_model_vector <- orchaRd::cvh1_ml(intercept_SMDH)
# CVH2 (CVH1) - 1.36 (1.16), 3.76 (1.94), and 12.1 (3.48) for lnRR
M2_main_model_vector <- orchaRd::m2_ml(intercept_SMDH)
M1_main_model_vector <- orchaRd::m1_ml(intercept_SMDH)
# M2 (M1) - 0.58 (0.54), 0.79 (0.66), and 0.78 for lnRR

# # # Total unstandardized raw variance (i.e. total heterogeneity)
round(sum(intercept_SMDH$sigma2), 3)

```

## Bivariate Meta-analytical model

For the overall effect size (i.e., the meta-analytical mean), we will run the bivariate multilevel meta-analytic model suggested by Yang et al. (2024), which models both lnRR and SMD(H) simultaneously (i.e., including them both as response variables and estimating their correlation; random).

```{r}
dataset_bivariate <- data.frame(paper_ID = c(dataset_lnRR$paper_ID, dataset_SMDH$paper_ID),
                      Observation_ID = c(dataset_lnRR$Observation_ID, dataset_SMDH$Observation_ID),
                      group_ID_coded = c(dataset_lnRR$group_ID_coded, dataset_SMDH$group_ID_coded),
                      population_ID = c(dataset_lnRR$population_ID, dataset_SMDH$population_ID),
                      bird_species = c(dataset_lnRR$bird_species, dataset_SMDH$bird_species),
                      ES = c(dataset_lnRR$lnRR_sign, dataset_SMDH$SMDH_sign),
                      V_ES = c(dataset_lnRR$lnRR_variance, dataset_SMDH$SMDH_variance),
                      ES_measure = c(rep("lnRR", nrow(dataset_lnRR)), rep("SMDH", nrow(dataset_SMDH)))
                      )

VCV_bivariate <- vcalc(vi = V_ES, # sampling variances of lnRR and SMD that are correlated within the same study;
             cluster = paper_ID, # study identity;
             type = ES_measure, # different types of effect size measures underlying the observed effect sizes;
             data = dataset_bivariate, # the long format data frame;
             rho = 0.2, # assuming that the effect sizes within the same study are correlated with rho = 0.5.
             nearpd = TRUE # specify whether impose the non positive definite VCV matrix (not invertible) to the nearest positive semi-definite matrix 
             )

Bivariate_model <- rma.mv(yi = ES, # specify the effect size estimate (the variable ES in our case);
                V = VCV_bivariate, # specify the imputed variance-covariance matrix;
                mods = ~ ES_measure - 1, # specify the variable "ES_measure" indicating the types of effect size measures;
                random = list(~ ES_measure | paper_ID, 
                              ~ ES_measure | Observation_ID,
                              ~ 1 | group_ID_coded,
                              ~ 1 | population_ID,
                              ~ 1 | bird_species
                              ), # add correlated random effects corresponding to the lnRR and SMD parameters in the same study;
                struct = "UN", # impose an unstructured variance-covariance of the study-specific random effects;
                test = "t", # t distribution is specified to test the overall effect against the null hypothesis and construct confidence intervals;
                method = "REML", # restricted likelihood maximum is assigned as the estimator for variance components as suggested;
                data = dataset_bivariate, # the long format dataset
                sparse = T
               )

# Saving the model
saveRDS(Bivariate_model, file = here::here("model/sensitivity_analysis/bivariate_model.rds"))
# saving the dataset for this model
write.csv(dataset_bivariate,here::here("data/05_data_sensitivity_analysis/dataset_bivariate.csv"))

summary(Bivariate_model)
```

```{r WARNING = F}
pacman::p_load(ellipse)
ab.l <- matreg(y=2, x=1, R=Bivariate_model$G, cov=TRUE, means=coef(Bivariate_model), n=Bivariate_model$g.levels.comb.k) # fit regression model

xy <- ellipse(Bivariate_model$G, centre=coef(Bivariate_model), level=0.95) # get 95% coverage region
ellipse_df <- data.frame(x = xy[, 1], y = xy[, 2]) # convert it int dataframe so that we can use ggplot to make a figure

pivot_wider(dataset_bivariate, names_from = ES_measure, values_from = c(ES, V_ES)) %>%
  ggplot() + 
  geom_point(aes(x = ES_lnRR, y = ES_SMDH, size = 1/sqrt(V_ES_lnRR)), color = "#1B9E77", alpha = 0.5) + 
  geom_point(aes(x = coef(Bivariate_model)[1], y = coef(Bivariate_model)[2]), color = "red", size = 2) + 
  geom_abline(intercept = ab.l$tab$beta[1], slope = ab.l$tab$beta[2]) +
  geom_path(data = ellipse_df, aes(x = x, y = y), color = "gray50") +
  #scale_x_continuous(limits = c(-2.5, 2.5)) + 
  #scale_y_continuous(limits = c(-2.5, 2.5)) + 
  guides(size = "none") + 
  labs(x = "lnRR", y = "SMD") + 
  theme_bw()
```

There is a very high negative correlation between the two effect sizes. This probably is an artifact of the data and not a true correlation. If you look at the model summary, the $tau^2$ for lnRR, it is extremely small 0.0008, i.e., there is very little variance. This could also be stemming from the fact that the log values in the dataset could be causing this. For SMDH on the other hand, the $tau^2$ is 0.1681 that is much larger. This difference in the variance estimate itself could be causing this artifact of the model. I will only go ahead to present this model as a senstivity analysis because I am not completely convinced by it's robustness.

# Sensitivity Analysis

### With published data only

```{r}

#----------------------------------------------------------------------------
# for lnRR as Effect Size
#----------------------------------------------------------------------------

# Preparing new variance-covariance matrix

VCV_lnRR_published <- vcalc(vi = lnRR_variance,
                  cluster = paper_ID,
                  obs = Observation_ID,
                  data = dataset_lnRR_published,
                  rho = 0.5 # assuming that the effect sizes within the same study are correlated with rho = 0.5
                  )

## Running intercept only model

SA_lnRR_published<- rma.mv(yi = lnRR_sign, # specify lnRR as the effect size measure;
                    V = VCV_lnRR_published, # specify lnRR's sampling variance-covariance matrix;
                    mods = ~ 1,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                    control = list(optimizer="optim"),
                    test = "t",
                    data = dataset_lnRR_published)

# Saving the model
saveRDS(SA_lnRR_published, file = here::here("model/sensitivity_analysis/SA_lnRR_published.rds"))
# saving the dataset for this model
write.csv(dataset_lnRR_published,here::here("data/05_data_sensitivity_analysis/dataset_lnRR_published.csv"))

# What is the effect now?

summary(SA_lnRR_published)
predict(SA_lnRR_published)

#----------------------------------------------------------------------------
# for SMDH as Effect Size
#----------------------------------------------------------------------------

# Preparing new variance-covariance matrix

VCV_SMDH_published <- vcalc(vi = SMDH_variance,
                  cluster = paper_ID,
                  obs = Observation_ID,
                  data = dataset_SMDH_published,
                  rho = 0.5 # assuming that the effect sizes within the same study are correlated with rho = 0.5
                  )

## Running intercept only model

SA_SMDH_published<- rma.mv(yi = SMDH_sign, # specify lnRR as the effect size measure;
                    V = VCV_SMDH_published, # specify lnRR's sampling variance-covariance matrix;
                    mods = ~ 1,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                    control = list(optimizer="optim"),
                    test = "t",
                    data = dataset_SMDH_published)

# Saving the model
saveRDS(SA_SMDH_published, file = here::here("model/sensitivity_analysis/SA_SMDH_published.rds"))
# saving the dataset for this model
write.csv(dataset_SMDH_published,here::here("data/05_data_sensitivity_analysis/dataset_SMDH_published.csv"))

# What is the effect now?

summary(SA_SMDH_published)
predict(SA_SMDH_published)

```

### With flagged proxies

Some traits for which the relationship with fitness is unclear or highly debated in the literature

```{r}

#----------------------------------------------------------------------------
# for lnRR as Effect Size
#----------------------------------------------------------------------------

# Preparing new variance-covariance matrix

VCV_lnRR_flagged <- vcalc(vi = lnRR_variance,
                  cluster = paper_ID,
                  obs = Observation_ID,
                  data = lnRR_con_flagged,
                  rho = 0.5 # assuming that the effect sizes within the same study are correlated with rho = 0.5
                  )

## Running intercept only model

SA_lnRR_flagged<- rma.mv(yi = lnRR_sign, # specify lnRR as the effect size measure;
                    V = VCV_lnRR_flagged, # specify lnRR's sampling variance-covariance matrix;
                    mods = ~ 1,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                    control = list(optimizer="optim"),
                    test = "t",
                    data = lnRR_con_flagged)

# Saving the model
saveRDS(SA_lnRR_flagged, file = here::here("model/sensitivity_analysis/SA_lnRR_flagged.rds"))
# saving the dataset for this model
write.csv(lnRR_con_flagged,here::here("data/05_data_sensitivity_analysis/dataset_lnRR_con_flagged.csv"))

# What is the effect now?

summary(SA_lnRR_flagged)
predict(SA_lnRR_flagged)

#----------------------------------------------------------------------------
# for SMDH as Effect Size
#----------------------------------------------------------------------------

# Preparing new variance-covariance matrix

VCV_SMDH_flagged <- vcalc(vi = SMDH_variance,
                  cluster = paper_ID,
                  obs = Observation_ID,
                  data = SMDH_con_flagged,
                  rho = 0.5 # assuming that the effect sizes within the same study are correlated with rho = 0.5
                  )

## Running intercept only model

SA_SMDH_flagged<- rma.mv(yi = SMDH_sign, # specify lnRR as the effect size measure;
                    V = VCV_SMDH_flagged, # specify lnRR's sampling variance-covariance matrix;
                    mods = ~ 1,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                    control = list(optimizer="optim"),
                    test = "t",
                    data = SMDH_con_flagged)

# Saving the model
saveRDS(SA_SMDH_flagged, file = here::here("model/sensitivity_analysis/SA_SMDH_flagged.rds"))
# saving the dataset for this model
write.csv(SMDH_con_flagged,here::here("data/05_data_sensitivity_analysis/dataset_SMDH_con_flagged.csv"))

# What is the effect now?

summary(SA_SMDH_flagged)
predict(SA_SMDH_flagged)

```

### Without 0 ES

```{r}

#----------------------------------------------------------------------------
# for lnRR as Effect Size
#----------------------------------------------------------------------------

# Preparing new variance-covariance matrix

VCV_lnRR_no_missingES <- vcalc(vi = lnRR_variance,
                  cluster = paper_ID,
                  obs = Observation_ID,
                  data = lnRR_sin_missingES,
                  rho = 0.5 # assuming that the effect sizes within the same study are correlated with rho = 0.5
                  )

## Running intercept only model

SA_lnRR_no_missingES<- rma.mv(yi = lnRR_sign, # specify lnRR as the effect size measure;
                    V = VCV_lnRR_no_missingES, # specify lnRR's sampling variance-covariance matrix;
                    mods = ~ 1,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                    control = list(optimizer="optim"),
                    test = "t",
                    data = lnRR_sin_missingES)

# Saving the model
saveRDS(SA_lnRR_no_missingES, file = here::here("model/sensitivity_analysis/SA_lnRR_no_missingES.rds"))
# saving the dataset for this model
write.csv(lnRR_sin_missingES,here::here("data/05_data_sensitivity_analysis/dataset_lnRR_sin_missingES.csv"))

# What is the effect now?

summary(SA_lnRR_no_missingES)
predict(SA_lnRR_no_missingES)

#----------------------------------------------------------------------------
# for SMDH as Effect Size
#----------------------------------------------------------------------------

# Preparing new variance-covariance matrix

VCV_SMDH_no_missingES <- vcalc(vi = SMDH_variance,
                  cluster = paper_ID,
                  obs = Observation_ID,
                  data = SMDH_sin_missingES,
                  rho = 0.5 # assuming that the effect sizes within the same study are correlated with rho = 0.5
                  )

## Running intercept only model

SA_SMDH_no_missingES<- rma.mv(yi = SMDH_sign, # specify lnRR as the effect size measure;
                    V = VCV_SMDH_no_missingES, # specify lnRR's sampling variance-covariance matrix;
                    mods = ~ 1,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                    control = list(optimizer="optim"),
                    test = "t",
                    data = SMDH_sin_missingES)

# Saving the model
saveRDS(SA_SMDH_no_missingES, file = here::here("model/sensitivity_analysis/SA_SMDH_no_missingES.rds"))
# saving the dataset for this model
write.csv(SMDH_sin_missingES,here::here("data/05_data_sensitivity_analysis/dataset_SMDH_sin_missingES.csv"))

# What is the effect now?

summary(SA_SMDH_no_missingES)
predict(SA_SMDH_no_missingES)

```

### With flagged proxies and no missing ES

```{r}

#----------------------------------------------------------------------------
# for lnRR as Effect Size
#----------------------------------------------------------------------------

# Preparing new variance-covariance matrix

VCV_flagged_no_missingES <- vcalc(vi = lnRR_variance,
                  cluster = paper_ID,
                  obs = Observation_ID,
                  data = lnRR_flagged_no_missingES,
                  rho = 0.5 # assuming that the effect sizes within the same study are correlated with rho = 0.5
                  )

## Running intercept only model

SA_lnRR_flagged_no_missingES<- rma.mv(yi = lnRR_sign, # specify lnRR as the effect size measure;
                    V = VCV_flagged_no_missingES, # specify lnRR's sampling variance-covariance matrix;
                    mods = ~ 1,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                    control = list(optimizer="optim"),
                    test = "t",
                    data = lnRR_flagged_no_missingES)

# Saving the model
saveRDS(SA_lnRR_flagged_no_missingES, file = here::here("model/sensitivity_analysis/SA_lnRR_flagged_no_missingES.rds"))

# saving the dataset for this model
write.csv(lnRR_flagged_no_missingES,here::here("data/05_data_sensitivity_analysis/dataset_lnRR_flagged_no_missingES.csv"))

# What is the effect now?

summary(SA_lnRR_flagged_no_missingES)
predict(SA_lnRR_flagged_no_missingES)

#----------------------------------------------------------------------------
# for SMDH as Effect Size
#----------------------------------------------------------------------------

# Preparing new variance-covariance matrix

VCV_SMDH_flagged_no_missingES <- vcalc(vi = SMDH_variance,
                  cluster = paper_ID,
                  obs = Observation_ID,
                  data = SMDH_flagged_no_missingES,
                  rho = 0.5 # assuming that the effect sizes within the same study are correlated with rho = 0.5
                  )

## Running intercept only model

SA_SMDH_flagged_no_missingES<- rma.mv(yi = SMDH_sign, # specify lnRR as the effect size measure;
                    V = VCV_SMDH_flagged_no_missingES, # specify lnRR's sampling variance-covariance matrix;
                    mods = ~ 1,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                    control = list(optimizer="optim"),
                    test = "t",
                    data = SMDH_flagged_no_missingES)

# Saving the model
saveRDS(SA_SMDH_flagged_no_missingES, file = here::here("model/sensitivity_analysis/SA_SMDH_flagged_no_missingES.rds"))

# saving the dataset for this model
write.csv(SMDH_flagged_no_missingES,here::here("data/05_data_sensitivity_analysis/dataset_SMDH_flagged_no_missingES.csv"))

# What is the effect now?

summary(SA_SMDH_flagged_no_missingES)
predict(SA_SMDH_flagged_no_missingES)
```

### No inferential stats for SMDH: effect sizes calculated from means, SDs, and sample size

(i.e., excluding any effect sizes calculated from inferential statistics)

```{r}

#----------------------------------------------------------------------------
# for SMDH as Effect Size
#----------------------------------------------------------------------------

# Preparing new variance-covariance matrix

VCV_SMDH_no_inferential <- vcalc(vi = SMDH_variance,
                  cluster = paper_ID,
                  obs = Observation_ID,
                  data = SMDH_no_inferential,
                  rho = 0.5 # assuming that the effect sizes within the same study are correlated with rho = 0.5
                  )

## Running intercept only model

SA_SMDH_no_inferential<- rma.mv(yi = SMDH_sign, # specify lnRR as the effect size measure;
                    V = VCV_SMDH_no_inferential, # specify lnRR's sampling variance-covariance matrix;
                    mods = ~ 1,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                    control = list(optimizer="optim"),
                    test = "t",
                    data = SMDH_no_inferential)

# Saving the model
saveRDS(SA_SMDH_no_inferential, file = here::here("model/sensitivity_analysis/SA_SMDH_no_inferential.rds"))

# saving the dataset for this model
write.csv(SMDH_no_inferential,here::here("data/05_data_sensitivity_analysis/dataset_SMDH_no_inferential.csv"))

# What is the effect now?

summary(SA_SMDH_no_inferential)
predict(SA_SMDH_no_inferential)
```

### Exluding failed geary's test for lnRR

```{r}
#----------------------------------------------------------------------------
# for  lnRR  as Effect Size
#----------------------------------------------------------------------------

VCV_lnRR_geary_passed <- vcalc(vi = lnRR_variance,
                  cluster = paper_ID,
                  obs = Observation_ID,
                  data = lnRR_geary_passed,
                  rho = 0.5 # assuming that the effect sizes within the same study are correlated with rho = 0.5
                  )

SA_lnRR_geary_passed<- rma.mv(yi = lnRR_sign, # specify lnRR as the effect size measure;
                    V = VCV_lnRR_geary_passed, # specify lnRR's sampling variance-covariance matrix;
                    mods = ~ 1,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                    control = list(optimizer="optim"),
                    test = "t",
                    data = lnRR_geary_passed)

# Saving the model
saveRDS(SA_lnRR_geary_passed, file = here::here("model/sensitivity_analysis/SA_lnRR_geary_passed.rds"))

# saving the dataset for this model
write.csv(lnRR_geary_passed,here::here("data/05_data_sensitivity_analysis/dataset_lnRR_geary_passed.csv"))

```

### Excluding the parasite and pathogen load fitness proxies

It excludes all the fitness proxies relating to parasite and pathogen loads/ abundance etc.

```{r}
#----------------------------------------------------------------------------
# for  lnRR  as Effect Size
#----------------------------------------------------------------------------

dataset_lnRR_no_para<-dataset_lnRR%>%filter(is.na(parasite_type))

## The dataset is reduced to almost half
nrow(dataset_lnRR)-nrow(dataset_lnRR_no_para)

VCV_lnRR_no_para <- vcalc(vi = lnRR_variance,
                  cluster = paper_ID,
                  obs = Observation_ID,
                  data = dataset_lnRR_no_para,
                  rho = 0.5 # assuming that the effect sizes within the same study are correlated with rho = 0.5
                  )

SA_lnRR_no_para<- rma.mv(yi = lnRR_sign, # specify lnRR as the effect size measure;
                    V = VCV_lnRR_no_para, # specify lnRR's sampling variance-covariance matrix;
                    mods = ~ 1,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                    control = list(optimizer="optim"),
                    test = "t",
                    data = dataset_lnRR_no_para)

# Saving the model
saveRDS(SA_lnRR_no_para, file = here::here("model/sensitivity_analysis/SA_lnRR_no_para.rds"))

# saving the dataset for this model
write.csv(dataset_lnRR_no_para,here::here("data/05_data_sensitivity_analysis/dataset_lnRR_no_para.csv"))

#----------------------------------------------------------------------------
# for  SMDH  as Effect Size
#----------------------------------------------------------------------------

dataset_SMDH_no_para<-dataset_SMDH%>%filter(is.na(parasite_type))

## The dataset is reduced to almost half
nrow(dataset_SMDH)-nrow(dataset_SMDH_no_para)

VCV_SMDH_no_para <- vcalc(vi = SMDH_variance,
                  cluster = paper_ID,
                  obs = Observation_ID,
                  data = dataset_SMDH_no_para,
                  rho = 0.5) # assuming that the effect sizes within the same study are correlated with rho = 0.5

SA_SMDH_no_para<- rma.mv(yi = SMDH_sign, # specify lnRR as the effect size measure;
                    V = VCV_SMDH_no_para, # specify lnRR's sampling variance-covariance matrix;
                    mods = ~ 1,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                    control = list(optimizer="optim"),
                    test = "t",
                    data = dataset_SMDH_no_para)

# Saving the model
saveRDS(SA_SMDH_no_para, file = here::here("model/sensitivity_analysis/SA_SMDH_no_para.rds"))

# saving the dataset for this model
write.csv(dataset_SMDH_no_para,here::here("data/05_data_sensitivity_analysis/dataset_SMDH_no_para.csv"))

```

## Different values of sampling variance correlation

[**Robust variance estimation to guard against arbitary value of sampling correlation**]{.underline} - 'robust()' function from R package 'metafor' (Viechtbauer 2010) using 'CR2' (bias reduced linearization correction) from the R package 'clubSandwich' (Pustejovsky 2024) for small-sample size adjustment.

At the moment, this is not working..

```{r}
#MLMA_lnRR_RVE<- robust(MLMA_lnRR_VCV,
#                        cluster = paper_ID,
#                        clubSandwich = TRUE,
#                        verbose=T)

# # summary(MLMA_lnRR_RVE)
# I2_CI <- i2_ml(MLMA_lnRR_RVE) # Relative heterogeneity
# round(I2_CI,2) # absolute heterogeneity is reported in the summary of the model
```

::: callout-important
Sensitivity Analysis

We will also perform a sensitivity analysis using 0.3 and 0.7 correlation between the sampling variances from the same paper_ID and report it in the supplementary materials. If this sensitivity analysis shows large differences in conclusions for the intercept-only model, we will proceed to perform such sensitivity analysis for the meta-regressions.
:::

First, set a series of ρs (i.e., 0.3, 0.5, 0.7) (we assume these values arbitrarily):

```{r lnRR rho sensitivity analysis}
# rho_range <- c(0.3, 0.5, 0.7)
# 
# MLMAlnRR_VCV_range <- list() # repeatedly run the specified model with varying rho
# for (i in 1:length(rho_range))
# {VCV_range <- vcalc(vi = lnRR_variance,
#                   cluster = paper_ID,
#                   subgroup = group_ID,
#                   obs = Observation_ID,
#                   data = dataset_lnRR,
#                    rho = rho_range[i],
#                     ) # impute sampling variance covariance matrix with varying rho
# 
# MLMAlnRR_VCV_range[[i]] <- rma.mv(yi = lnRR_sign,
#                                   V = VCV_range, # sampling variance covariance matrix with varying values of rho.
#                                  random = list(~ 1 | paper_ID,
#                                   ~ 1 | Observation_ID,
#                                   ~ 1 | experiment_ID,
#                                   ~ 1 | group_ID,
#                                   ~ 1 | repeated_trait_ID),
#                                   method = "REML",
#                                   data = dataset_lnRR # run the model with varying rho
# )
# }
# 
# save(VCV_range, file = here::here("GNM/figures/VCV_range.Rdata"))
# save(MLMAlnRR_VCV_range, file = here::here("GNM/figures/MLMAlnRR_VCV_range.Rdata"))

```

```{r Table for lnRR rho sensitivity analysis}

# t4 <- data.frame(rho  = rho_range,
#                  "overall effect"  = sapply(MLMAlnRR_VCV_range, function(x) coef(x)),
#                  "standard error" = sapply(MLMAlnRR_VCV_range, function(x) x$se),
#                  "p-value" = sapply(MLMAlnRR_VCV_range, function(x) x$pval),
#                  "Lower CI" = sapply(MLMAlnRR_VCV_range, function(x) x$ci.lb),
#                  "Upper CI" = sapply(MLMAlnRR_VCV_range, function(x) x$ci.ub),
#                  "Log-likehood" = sapply(MLMAlnRR_VCV_range, function(x) fitstats(x)[1,1]),
#                  "AIC" = sapply(MLMAlnRR_VCV_range, function(x) fitstats(x)[3,1]),
#                  "BIC" = sapply(MLMAlnRR_VCV_range, function(x) fitstats(x)[4,1]),
#                  "AICc" = sapply(MLMAlnRR_VCV_range, function(x) fitstats(x)[5,1]))
# 
# colnames(t4) <- c("Sampling correlation", "Overall effect (lnRR)", "Standard error", "p-value", "Lower CI", "Upper CI", "Log-likehood", "AIC", "BIC", "AICc")
# 
# t4 %>% dfround(4) %>% DT::datatable()
```

# Mechanistic Hypothesis

To address our secondary questions "Does the addition of green nest material by birds increase their reproductive success? (Courtship hypothesis)" and "Does the addition of green nest material have protective effects and health benefits for the nestlings? (Parental care hypothesis)", we will make use of the unimoderator multilevel meta-regressions hypothesis type (levels: CH, PCH, Both) as the moderator.

```{r hypothesis-model}
lnRR_hypothesis <-
  metafor::rma.mv(yi = lnRR_sign, # specify lnRR as the effect size measure;
                    V = VCV_lnRR,# specify lnRR's sampling variance;
                    mods = ~ - 1 + Hypothesis,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                  test = "t",
                  control = list(optimizer="optim"),
                  data = dataset_lnRR)
# Save model
saveRDS(lnRR_hypothesis, file = here::here("model/lnRR_hypothesis.rds"))

# saving the dataset for this model
write.csv(dataset_lnRR,here::here("data/04_data_analysis/hypothesis_lnRR.csv"))

lnRR_hypothesis_R2 <- r2_ml(lnRR_hypothesis)
round(lnRR_hypothesis_R2 * 100, 1)

SMDH_hypothesis <-
  metafor::rma.mv(yi = SMDH_sign, # specify lnRR as the effect size measure;
                    V = VCV_SMDH,# specify lnRR's sampling variance;
                    mods = ~ - 1 + Hypothesis,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                  test = "t",
                  control = list(optimizer="optim"),
                  data = dataset_SMDH)

# Save model
saveRDS(SMDH_hypothesis, file = here::here("model/SMDH_hypothesis.rds"))

# saving the dataset for this model
write.csv(dataset_SMDH,here::here("data/04_data_analysis/hypothesis_SMDH.csv"))
```

# Meta-regressions

For understanding the heterogeneity in our intercept-only model, we will run unimoderator multilevel meta-regressions (see section Exploratory Analyses in pre-registration or manuscript) using the following moderators:

## Type of experimental design

(levels: 1 = non-aromatic vs. aromatic, 2 = no added material vs. aromatic, 3 = no added material vs. non-aromatic).

```{r}
# This analysis only includes dataset that has a proxy relating to comaparision type
dataset_lnRR_design <- dataset_lnRR %>% filter(!is.na(comparision_type))

VCV_lnRR_design <- vcalc(vi = lnRR_variance,
                  cluster = paper_ID,
                  obs = Observation_ID,
                  data = dataset_lnRR_design,
                  rho = 0.5 # assuming that the effect sizes within the same study are correlated with rho = 0.5
                  )
lnRR_design <-
  metafor::rma.mv(yi = lnRR_sign, # specify lnRR as the effect size measure;
                    V = VCV_lnRR_design ,# specify lnRR's sampling variance;
                    mods = ~ - 1 + comparision_type,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                  test = "t",
                  control = list(optimizer="optim"),
                  data = dataset_lnRR_design )

# Saving the model
saveRDS(lnRR_design, file = here::here("model/lnRR_design.rds"))

# saving the dataset for this model
write.csv(dataset_lnRR_design,here::here("data/04_data_analysis/design_lnRR.csv"))

# This analysis only includes dataset that has a proxy relating to comaparision type
dataset_SMDH_design <- dataset_SMDH %>% filter(!is.na(comparision_type))

VCV_SMDH_design <- vcalc(vi = SMDH_variance,
                  cluster = paper_ID,
                  obs = Observation_ID,
                  data = dataset_SMDH_design,
                  rho = 0.5 # assuming that the effect sizes within the same study are correlated with rho = 0.5
                  )
SMDH_design <-
  metafor::rma.mv(yi = SMDH_sign, # specify lnRR as the effect size measure;
                    V = VCV_SMDH_design ,# specify lnRR's sampling variance;
                    mods = ~ - 1 + comparision_type,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                  test = "t",
                  control = list(optimizer="optim"),
                  data = dataset_SMDH_design )

# Saving the model
saveRDS(SMDH_design, file = here::here("model/SMDH_design.rds"))

# saving the dataset for this model
write.csv(dataset_SMDH_design,here::here("data/04_data_analysis/design_SMDH.csv"))
```

Let me try to do a sensitivity analysis to see what happens when I remove the control condition of 3 in the overall model since that is expected to not have an effect according to the authors as well.. i.e. it is a double control..

### Excluding non-aromatic and blank comparison (i.e. the two control group comparison)

```{r}

## For the lnRR effect size...
dataset_lnRR_no_comp3<-dataset_lnRR%>%filter(comparision_type!=3)
## Only some rows are removed... 
nrow(dataset_lnRR)-nrow(dataset_lnRR_no_comp3)

VCV_lnRR_no_comp3 <- vcalc(vi = lnRR_variance,
                  cluster = paper_ID,
                  obs = Observation_ID,
                  data = dataset_lnRR_no_comp3,
                  rho = 0.5 # assuming that the effect sizes within the same study are correlated with rho = 0.5
                  )

SA_lnRR_no_comp3<- rma.mv(yi = lnRR_sign, # specify lnRR as the effect size measure;
                    V = VCV_lnRR_no_comp3, # specify lnRR's sampling variance-covariance matrix;
                    mods = ~ 1,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                    control = list(optimizer="optim"),
                    test = "t",
                    data = dataset_lnRR_no_comp3)

# Saving the model
saveRDS(SA_lnRR_no_comp3, file = here::here("model/sensitivity_analysis/SA_lnRR_no_comp3.rds"))

# saving the dataset for this model
write.csv(dataset_lnRR_no_comp3,here::here("data/05_data_sensitivity_analysis/dataset_lnRR_no_comp3.csv"))

#----------------------------------------------------------------------------
# for  SMDH  as Effect Size
#----------------------------------------------------------------------------

dataset_SMDH_no_comp3<-dataset_SMDH%>%filter(comparision_type!=3)
## Only some rows are removed... 
nrow(dataset_SMDH)-nrow(dataset_SMDH_no_comp3)

VCV_SMDH_no_comp3 <- vcalc(vi = SMDH_variance,
                  cluster = paper_ID,
                  obs = Observation_ID,
                  data = dataset_SMDH_no_comp3,
                  rho = 0.5 # assuming that the effect sizes within the same study are correlated with rho = 0.5
                  )

SA_SMDH_no_comp3<- rma.mv(yi = SMDH_sign, # specify lnRR as the effect size measure;
                    V = VCV_SMDH_no_comp3, # specify lnRR's sampling variance-covariance matrix;
                    mods = ~ 1,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                    control = list(optimizer="optim"),
                    test = "t",
                    data = dataset_SMDH_no_comp3)

# Saving the model
saveRDS(SA_SMDH_no_comp3, file = here::here("model/sensitivity_analysis/SA_SMDH_no_comp3.rds"))

# saving the dataset for this model
write.csv(dataset_SMDH_no_comp3,here::here("data/05_data_sensitivity_analysis/dataset_SMDH_no_comp3.csv"))
```

Results remain practically the same irrespective..

## Time of addition of green nest material

(levels: before egg hatching, after egg hatching, continuously throughout the nesting phase).

```{r}
## We use the complete dataset in this case
lnRR_time_gnm_addition <-
  metafor::rma.mv(yi = lnRR_sign, # specify lnRR as the effect size measure;
                    V = VCV_lnRR,# specify lnRR's sampling variance using the variance-covariance matrix we created for the intercept-only model since the whole dataset is being used in this case as well
                    mods = ~ - 1 + time_of_gnm_addition,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                  test = "t",
                  control = list(optimizer="optim"),
                  data = dataset_lnRR)

## Saving the model
saveRDS(lnRR_time_gnm_addition, file = here::here("model/lnRR_time_gnm_addition.rds"))

# saving the dataset for this model
write.csv(dataset_lnRR,here::here("data/04_data_analysis/time_gnm_lnRR.csv"))

## Now similarly for the SMDH effect size using the complete dataset
SMDH_time_gnm_addition <-
  metafor::rma.mv(yi = SMDH_sign, # specify lnRR as the effect size measure;
                    V = VCV_SMDH,# specify lnRR's sampling variance using the variance-covariance matrix we created for the intercept-only model since the whole dataset is being used in this case as well
                    mods = ~ - 1 + time_of_gnm_addition,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                  test = "t",
                  control = list(optimizer="optim"),
                  data = dataset_SMDH)

## Saving the model
saveRDS(SMDH_time_gnm_addition, file = here::here("model/SMDH_time_gnm_addition.rds"))

# saving the dataset for this model
write.csv(dataset_SMDH,here::here("data/04_data_analysis/time_gnm_SMDH.csv"))
```

## Type of parasites

(levels: arthropods, micro-organisms)

```{r Types of parasite}
# This analysis only includes dataset that has a proxy relating to parasite or pathogen
dataset_lnRR_parasite <- dataset_lnRR%>%
  filter(!is.na(parasite_type))%>%
  mutate(parasite_type=as.factor(parasite_type),
         paper_ID=as.factor(paper_ID),
         Observation_ID=as.factor(Observation_ID),
         group_ID_coded=as.factor(group_ID_coded),
         bird_species=as.factor(bird_species))

VCV_lnRR_parasite <- vcalc(vi = lnRR_variance,
                  cluster = paper_ID,
                  obs = Observation_ID,
                  data = dataset_lnRR_parasite,
                  rho = 0.5 # assuming that the effect sizes within the same study are correlated with rho = 0.5
                  )

lnRR_parasite <-
  metafor::rma.mv(yi = lnRR_sign, # specify lnRR as the effect size measure;
                    V = VCV_lnRR_parasite,# specify lnRR's sampling variance;
                    mods = ~ - 1 + parasite_type,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                   test = "t",
                  control = list(optimizer="optim"),
                  data = dataset_lnRR_parasite)

# Saving the model
saveRDS(lnRR_parasite, file = here::here("model/lnRR_parasite.rds"))

# saving the dataset for this model
write.csv(dataset_lnRR_parasite,here::here("data/04_data_analysis/parasite_lnRR.csv"))

### Similiarly for SMDH as well selecting the dataset with parasite type
# This analysis only includes dataset that has a proxy relating to parasite or pathogen
dataset_SMDH_parasite <- dataset_SMDH%>%
  filter(!is.na(parasite_type))
VCV_SMDH_parasite <- vcalc(vi = SMDH_variance,
                  cluster = paper_ID,
                  obs = Observation_ID,
                  data = dataset_SMDH_parasite,
                  rho = 0.5 # assuming that the effect sizes within the same study are correlated with rho = 0.5
                  )

SMDH_parasite <-
  metafor::rma.mv(yi = SMDH_sign, # specify lnRR as the effect size measure;
                    V = VCV_SMDH_parasite,# specify lnRR's sampling variance;
                    mods = ~ - 1 + parasite_type,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                   test = "t",
                  control = list(optimizer="optim"),
                  data = dataset_SMDH_parasite)

# Save the model
saveRDS(SMDH_parasite, file = here::here("model/SMDH_parasite.rds"))

# saving the dataset for this model
write.csv(dataset_SMDH_parasite,here::here("data/04_data_analysis/parasite_SMDH.csv"))
```

#### Bird species

(levels: Cyanistes caeruleus, Sturnus unicolor, Tachycineta bicolor, Sturnus vulgaris; note that these levels reflect the list of species studied in our current database, which may increase after updating our search and/or receiving unpublished data from authors)

```{r}
lnRR_birds <-
  metafor::rma.mv(yi = lnRR_sign, # specify lnRR as the effect size measure;
                    V = VCV_lnRR,# specify lnRR's sampling variance;
                    mods = ~ - 1 + bird_species,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                  test = "t",
                  control = list(optimizer="optim"),
                  data = dataset_lnRR)

# Saving the model
saveRDS(lnRR_birds, file = here::here("model/lnRR_birds.rds"))

# Saving the dataset for this model
write.csv(dataset_lnRR,here::here("data/04_data_analysis/birds_lnRR.csv"))

## Now doing this for SMDH

SMDH_birds <-
  metafor::rma.mv(yi = SMDH_sign, # specify lnRR as the effect size measure;
                    V = VCV_SMDH,# specify lnRR's sampling variance;
                    mods = ~ - 1 + bird_species,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                  test = "t",
                  control = list(optimizer="optim"),
                  data = dataset_SMDH)

# Saving the model
saveRDS(SMDH_birds, file = here::here("model/SMDH_birds.rds"))

# Saving the dataset for this model
write.csv(dataset_SMDH,here::here("data/04_data_analysis/birds_SMDH.csv"))
```

#### Type of trait studied

(levels:physiology, morphology, reproduction, behaviour, parasite and pathogenic load, phenology)

```{r}

lnRR_trait <-
  metafor::rma.mv(yi = lnRR_sign, # specify lnRR as the effect size measure;
                    V = VCV_lnRR,# specify lnRR's sampling variance;
                    mods = ~ - 1 + trait_type,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                  test = "t",
                  control = list(optimizer="optim"),
                  data = dataset_lnRR)

# Saving the model
saveRDS(lnRR_trait, file = here::here("model/lnRR_trait.rds"))

# Saving the dataset for this model
write.csv(dataset_lnRR,here::here("data/04_data_analysis/trait_lnRR.csv"))

## Now similarly for SMDH

SMDH_trait <-
  metafor::rma.mv(yi = SMDH_sign, # specify lnRR as the effect size measure;
                    V = VCV_SMDH,# specify lnRR's sampling variance;
                    mods = ~ - 1 + trait_type,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                  test = "t",
                  control = list(optimizer="optim"),
                  data = dataset_SMDH)

# Saving the model
saveRDS(SMDH_trait, file = here::here("model/SMDH_trait.rds"))

# Saving the dataset for this model
write.csv(dataset_SMDH,here::here("data/04_data_analysis/SMDH_lnRR.csv"))
```

# Risk of Bias

6\. Blinding: (Levels: yes, no(

7\. Random Assignment: (Levels: yes, no)

8\. Missing Data: (Levels: yes, no)

Note that the specific subset of data for each meta-regression may differ based on the data availability (e.g., information about the type of parasites is only available for effect sizes based on parasite/pathogen related fitness proxy). We will conduct meta-regression on suitable subsets of data for each analysis, provided we have sufficient effect sizes per moderator level (i.e., at least 5 effect sizes per moderator level).

::: callout-important
## Phylogenetic Effects

Note also that if we obtain additional effect sizes for other bird species after we update our search and/or contact authors, our models may need to account for phylogenetic nonindependence via the inclusion of two additional random effects: "species" and "phylogeny" following Cinar et al. (2022).
:::

## Other Analysis

### Risk of Bias Assessment

-   RoB - The outcome will be presented in the form of a [**summary table**]{.underline}. We will conduct uni-moderator meta-regressions using these variables.

# Publication Bias

Small study effects:

```{r}

dataset_lnRR_published<-dataset_lnRR_published%>%
  mutate(effective_n_inv=1/((4*effective_n_experiment*effective_n_control)/(effective_n_experiment+effective_n_control)))%>%
  mutate(sqrt_effective_n_inv=sqrt(effective_n_inv))

lnRR_small_study <-
  metafor::rma.mv(yi = lnRR_sign, # specify lnRR as the effect size measure;
                    V = VCV_lnRR_published,# specify lnRR's sampling variance;
                    mods = ~sqrt_effective_n_inv,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                  test = "t",
                  control = list(optimizer="optim"),
                  data = dataset_lnRR_published)

# saving the model
saveRDS(lnRR_small_study, file = here::here("model/lnRR_small_study.rds"))

summary(lnRR_small_study)

round(r2_ml(lnRR_small_study)*100,2)

## Now for SMDH model

dataset_SMDH_published<-dataset_SMDH_published%>%
  mutate(effective_n_inv=1/((4*effective_n_experiment*effective_n_control)/(effective_n_experiment+effective_n_control)))%>%
  mutate(sqrt_effective_n_inv=sqrt(effective_n_inv))

SMDH_small_study <-
  metafor::rma.mv(yi = SMDH_sign, # specify lnRR as the effect size measure;
                    V = VCV_SMDH_published,# specify lnRR's sampling variance;
                    mods = ~sqrt_effective_n_inv,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                  test = "t",
                  control = list(optimizer="optim"),
                  data = dataset_SMDH_published)

# saving the model
saveRDS(SMDH_small_study, file = here::here("model/SMDH_small_study.rds"))

summary(SMDH_small_study)

round(r2_ml(SMDH_small_study)*100,2)

```

Decline effects (Time lag bias)

Next, run a uni-moderator meta-regression to test for evidence of decline effects (also known as time-lag bias) which includes year of publication (mean-centered) as the only moderator, and whose slope would provide information about whether the overall effect size has changed (decline) over time (Nakagawa et al. 2022; Sánchez-Tójar et al. 2018).

```{r}
dataset_lnRR_published<-dataset_lnRR_published%>%
  mutate(year_MeanCenter=scale(as.numeric(year_publication),scale=F)[,1])

lnRR_decline_eff <-
  metafor::rma.mv(yi = lnRR_sign, # specify lnRR as the effect size measure;
                    V = VCV_lnRR_published,# specify lnRR's sampling variance;
                    mods = ~ year_MeanCenter,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                  test = "t",
                  control = list(optimizer="optim"),
                  data = dataset_lnRR_published)

# saving the model
saveRDS(lnRR_decline_eff, file = here::here("model/lnRR_decline_eff.rds"))

summary(lnRR_decline_eff)

round(r2_ml(lnRR_decline_eff)*100,2)

## --- Now for SMDH -----
dataset_SMDH_published<-dataset_SMDH_published%>%
  mutate(year_MeanCenter=scale(as.numeric(year_publication),scale=F)[,1])

SMDH_decline_eff <-
  metafor::rma.mv(yi = SMDH_sign, # specify lnRR as the effect size measure;
                    V = VCV_SMDH_published,# specify lnRR's sampling variance;
                    mods = ~ year_MeanCenter,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                  test = "t",
                  control = list(optimizer="optim"),
                  data = dataset_SMDH_published)
# saving the model
saveRDS(SMDH_decline_eff, file = here::here("model/SMDH_decline_eff.rds"))

summary(SMDH_decline_eff)

round(r2_ml(SMDH_decline_eff)*100,2)

```

Last, run an all-in meta-regression including both the (square root of the) inverse of the 'effective sample size' and the mean-centered year of publication as well as all the other moderators tested in our study to estimate how much heterogeneity is explained by all moderators combined as well as to explore whether evidence remains similar after accounting for all the other moderators (see equation 29 in Nakagawa et al. 2022 for more information about this approach).

## All-in Model for heterogeneity

```{r}
# ## For lnRR complete dataset
# 
# ## I am changing the year of unpublished studies to 2025 which is to now, assuming that if they were to be published now, what would be the effect. I will keep in mind that in this model if I do find an evidence of publication bias over years, I should keep in mind it is non-independent of the unpublished studies
# 
# dataset_lnRR_all_in<-dataset_lnRR%>%mutate(year_publication = case_when(is.na(year_publication) ~ 2025, TRUE ~ year_publication))%>%
#   mutate(year_MeanCenter=scale(as.numeric(year_publication),scale=F)[,1])%>%
#   mutate(effective_n_inv=1/((4*effective_n_experiment*effective_n_control)/(effective_n_experiment+effective_n_control)))%>%
#   mutate(sqrt_effective_n_inv=sqrt(effective_n_inv))%>%
#   mutate(comparision_type= case_when(is.na(comparision_type) ~ 4, TRUE ~ comparision_type))
# 
# VCV_lnRR_all_in<- vcalc(vi = lnRR_variance,
#                   cluster = paper_ID,
#                   obs = Observation_ID,
#                   data = dataset_lnRR_all_in,
#                   rho = 0.5 # assuming that the effect sizes within the same study are correlated with rho = 0.5
#                   )
# 
# lnRR_all_in <- rma.mv(yi = lnRR_sign, # specify lnRR as the effect size measure;
#                     V = VCV_lnRR_all_in, # specify lnRR's sampling variance-covariance matrix;
#                     mods = ~ time_of_gnm_addition + 
#                       year_MeanCenter + 
#                       sqrt_effective_n_inv + 
#                       as.factor(comparision_type) + 
#                       bird_species + 
#                       trait_type,
#                     random = list(~ 1 | paper_ID,
#                                   ~ 1 | Observation_ID,
#                                   ~ 1 | group_ID_coded,
#                                   ~ 1 | population_ID,
#                                   ~ 1 | bird_species),
#                     method = "REML",
#                     control = list(optimizer="optim"),
#                     test = "t",
#                     data = dataset_lnRR_all_in)
# 
# 
# 
# ## For SMDH complete published dataset
# 
# dataset_SMDH_all_in<-dataset_SMDH%>%mutate(year_publication = case_when(is.na(year_publication) ~ 2025, TRUE ~ year_publication))%>%
#   mutate(year_MeanCenter=scale(as.numeric(year_publication),scale=F)[,1])%>%
#   mutate(effective_n_inv=1/((4*effective_n_experiment*effective_n_control)/(effective_n_experiment+effective_n_control)))%>%
#   mutate(sqrt_effective_n_inv=sqrt(effective_n_inv))%>%
#   mutate(comparision_type= case_when(is.na(comparision_type) ~ 4, TRUE ~ comparision_type))
# 
# VCV_SMDH_all_in<- vcalc(vi = SMDH_variance,
#                   cluster = paper_ID,
#                   obs = Observation_ID,
#                   data = dataset_SMDH_all_in,
#                   rho = 0.5 # assuming that the effect sizes within the same study are correlated with rho = 0.5
#                   )
# 
# SMDH_all_in<- rma.mv(yi = SMDH_sign, # specify lnRR as the effect size measure;
#                     V = VCV_SMDH_all_in, # specify lnRR's sampling variance-covariance matrix;
#                     mods = ~ time_of_gnm_addition + 
#                       year_MeanCenter + 
#                       sqrt_effective_n_inv + 
#                       as.factor(comparision_type) + 
#                       bird_species + 
#                       trait_type,
#                     random = list(~ 1 | paper_ID,
#                                   ~ 1 | Observation_ID,
#                                   ~ 1 | group_ID_coded,
#                                   ~ 1 | population_ID,
#                                   ~ 1 | bird_species),
#                     method = "REML",
#                     control = list(optimizer="optim"),
#                     test = "t",
#                     data = dataset_SMDH_all_in)
```

# Pair-wise comparison for meta-regression

including including post-hoc Wald tests

For this I have to run all the meta-regressions again without removing the intercepts, good luck to me!

```{r}
#----------------------------------------------------------------------------------
# For Different hypothesis
#----------------------------------------------------------------------------------

# lnRR_hypothesis_pc <-
#   metafor::rma.mv(yi = lnRR_sign, # specify lnRR as the effect size measure;
#                     V = VCV_lnRR,# specify lnRR's sampling variance;
#                     mods = ~ Hypothesis,
#                     random = list(~ 1 | paper_ID,
#                                   ~ 1 | Observation_ID,
#                                   ~ 1 | group_ID_coded,
#                                   ~ 1 | population_ID,
#                                   ~ 1 | bird_species),
#                     method = "REML",
#                   test = "t",
#                   control = list(optimizer="optim"),
#                   data = dataset_lnRR)
# 
# # I now want to test for statistical differences between groups. Male and fem
# # information for androgens can be obtained from the outcome of the table. For thyroid
# # hormones I need to do Post-hoc Wald tests: 
# car::linearHypothesis(lnRR_hypothesis_pc, rbind(c(0,1,-1)))
```