### 2025-07-09 code/08_figures-tables.Rmd
No build, pushing to switch machines :D

---
title: "Figures"
author: "Shreya D."
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# ---- FIGURES --------------

# Visualising location and bird species

I am using the figure code from here and adapting it for my needs..

Mentesana, L., Hau, M., D'Amelio, P.B., Adreani, N.M. and Sánchez-Tójar, A. (2025), Do Egg Hormones Have Fitness Consequences in Wild Birds? A Systematic Review and Meta-Analysis. Ecology Letters, 28: e70100. <https://doi.org/10.1111/ele.70100>

Code here: `https://github.com/ASanchez-Tojar/meta-analysis_egg_hormones_and_fitness/blob/main/code/003_figures.R`

```{r message=FALSE, warning=FALSE}
## Let's get the needed packages
if (!requireNamespace("devtools", quietly = TRUE))
    install.packages("devtools")

if (!requireNamespace("orchaRd", quietly = TRUE))
    devtools::install_github("daniel1noble/orchaRd", ref = "main")

pacman::p_load(ggplot2,dplyr,stringr,readr,cowplot,maps)

## Let's first get the data in for this plot

# Loading cleaned dataset
dataset_analysis<- read_csv(here::here("data/04_data_analysis/dataset_analysis.csv"))

```

```{r location-map}

# To split population_ID into location and country
dataset_analysis[c('Location','Country')] <- str_split_fixed(dataset_analysis$population_ID, ',', 2)

### For the purpose of this map, I am adding the the longitude and latitude of the study area manually. They exist in the dataset_analysis$population_location as they were extracted from the paper. I have also added some by just searching the general location since the coordinates were not provided for these..

data_coordinates<-data.frame(Location = unique(dataset_analysis$Location),
                             Lat_Location=c(51.50,38.41,40.45,
                                            45.40,52.39, 37.18,
                                            54,39.05,
                                            51.57,48,42.54,39.95,
                                            40.13,28.10,52.06,60.39),
                             Long_Location=c(19.29,-9.02,-3.50,
                                             -64.10,6.05,-3.11,
                                             -124,-76.81,
                                             -2.26,11,8.91,-75.16,
                                             -8.27,107.10,8.25,5.32))

## Let's now merge the datasets by the location names

dataset_analysis <- dataset_analysis %>%
  left_join(data_coordinates, by = "Location")

## Count how many times each Location appears in the full dataset
location_counts <- dataset_analysis %>%
  group_by(Location) %>%
  summarise(size.freq = n_distinct(paper_ID), .groups = 'drop')

# I will subset the data base so that I get one row per Location AND bird_species.
# Like this, I get one point per location independently of the number of effect sizes 
# we have per place.
dataset_analysis_unique_rows <- dataset_analysis %>%
  distinct(Location, bird_species, .keep_all = TRUE)

# Merge the counts into the unique rows dataset
dataset_analysis_unique_rows <- dataset_analysis_unique_rows %>%
  left_join(location_counts, by = "Location")

# create data for world coordinates using  
# map_data() function 
world_coordinates <- map_data("world")

# create world map using ggplot() function 
# geom_map() function takes world coordinates as input to plot world map 
fig_map <- ggplot(world_coordinates, aes(x = long, y = lat)) + 
  geom_map(map = world_coordinates, 
    aes(map_id = region), 
    color = "white", fill= "#D9EAF7") + 
  geom_point(data = dataset_analysis_unique_rows, 
             aes(x = Long_Location, y = Lat_Location,
                 size = size.freq, colour=bird_species), alpha = 0.6) +
scale_size(range = c(3, 5)) +
  scale_fill_manual(values = c("#00B0BE","#FFB255",  "#FF8CAE", 
                                 "#F47A00", "#98C127","#003A7D","#7E4794")) +

  scale_colour_manual(values = c("#00B0BE","#FFB255",  "#FF8CAE", 
                                 "#F47A00", "#98C127","#003A7D","#7E4794")) +
  guides(size = "none", alpha = "none") +
  labs(x = "Longitude",
       y = "Latitude",
       col = "Bird Species")+
  theme_classic() +
  theme(legend.position = c(0.99, 0.15),
        legend.justification = c("right", "bottom"),
        legend.background = element_rect(fill = "white", color = "black"),
        legend.text = element_text(size=6),
        axis.title = element_text(size = 10),
        axis.text.x = element_text(size = 10),
        axis.text.y = element_blank())

# To add the zoomed in map.. to show the points more clearly
zoom_map <- ggplot() + 
  geom_map(data = world_coordinates, map = world_coordinates,
           aes(map_id = region),
           color = "white", fill = "#D4D4D4") +
  geom_point(data = dataset_analysis_unique_rows,
             aes(x = Long_Location, y = Lat_Location,
                 size = size.freq, colour = bird_species),
             alpha = 0.9) +
  scale_size(range = c(2, 4)) +
  scale_colour_manual(values = c("#00B0BE","#FFB255",  "#FF8CAE", 
                                 "#F47A00", "#98C127","#003A7D","#7E4794")) +
  coord_cartesian(xlim = c(-10, 20), ylim = c(25, 65)) +  # Northern Hemisphere
  theme_void() +
  theme(
    legend.position = "none",
    plot.background = element_rect(color = "white", fill="white"),
    panel.border = element_rect(colour = "black", fill = NA, size = 1)
  )

geographic_distribution <- ggdraw() +
  draw_plot(fig_map) +
  draw_plot(zoom_map, x = 0.05, y = 0.15, width = 0.28, height = 0.38) + theme(panel.border = element_rect(colour = "black", fill=NA, size=1.5))

ggsave(here::here("figures/Figure_Distribution.png"), plot = geographic_distribution, height = 100, width = 170, units= 'mm', dpi = 600)
```

# Visualizing author network

```{r}
# # Loading cleaned dataset
# dataset_analysis<- read_csv(here::here("data/03_data_cleaning/dataset_analysis.csv"))
# 
# library(tidyverse)
# library(tidyr)
# library(igraph)
# installed.packages("ggraph")
# library(ggraph)
# 
# clean_author <- function(name) {
#   name %>%
#     str_replace_all(",", "") %>%   # remove commas
#     str_squish() %>%              # remove excess whitespace
#     str_to_title()                # standardize case (optional, if needed)
# }
# 
# # Apply cleaning to first and last author fields
# author_roles <- dataset_analysis %>%
#   distinct(paper_ID, .keep_all = TRUE) %>%
#   select(paper_ID, authors, bird_species) %>%
#   mutate(
#     author_list = str_split(authors, ";\\s*"),
#     first_author = map_chr(author_list, 1),
#     last_author = map_chr(author_list, ~ tail(.x, 1))
#   )%>%
#   mutate(
#     first_author=case_when(
#       first_author == "Polo, V" | first_author == "Polo V" ~ "Polo V.",
#       first_author == "Soler J J" ~ "Soler J. J.",
#       first_author == "Glądalski, M" ~ "Gladalski M.",
#       first_author == "Gwinner, H" | first_author == "Gwinner H" ~ "Gwinner H.",
#       TRUE ~ as.character(first_author)
#     ))%>%
#   mutate(
#     last_author=case_when(
#       last_author == "Veiga J P" ~ "Veiga J. P.",
#       TRUE ~ as.character(last_author)
#     )
#   )
# 
# # 2. Create Core Author Pairs (First → Last)
# author_pairs_core <- author_roles %>%
#   select(paper_ID, first_author, last_author) %>%
#   rename(from = first_author, to = last_author) %>%
#   filter(from != to) %>%
#   count(from, to, name = "weight")
# 
# # 3. Build Author Graph
# author_graph <- graph_from_data_frame(author_pairs_core, directed = TRUE)
# 
# # 4. Assign Dominant Bird Species to PIs (last authors)
# author_species <- author_roles %>%
#   select(last_author, bird_species) %>%
#   group_by(last_author, bird_species) %>%
#   count() %>%
#   slice_max(n, n = 1) %>%
#   distinct(last_author, .keep_all = TRUE)
# 
# # Add species as vertex attribute
# V(author_graph)$bird_species <- author_species$bird_species[
#   match(V(author_graph)$name, author_species$last_author)
# ]
# 
# # 5. Optional: Scale node size by number of papers per author
# author_paper_counts <- author_roles %>%
#   pivot_longer(cols = c(first_author, last_author), names_to = "role", values_to = "author") %>%
#   count(author, name = "paper_count")
# 
# # Add paper count attribute
# V(author_graph)$paper_count <- author_paper_counts$paper_count[
#   match(V(author_graph)$name, author_paper_counts$author)
# ]
# 
# # Replace NAs with small size for isolated authors
# V(author_graph)$paper_count[is.na(V(author_graph)$paper_count)] <- 1
# 
# # 6. Plot with ggraph
# ggraph(author_graph, layout = "fr") +
#   geom_edge_link(aes(width = weight), alpha = 0.4, color = "grey50") +
#   geom_node_point(aes(color = bird_species, size = paper_count)) +
#   geom_node_text(aes(label = name), repel = TRUE, size = 3) +
#   scale_edge_width(range = c(0.3, 2)) +
#   scale_size(range = c(3, 10)) +
#   theme_void() +
#   labs(title = "Author Collaboration Network by Bird Species",
#        size = "Paper Count") +
#   theme(legend.position = "right")
#

```

# Intercept only model

Orchard plot for overall effect of green nest material on fitness using lnRR and SMDH

```{r orchard plot for intercept only models}
pacman::p_load(here,tidyverse,metafor,orchaRd,patchwork,cowplot)

# Loading models for SMDH and lnRR

intercept_lnRR<-readRDS(here::here("model/intercept_lnRR.rds"))
intercept_SMDH<-readRDS(here::here("model/intercept_SMDH.rds"))

# Plot for lnRR

intercept_lnRR_plot<- orchaRd::orchard_plot(intercept_lnRR,
                                             mod = "1",
                                             group = "paper_ID",
                                             trunk.size = 0.3,
                                             branch.size = 2,
                                             twig.size = 0.5,
                                             alpha = 0.2,
                                             xlab = "Effect size (lnRR)", 
                                             fill =  T) +
  scale_fill_manual(values = "#98C127") +
  scale_colour_manual(values = "#98C127" ) +
theme_minimal(base_size = 14) +
  theme(
    panel.border = element_rect(colour = "black", fill = NA, size = 0.8),
    legend.position = c(0.98, 0.05),
    legend.justification = c(1, 0),
    legend.direction = "horizontal",
    legend.title = element_text(size = 7),
    legend.text = element_text(size = 7),
    axis.title = element_text(size = 10),
    axis.text.x = element_text(size = 10),
    axis.text.y = element_blank()
  ) +
  labs(x = "Intercept")

# Plot for SMDH
intercept_SMDH_plot <- orchaRd::orchard_plot(intercept_SMDH,
                                    mod = "1",
                                    group = "paper_ID",
                                    trunk.size = 0.3,
                                    branch.size = 2,
                                    twig.size = 0.5,
                                    alpha = 0.2,
                                    xlab = "Effect size (SMDH)",
                                    fill = TRUE) +
  scale_fill_manual(values = "#F47A00") +
  scale_colour_manual(values = "#F47A00") +
  theme_minimal(base_size = 14) +
  theme(
    panel.border = element_rect(colour = "black", fill = NA, size = 0.8),
    legend.position = c(0.98, 0.05),
    legend.justification = c(1, 0), 
    legend.direction = "horizontal",
    legend.title = element_text(size = 7),
    legend.text = element_text(size = 7),
    axis.title = element_text(size = 10),
    axis.text.x = element_text(size = 10),
    axis.text.y = element_blank()
  ) +
  labs(x = "Intercept")

# Combine using patchwork
Figure_intercept <- intercept_lnRR_plot / intercept_SMDH_plot +
  plot_annotation(tag_level = c("A","B"))

ggsave(here::here("figures/Figure_overall_effect.png"), plot = Figure_intercept, height = 150, width = 170, units = "mm", dpi = 600)

###----------------------------------------------------------------------------------

## Now the cropped versions presented in the main manuscript..

###----------------------------------------------------------------------------------

intercept_lnRR_plot_crop <- orchaRd::orchard_plot(intercept_lnRR,
                                             mod = "1",
                                             group = "paper_ID",
                                             trunk.size = 0.5,
                                             branch.size = 2,
                                             twig.size = 0.5,
                                             alpha = 0.2,
                                             xlab = "Effect size (lnRR)", 
                                             fill =  T) +
  scale_fill_manual(values = "#98C127") +
  scale_colour_manual(values = "#98C127" ) +
    scale_y_continuous(limits = c(-0.7, 0.7)) +
theme_minimal(base_size = 14) +
  theme(
    panel.border = element_rect(colour = "black", fill = NA, size = 0.8),
    legend.position = c(0.98, 0.05),
    legend.justification = c(1, 0),
    legend.direction = "horizontal",
    legend.title = element_text(size = 7),
    legend.text = element_text(size = 7),
    axis.title = element_text(size = 10),
    axis.text.x = element_text(size = 10),
    axis.text.y = element_blank()
  ) +
  labs(x = "Intercept")

# Plot for SMDH
intercept_SMDH_plot_crop <- orchaRd::orchard_plot(intercept_SMDH,
                                    mod = "1",
                                    group = "paper_ID",
                                    trunk.size = 0.5,
                                    branch.size = 2,
                                    twig.size = 0.5,
                                    alpha = 0.2,
                                    xlab = "Effect size (SMDH)",
                                    fill = TRUE) +
  scale_fill_manual(values = "#F47A00") +
  scale_colour_manual(values = "#F47A00") +
  scale_y_continuous(limits = c(-2.7, 2.7)) +
  theme_minimal(base_size = 14) +
  theme(
    panel.border = element_rect(colour = "black", fill = NA, size = 0.8),
    legend.position = c(0.98, 0.05),
    legend.justification = c(1, 0), 
    legend.direction = "horizontal",
    legend.title = element_text(size = 7),
    legend.text = element_text(size = 7),
    axis.title = element_text(size = 10),
    axis.text.x = element_text(size = 10),
    axis.text.y = element_blank()
  ) +
  labs(x = "Intercept")

# Combine using patchwork
Figure_intercept_crop <- intercept_lnRR_plot_crop / intercept_SMDH_plot_crop +
  plot_annotation(tag_level = c("A","B"))

ggsave(here::here("figures/Figure_overall_effect_crop.png"), plot = Figure_intercept_crop, height = 150, width = 170, units = "mm", dpi = 600)
```

# Hypothesis meta-regression

```{r orchard plot hypothesis meta-regression}

# Loading models for SMDH and lnRR

lnRR_hypothesis<-readRDS(here::here("model/lnRR_hypothesis.rds"))
SMDH_hypothesis<-readRDS(here::here("model/SMDH_hypothesis.rds"))

# Plot lnRR meta-regression on hypothesis

lnRR_hypothesis_plot <-
  orchard_plot(mod_results(lnRR_hypothesis,
                           mod = "Hypothesis",
                           group = "paper_ID"),
               trunk.size = 0.25,
               branch.size = 1,
               twig.size = 0.5,
               alpha = 0.4,
               xlab ="") +
  scale_colour_manual(values = c("#FF8CAE", "#003A7D", "#7E4794")) +
  scale_fill_manual(values = c("#FF8CAE", "#003A7D", "#7E4794")) +
  labs(x = "",
       y = "Effect size (lnRR)")+
  theme(
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.8),
    legend.position.inside = c(0.99, 0.05),
    legend.justification = c(1, 0), 
    legend.direction = "horizontal",
    legend.title = element_text(size = 6),
    legend.text = element_text(size = 6),
    axis.title = element_text(size = 10),
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(angle = 0, vjust = 0.5, hjust = 1))+ 
  scale_x_discrete(labels = c("Both","Courtship","Parental Care"))

# Plot SMDH meta-regression on hypothesis

SMDH_hypothesis_plot <-
  orchard_plot(
    mod_results(
      SMDH_hypothesis,
      mod = "Hypothesis",
      group = "paper_ID"
    ),
    trunk.size = 0.25,
    branch.size = 1,
    twig.size = 0.5,
    alpha = 0.4,
    xlab ="") +
  scale_colour_manual(values = c("#FF8CAE", "#003A7D", "#7E4794")) +
  scale_fill_manual(values = c("#FF8CAE", "#003A7D", "#7E4794")) +
  labs(x = "",
       y = "Effect size (SMDH)")+
    theme(
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.8),
    legend.position.inside = c(0.99, 0.05),
    legend.justification = c(1, 0), 
    legend.direction = "horizontal",
    legend.title = element_text(size = 6),
    legend.text = element_text(size = 6),
    axis.title = element_text(size = 10),
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(angle = 0, vjust = 0.5, hjust = 1))+ 
  scale_x_discrete(labels = c("Both","Courtship","Parental Care"))

SMDH_hypothesis_plot <- SMDH_hypothesis_plot +
  theme(axis.text.y  = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank())

# Combine using patchwork
Figure_hypothesis <- lnRR_hypothesis_plot + SMDH_hypothesis_plot +
  plot_annotation(tag_level = c("A","B"))

ggsave(here::here("figures/Figure_hypothesis.png"), plot = Figure_hypothesis, height = 120, width = 250, units = "mm", dpi = 600)

##-----------------------------------------------------------------------------

# Now the cropped figures for the main manuscript

##-----------------------------------------------------------------------------

# Plot lnRR meta-regression on hypothesis

lnRR_hypothesis_plot_crop <-
  orchard_plot(mod_results(lnRR_hypothesis,
                           mod = "Hypothesis",
                           group = "paper_ID"),
               trunk.size = 0.5,
               branch.size = 1.3,
               twig.size = 0.5,
               alpha = 0.4,
               fill = T,
               xlab ="") +
  scale_colour_manual(values = c("#FF8CAE", "#003A7D", "#7E4794")) +
  scale_fill_manual(values = c("#FF8CAE", "#003A7D", "#7E4794")) +
    scale_y_continuous(limits = c(-0.7, 0.7)) +
  labs(x = "",
       y = "Effect size (lnRR)")+
  theme(
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.8),
    legend.position.inside = c(0.99, 0.05),
    legend.justification = c(1, 0), 
    legend.direction = "horizontal",
    legend.title = element_text(size = 6),
    legend.text = element_text(size = 6),
    axis.title = element_text(size = 10),
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(angle = 0, vjust = 0.5, hjust = 1))+ 
  scale_x_discrete(labels = c("Both","Courtship","Parental\nCare"))

# Plot SMDH meta-regression on hypothesis

SMDH_hypothesis_plot_crop <-
  orchard_plot(
    mod_results(
      SMDH_hypothesis,
      mod = "Hypothesis",
      group = "paper_ID"
    ),
    trunk.size = 0.5,
    branch.size = 1.3,
    twig.size = 0.5,
    alpha = 0.4,
    fill = T,
    xlab ="") +
  scale_colour_manual(values = c("#FF8CAE", "#003A7D", "#7E4794")) +
  scale_fill_manual(values = c("#FF8CAE", "#003A7D", "#7E4794")) +
  scale_y_continuous(limits = c(-2.7, 2.7)) +
  labs(x = "",
       y = "Effect size (SMDH)")+
  theme(
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.8),
    legend.position.inside = c(0.99, 0.05),
    legend.justification = c(1, 0), 
    legend.direction = "horizontal",
    legend.title = element_text(size = 6),
    legend.text = element_text(size = 6),
    axis.title = element_text(size = 10),
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(angle = 0, vjust = 0.5, hjust = 1))+ 
  scale_x_discrete(labels = c("Both","Courtship","Parental\nCare"))

SMDH_hypothesis_plot_crop <- SMDH_hypothesis_plot_crop +
  theme(axis.text.y  = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank())

# Combine using patchwork
Figure_hypothesis_crop <- lnRR_hypothesis_plot_crop + SMDH_hypothesis_plot_crop +
  plot_annotation(tag_level = c("A","B"))

ggsave(here::here("figures/Figure_hypothesis_crop.png"), plot = Figure_hypothesis_crop, height = 120, width = 250, units = "mm", dpi = 600)
```

# Other meta-regressions

# A. Combined Plot

## Time of green nest material addition

```{r orchard plot time of gnm addition meta-regression}

# Loading models for SMDH and lnRR

lnRR_time_gnm_addition<-readRDS(here::here("model/lnRR_time_gnm_addition.rds"))
SMDH_time_gnm_addition<-readRDS(here::here("model/SMDH_time_gnm_addition.rds"))

lnRR_time_gnm_addition_plot <-
  orchard_plot(mod_results(
      lnRR_time_gnm_addition,
      mod = "time_of_gnm_addition",
      group = "paper_ID"
    ),
    trunk.size = 0.25,
    branch.size = 1,
    twig.size = 0.5,
    alpha = 0.2,
    xlab ="") +
 scale_colour_manual(values = c("#FF8CAE", "#003A7D", "#7E4794")) +
  scale_fill_manual(values = c("#FF8CAE", "#003A7D", "#7E4794")) +
  labs(x = "",
       y = "lnRR")+
    theme(
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.8),
    legend.position = "bottom",
    # legend.position.inside = c(0.98, 0.05),
    # legend.justification = c(1, 0), 
    # legend.direction = "horizontal",
    legend.title = element_text(size = 6),
    legend.text = element_text(size = 6),
    axis.title = element_text(size = 10),
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(angle = 0, vjust = 0.5, hjust = 1)
  ) + scale_x_discrete(labels = c("After egg laying", "Before egg laying", "Continous addition"))

SMDH_time_gnm_addition_plot <-
  orchard_plot(mod_results(
      SMDH_time_gnm_addition,
      mod = "time_of_gnm_addition",
      group = "paper_ID"
    ),
    trunk.size = 0.25,
    branch.size = 1,
    twig.size = 0.5,
    alpha = 0.2,
    xlab ="") +
 scale_colour_manual(values = c("#FF8CAE", "#003A7D", "#7E4794")) +
  scale_fill_manual(values = c("#FF8CAE", "#003A7D", "#7E4794")) +
  labs(x = "",
       y = "SMD(H)")+
    theme(
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.8),
    legend.position = "bottom",
    # legend.position.inside = c(0.98, 0.05),
    # legend.justification = c(1, 0), 
    # legend.direction = "horizontal",
    legend.title = element_text(size = 6),
    legend.text = element_text(size = 6),
    axis.title = element_text(size = 10),
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(angle = 0, vjust = 0.5, hjust = 1)
  )

```

## Experimental Design

```{r orchard plot experimental design meta-regression}

# Loading models for SMDH and lnRR

lnRR_design<-readRDS(here::here("model/lnRR_design.rds"))
SMDH_design<-readRDS(here::here("model/SMDH_design.rds"))

lnRR_design_plot <-
  orchard_plot(mod_results(
      lnRR_design,
      mod = "comparision_type",
      group = "paper_ID"
    ),
    trunk.size = 0.25,
    branch.size = 1,
    twig.size = 0.5,
    alpha = 0.2,
    xlab ="") +
 scale_colour_manual(values = c("#FF8CAE", "#003A7D", "#7E4794")) +
  scale_fill_manual(values = c("#FF8CAE", "#003A7D", "#7E4794")) +
  labs(x = "",
       y = "lnRR")+
    theme(
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.8),
    legend.position = "none",
    # legend.position.inside = c(0.98, 0.05),
    # legend.justification = c(1, 0),
    # legend.direction = "horizontal",
    legend.title = element_text(size = 6),
    legend.text = element_text(size = 6),
    axis.title = element_text(size = 10),
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(angle = 0, vjust = 0.5, hjust = 1)
  ) + scale_x_discrete(labels = c("Non-aromatic Vs. Aromatic","No material Vs. Aromatic", "No material Vs. Non-Aromatic"))

SMDH_design_plot <-
  orchard_plot(mod_results(
      SMDH_design,
      mod = "comparision_type",
      group = "paper_ID"
    ),
    trunk.size = 0.25,
    branch.size = 1,
    twig.size = 0.5,
    alpha = 0.2,
    xlab ="") +
 scale_colour_manual(values = c("#FF8CAE", "#003A7D", "#7E4794")) +
  scale_fill_manual(values = c("#FF8CAE", "#003A7D", "#7E4794")) +
  labs(x = "",
       y = "SMD(H)")+
    theme(
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.8),
    legend.position = "none",
    # legend.position.inside = c(0.98, 0.05),
    # legend.justification = c(1, 0),
    # legend.direction = "horizontal",
    legend.title = element_text(size = 6),
    legend.text = element_text(size = 6),
    axis.title = element_text(size = 10),
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(angle = 0, vjust = 0.5, hjust = 1)
  )

```

### Now we combine these plots

```{r combining meta-regression plots, effect}

## I am going to remove the y ticks from the SMDH plots to make it smaller..
SMDH_design_plot <- SMDH_design_plot +
  theme(axis.text.y  = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank())

SMDH_time_gnm_addition_plot <- SMDH_time_gnm_addition_plot +
  theme(axis.text.y  = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank())

# Combine using patchwork
Figure_design_time_gnm_addition <- ((lnRR_design_plot + SMDH_design_plot) /
                               (lnRR_time_gnm_addition_plot + SMDH_time_gnm_addition_plot)) +
  plot_annotation(tag_levels = "A")

ggsave(here::here("figures/Figure_design_time_gnm_addition.png"), plot = Figure_design_time_gnm_addition, height = 240, width = 250, units = "mm", dpi = 600)

```

# B. Combined Plot

## Parasite Type

```{r orchard plot parasite type meta-regression}

# Loading models for SMDH and lnRR

lnRR_parasite<-readRDS(here::here("model/lnRR_parasite.rds"))
SMDH_parasite<-readRDS(here::here("model/SMDH_parasite.rds"))

lnRR_parasite_plot <-
  orchard_plot(mod_results(
      lnRR_parasite,
      mod = "parasite_type",
      group = "paper_ID"
    ),
    trunk.size = 0.25,
    branch.size = 1,
    twig.size = 0.5,
    alpha = 0.2,
    xlab ="") +
  scale_colour_manual(values = c("#FF8CAE", "#003A7D")) +
  scale_fill_manual(values = c("#FF8CAE", "#003A7D")) +
  labs(x = "",
       y = "lnRR")+
    theme(
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.8),
    legend.position = "none",
    # legend.position.inside = c(0.98, 0.05),
    # legend.justification = c(1, 0), 
    # legend.direction = "horizontal",
    legend.title = element_text(size = 6),
    legend.text = element_text(size = 6),
    axis.title = element_text(size = 10),
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size =6,angle = 0, vjust = 0.5, hjust = 1)
  )

SMDH_parasite_plot <-
  orchard_plot(mod_results(
      SMDH_parasite,
      mod = "parasite_type",
      group = "paper_ID"
    ),
    trunk.size = 0.25,
    branch.size = 1,
    twig.size = 0.5,
    alpha = 0.2,
    xlab ="") +
  scale_colour_manual(values = c("#FF8CAE", "#003A7D")) +
  scale_fill_manual(values = c("#FF8CAE", "#003A7D")) +
  labs(x = "",
       y = "SMD(H)")+
    theme(
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.8),
    legend.position = "none",
    # legend.position.inside = c(0.98, 0.05),
    # legend.justification = c(1, 0), 
    # legend.direction = "horizontal",
    legend.title = element_text(size = 6),
    legend.text = element_text(size = 6),
    axis.title = element_text(size = 10),
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size =6,angle = 0, vjust = 0.5, hjust = 1)
  )

```

## Bird Species

```{r orchard plot bird species meta-regression}

# Loading models for SMDH and lnRR

lnRR_birds<-readRDS(here::here("model/lnRR_birds.rds"))
SMDH_birds<-readRDS(here::here("model/SMDH_birds.rds"))

lnRR_birds_plot <-
  orchard_plot(mod_results(
      lnRR_birds,
      mod = "bird_species",
      group = "paper_ID"
    ),
    trunk.size = 0.25,
    branch.size = 1,
    twig.size = 0.5,
    alpha = 0.2,
    xlab ="") +
  scale_colour_manual(values =  c("#00B0BE","#FFB255",  "#FF8CAE", 
                                 "#F47A00", "#98C127","#003A7D","#7E4794")) +
  scale_fill_manual(values =  c("#00B0BE","#FFB255",  "#FF8CAE", 
                                 "#F47A00", "#98C127","#003A7D","#7E4794")) +
  labs(x = "",
       y = "lnRR")+
    theme(
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.8),
    legend.position = "none",
    # legend.position.inside = c(0.98, 0.05),
    # legend.justification = c(1, 0), 
    # legend.direction = "horizontal",
    legend.title = element_text(size = 6),
    legend.text = element_text(size = 6),
    axis.title = element_text(size = 10),
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size =6,angle = 0, vjust = 0.5, hjust = 1)
  )

SMDH_birds_plot <-
  orchard_plot(mod_results(
      SMDH_birds,
      mod = "bird_species",
      group = "paper_ID"
    ),
    trunk.size = 0.25,
    branch.size = 1,
    twig.size = 0.5,
    alpha = 0.2,
    xlab ="") +
  scale_colour_manual(values =  c("#00B0BE","#FFB255",  "#FF8CAE", 
                                 "#F47A00", "#98C127","#003A7D","#7E4794")) +
  scale_fill_manual(values =  c("#00B0BE","#FFB255",  "#FF8CAE", 
                                 "#F47A00", "#98C127","#003A7D","#7E4794")) +
  labs(x = "",
       y = "SMD(H)")+
    theme(
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.8),
    legend.position = "none",
    # legend.position.inside = c(0.98, 0.05),
    # legend.justification = c(1, 0), 
    # legend.direction = "horizontal",
    legend.title = element_text(size = 6),
    legend.text = element_text(size = 6),
    axis.title = element_text(size = 10),
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size =6,angle = 0, vjust = 0.5, hjust = 1)
  )
```

## Trait studied

```{r orchard plot trait-studied meta-regression}

# Loading models for SMDH and lnRR

lnRR_trait<-readRDS(here::here("model/lnRR_trait.rds"))
SMDH_trait<-readRDS(here::here("model/SMDH_trait.rds"))

lnRR_trait_plot <-
  orchard_plot(mod_results(
      lnRR_trait,
      mod = "trait_type",
      group = "paper_ID"
    ),
    trunk.size = 0.25,
    branch.size = 1,
    twig.size = 0.5,
    alpha = 0.2,
    xlab ="") +
  scale_colour_manual(values =  c("#00B0BE","#FFB255",  "#FF8CAE", 
                                 "#F47A00", "#98C127","#003A7D","#7E4794")) +
  scale_fill_manual(values =  c("#00B0BE","#FFB255",  "#FF8CAE", 
                                 "#F47A00", "#98C127","#003A7D","#7E4794")) +
  labs(x = "",
       y = "lnRR")+
    theme(
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.8),
    legend.position = "none",
    # legend.position.inside = c(0.98, 0.05),
    # legend.justification = c(1, 0), 
    # legend.direction = "horizontal",
    legend.title = element_text(size = 6),
    legend.text = element_text(size = 6),
    axis.title = element_text(size = 10),
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size =6,angle = 0, vjust = 0.5, hjust = 1)
  )

SMDH_trait_plot <-
  orchard_plot(mod_results(
      SMDH_trait,
      mod = "trait_type",
      group = "paper_ID"
    ),
    trunk.size = 0.25,
    branch.size = 1,
    twig.size = 0.5,
    alpha = 0.2,
    xlab ="") +
  scale_colour_manual(values =  c("#00B0BE","#FFB255",  "#FF8CAE", 
                                 "#F47A00", "#98C127","#003A7D","#7E4794")) +
  scale_fill_manual(values =  c("#00B0BE","#FFB255",  "#FF8CAE", 
                                 "#F47A00", "#98C127","#003A7D","#7E4794")) +
  labs(x = "",
       y = "SMD(H)")+
    theme(
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.8),
    legend.position = "none",
    # legend.position.inside = c(0.98, 0.05),
    # legend.justification = c(1, 0), 
    # legend.direction = "horizontal",
    legend.title = element_text(size = 6),
    legend.text = element_text(size = 6),
    axis.title = element_text(size = 10),
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size =6, angle = 0, vjust = 0.5, hjust = 1)
  )
```

### Now we combine these plot

```{r combining meta-regression plots, no effect}

## I am going to remove the y ticks from the SMDH plots to make it smaller..
SMDH_parasite_plot <- SMDH_parasite_plot +
  theme(axis.text.y  = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank())

SMDH_birds_plot <- SMDH_birds_plot +
  theme(axis.text.y  = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank())

SMDH_trait_plot <- SMDH_trait_plot +
  theme(axis.text.y  = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank())

# Combine using patchwork
Figure_parasite_birds_trait <- ((lnRR_parasite_plot + SMDH_parasite_plot) / 
                                  (lnRR_birds_plot + SMDH_birds_plot) / 
                                  (lnRR_trait_plot + SMDH_trait_plot)) +
  plot_annotation(tag_levels = "A")

ggsave(here::here("figures/Figure_parasite_birds_trait.png"), plot = Figure_parasite_birds_trait, height = 250, width = 400, units = "mm", dpi = 600)

```

# Publication Bias Tests

```{r}

###---Small study Effects ----###

lnRR_small_study<-readRDS(here::here("model/lnRR_small_study.rds"))
SMDH_small_study<-readRDS(here::here("model/SMDH_small_study.rds"))

lnRR_small_study_bubble<-orchaRd::bubble_plot(mod_results(lnRR_small_study,
                     mod="sqrt_effective_n_inv",
                     group="paper_ID",
                     weights="prop"),
                     group="paper_ID") +
  scale_colour_manual(values ="mediumpurple4") +
  scale_fill_manual(values = "mediumpurple4") +
  labs(y = "Effect size (lnRR)",
       x = expression(sqrt(1 / N[effective])))+
    theme(
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.8),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 10),
    axis.title = element_text(size = 14),
    axis.text.x = element_text(size = 14),
    axis.text.y = element_text(angle = 0, vjust = 0.5, hjust = 1))

SMDH_small_study_bubble<-orchaRd::bubble_plot(mod_results(SMDH_small_study,
                     mod="sqrt_effective_n_inv",
                     group="paper_ID",
                     weights="prop"),
                     group="paper_ID") +
  scale_colour_manual(values ="mediumpurple4") +
  scale_fill_manual(values = "mediumpurple4") +
  labs(y = "Effect size (SMDH)",
       x = expression(sqrt(1 / N[effective])))+
    theme(
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.8),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 10),
    axis.title = element_text(size = 14),
    axis.text.x = element_text(size = 14),
    axis.text.y = element_text(angle = 0, vjust = 0.5, hjust = 1))

###---Decline Effects ----###

lnRR_decline_eff<-readRDS(here::here("model/lnRR_decline_eff.rds"))
SMDH_decline_eff<-readRDS(here::here("model/SMDH_decline_eff.rds"))

lnRR_decline_eff_bubble<-orchaRd::bubble_plot(mod_results(lnRR_decline_eff,
                     mod="year_MeanCenter",
                     group="paper_ID",
                     weights="prop"),
                     group="paper_ID")+
  scale_colour_manual(values ="#218833") +
  scale_fill_manual(values = "#218833") +
  labs(y = "Effect size (lnRR)",
       x = "Publication Year (Mean Cenetered)")+
    theme(
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.8),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 10),
    axis.title = element_text(size = 14),
    axis.text.x = element_text(size = 14),
    axis.text.y = element_text(angle = 0, vjust = 0.5, hjust = 1))

SMDH_decline_effect_bubble<-orchaRd::bubble_plot(mod_results(SMDH_decline_eff,
                     mod="year_MeanCenter",
                     group="paper_ID",
                     weights="prop"),
                     group="paper_ID")+
  scale_colour_manual(values ="#218833") +
  scale_fill_manual(values = "#218833") +
  labs(y = "Effect size (SMDH)",
       x = "Publication Year (Mean Cenetered)")+
    theme(
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.8),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 10),
    axis.title = element_text(size = 14),
    axis.text.x = element_text(size = 14),
    axis.text.y = element_text(angle = 0, vjust = 0.5, hjust = 1))

# Combine using patchwork
Figure_publication_bias <- ((lnRR_small_study_bubble + SMDH_small_study_bubble) /
                               (lnRR_decline_eff_bubble + SMDH_decline_effect_bubble)) +
  plot_annotation(tag_levels = "A")

ggsave(here::here("figures/Figure_publication_bias.png"), plot = Figure_publication_bias, height = 300, width = 400, units = "mm", dpi = 600)
```

# -------- TABLES ----------

## Table: Sensitivity analysis for intercept only model

```{r}
#| label: SA_intercept_only_table

pacman::p_load(gt,dplyr,metafor)

# Loading the models including flagged proxies
SA_lnRR_flagged<-readRDS(here::here("model/sensitivity_analysis/SA_lnRR_flagged.rds"))
SA_SMDH_flagged<-readRDS(here::here("model/sensitivity_analysis/SA_SMDH_flagged.rds"))

# Loading the models without data for those using 0 as ES because authors reported no effect
SA_lnRR_no_missingES<-readRDS(here::here("model/sensitivity_analysis/SA_lnRR_no_missingES.rds"))
SA_SMDH_no_missingES<-readRDS(here::here("model/sensitivity_analysis/SA_SMDH_no_missingES.rds"))

# Loading the models without data for those using 0 as ES because authors reported no effect and including flagged proxies
SA_lnRR_flagged_no_missingES<-readRDS(here::here("model/sensitivity_analysis/SA_lnRR_flagged_no_missingES.rds"))
SA_SMDH_flagged_no_missingES<-readRDS(here::here("model/sensitivity_analysis/SA_SMDH_flagged_no_missingES.rds"))

# Loading the model for SMDH without inferential statistics. There is none for lnRR because we can only uses means and deviation to estimate lnRR
SA_SMDH_no_inferential<-readRDS(here::here("model/sensitivity_analysis/SA_SMDH_no_inferential.rds"))

# Now let's build the table

SA_intercept_only <- tibble::tibble(
  effect_size = c("lnRR","lnRR","lnRR","SMDH","SMDH","SMDH","SMDH"),
  data_used = c("Including flagged proxies", "Without those cases where author reported no effect of the treatment", "Including flagged proxies and without the cases with no effect of the treatment", "Including flagged proxies", "Without those cases where author reported no effect of the treatment", "Including flagged proxies and without the cases with no effect of the treatment", "Without the cases where SMDH was calculated from inferential statistics"),

  Estimate = c(SA_lnRR_flagged$b[1],
               SA_lnRR_no_missingES$b[1],
               SA_lnRR_flagged_no_missingES$b[1],
               SA_SMDH_flagged$b[1],
               SA_SMDH_no_missingES$b[1],
               SA_SMDH_flagged_no_missingES$b[1],
               SA_SMDH_no_inferential$b[1]),

  CI_Lower = c(SA_lnRR_flagged$ci.lb,
               SA_lnRR_no_missingES$ci.lb,
               SA_lnRR_flagged_no_missingES$ci.lb,
               SA_SMDH_flagged$ci.lb,
               SA_SMDH_no_missingES$ci.lb,
               SA_SMDH_flagged_no_missingES$ci.lb,
               SA_SMDH_no_inferential$ci.lb),

  CI_Upper = c(SA_lnRR_flagged$ci.ub,
               SA_lnRR_no_missingES$ci.ub,
               SA_lnRR_flagged_no_missingES$ci.ub,
               SA_SMDH_flagged$ci.ub,
               SA_SMDH_no_missingES$ci.ub,
               SA_SMDH_flagged_no_missingES$ci.ub,
               SA_SMDH_no_inferential$ci.ub),

  PI_Lower = c(predict(SA_lnRR_flagged)$pi.lb,
               predict(SA_lnRR_no_missingES)$pi.lb,
               predict(SA_lnRR_flagged_no_missingES)$pi.lb,
               predict(SA_SMDH_flagged)$pi.lb,
               predict(SA_SMDH_no_missingES)$pi.lb,
               predict(SA_SMDH_flagged_no_missingES)$pi.lb,
               predict(SA_SMDH_no_inferential)$pi.lb),

  PI_Upper = c(predict(SA_lnRR_flagged)$pi.ub,
               predict(SA_lnRR_no_missingES)$pi.ub,
               predict(SA_lnRR_flagged_no_missingES)$pi.ub,
               predict(SA_SMDH_flagged)$pi.ub,
               predict(SA_SMDH_no_missingES)$pi.ub,
               predict(SA_SMDH_flagged_no_missingES)$pi.ub,
               predict(SA_SMDH_no_inferential)$pi.ub),

  P_value = c(SA_lnRR_flagged$pval,
              SA_lnRR_no_missingES$pval,
              SA_lnRR_flagged_no_missingES$pval,
              SA_SMDH_flagged$pval,
              SA_SMDH_no_missingES$pval,
              SA_SMDH_flagged_no_missingES$pval,
              SA_SMDH_no_inferential$pval),

  Q_value = c(SA_lnRR_flagged$QE,
              SA_lnRR_no_missingES$QE,
              SA_lnRR_flagged_no_missingES$QE,
              SA_SMDH_flagged$QE,
              SA_SMDH_no_missingES$QE,
              SA_SMDH_flagged_no_missingES$QE,
              SA_SMDH_no_inferential$QE),

  k_effect_sizes = c(SA_lnRR_flagged$k,
        SA_lnRR_no_missingES$k,
        SA_lnRR_flagged_no_missingES$k,
        SA_SMDH_flagged$k,
        SA_SMDH_no_missingES$k,
        SA_SMDH_flagged_no_missingES$k,
        SA_SMDH_no_inferential$k),

  n_studies = c(length(SA_lnRR_flagged$s.levels.f[[1]]),
                      length(SA_lnRR_no_missingES$s.levels.f[[1]]),
                      length(SA_lnRR_flagged_no_missingES$s.levels.f[[1]]),
                      length(SA_SMDH_flagged$s.levels.f[[1]]),
                      length(SA_SMDH_no_missingES$s.levels.f[[1]]),
                      length(SA_SMDH_flagged_no_missingES$s.levels.f[[1]]),
                      length(SA_SMDH_no_inferential$s.levels.f[[1]])
))

# Combine CI and PI into a single column each for compactness
SA_intercept_only <- SA_intercept_only %>%
  mutate(
    CI = paste0("[", sprintf("%.3f", CI_Lower), ", ", sprintf("%.3f", CI_Upper), "]"),
    PI = paste0("[", sprintf("%.3f", PI_Lower), ", ", sprintf("%.3f", PI_Upper), "]"),
    P_value = scales::pvalue(P_value),
    Q_value = sprintf("%.2f", Q_value)
  ) %>%
  select(
    effect_size,
    data_used,
    Estimate,
    CI,
    PI,
    P_value,
    Q_value,
    k_effect_sizes,
    n_studies
  )

# Let's make this table pretty
SA_intercept_only<-SA_intercept_only%>%
  gt()%>%
  tab_options(
    table.font.size = px(12),
    table.border.top.color = "black",
    table.border.bottom.color = "black",
    heading.title.font.size = px(14),
    heading.subtitle.font.size = px(12),
    column_labels.font.weight = "bold"
  ) %>%
  tab_header(
    title = md("**Summary of Sensitivity Analyses for lnRR and SMDH Intercept-only models**")
  ) %>%
  cols_label(
    effect_size = "Effect Size",
    data_used = "Data used in model",
    Estimate = "Estimate",
    CI = "95% CI",
    PI = "95% PI",
    P_value = "P-value",
    Q_value = "Q value",
    k_effect_sizes = "k (No. of Effect Sizes)",
    n_studies = "n (No. of Studies)"
  ) %>%
  fmt_number(
    columns = c(Estimate),
    decimals = 3
  ) %>%
  tab_style(
    style = cell_borders(
      sides = "bottom",
      color = "black",
      weight = px(1)
    ),
    locations = cells_body()
  ) %>%
  tab_options(
    table.font.size = "small",
    heading.title.font.size = 14,
    table.border.top.width = px(2),
    column_labels.font.weight = "bold",
    row_group.font.weight = "bold"
  ) %>%
  opt_table_outline()

#saving as a docx file
gtsave(SA_intercept_only,here::here("tables/SA_intercept_only.docx"))
gtsave(SA_intercept_only,here::here("tables/SA_intercept_only.html"))
```

## Table: Output from meta-regressions

```{r}

pacman::p_load(gt, tidyverse, metafor, purrr, stringr)

# Loading models for experimental design
lnRR_design<-readRDS(here::here("model/lnRR_design.rds"))
SMDH_design<-readRDS(here::here("model/SMDH_design.rds"))

# Loading models for parasite type
lnRR_parasite<-readRDS(here::here("model/lnRR_parasite.rds"))
SMDH_parasite<-readRDS(here::here("model/SMDH_parasite.rds"))

# Loading models for time of green nest material addition
lnRR_time_gnm_addition<-readRDS(here::here("model/lnRR_time_gnm_addition.rds"))
SMDH_time_gnm_addition<-readRDS(here::here("model/SMDH_time_gnm_addition.rds"))

# Loading models for the type of trait studied
lnRR_trait<-readRDS(here::here("model/lnRR_trait.rds"))
SMDH_trait<-readRDS(here::here("model/SMDH_trait.rds"))

# Loading models for the bird species studied
lnRR_birds<-readRDS(here::here("model/lnRR_birds.rds"))
SMDH_birds<-readRDS(here::here("model/SMDH_birds.rds"))
```

```{r}

# Let's source our custom function to extract values from the meta-regressions
source(here::here("functions/extractor_metareg.R"))

## For lnRR Values
lnRR_results <- extractor_metareg(lnRR_parasite,"parasite_type")%>%
  bind_rows(extractor_metareg(lnRR_birds,"bird_species"))%>%
  bind_rows(extractor_metareg(lnRR_trait,"trait_type"))%>%
  bind_rows(extractor_metareg(lnRR_design,"comparision_type"))%>%
  bind_rows(extractor_metareg(lnRR_time_gnm_addition,"time_of_gnm_addition"))%>%
  mutate(effect_size = "lnRR")

## For SMDH Values
SMDH_results <- extractor_metareg(SMDH_parasite,"parasite_type")%>%
  bind_rows(extractor_metareg(SMDH_birds,"bird_species"))%>%
  bind_rows(extractor_metareg(SMDH_trait,"trait_type"))%>%
  bind_rows(extractor_metareg(SMDH_design,"comparision_type"))%>%
  bind_rows(extractor_metareg(SMDH_time_gnm_addition,"time_of_gnm_addition"))%>%
  mutate(effect_size = "SMDH")

## Now let's write the table for publication:
# Combine both results
combined_results <- bind_rows(lnRR_results, SMDH_results)%>%
  pivot_wider(
    id_cols = c(moderator, level),
    names_from = effect_size,
    values_from = c(estimate, p_value, CI, PI, k, n),
    names_glue = "{.value}_{effect_size}"
  )%>% # Prepare row labels to mimic grouping
  arrange(moderator, level)%>%
  mutate(row_type = "level")

# Create moderator header rows for heterogeneity metrics
  moderator_header<-bind_rows(lnRR_results, SMDH_results) %>%
  select(moderator, effect_size, Q, Q_pval, R2_marginal) %>%
    distinct()%>%
    pivot_wider(
      names_from = effect_size,
      values_from = c(Q, Q_pval, R2_marginal),
      names_glue = "{.value}_{effect_size}"
    )%>%
  mutate(level = "Heterogeneity ",
         row_type = "moderator")

# Final combined table with headers and levels
combined_results <- bind_rows(moderator_header, combined_results) %>%
  arrange(moderator, row_type)%>% select(-row_type)%>%
  relocate(level, .after = moderator)%>%
  mutate(moderator=case_when(moderator=="parasite_type"~"Parasite Type",
                             moderator=="bird_species"~"Bird Species",
                             moderator=="trait_type"~"Trait Type Category",
                             moderator=="comparision_type"~"Experimental Design",
                             moderator=="time_of_gnm_addition"~"Time of Green Nest Material Addition"))%>%
  mutate(level = case_when(level==1 ~ "Non-aromatic vs. Aromatic",
                           level==2 ~ "No added material vs. Aromatic",
                           level==3 ~ "No added material vs. Non-aromatic",
                           level=="a" ~ "After egg hatching",
                           level=="b" ~ "Before egg hatching",
                           level=="c" ~ "Continously through nesting phase",
                           level=="Parasitic_and_pathogenic"~"Parasitic and Pathogen Related", .default=level))

# Optional: clean up or reorder columns for display
# Select and order columns for clarity
MLMR_table <- combined_results %>%
  gt(groupname_col = "moderator") %>%
  tab_header(title = md("**Multilevel Meta-analytic Models for Exploratory Analyses**")) %>%
  cols_label(
    Q_lnRR = "Q",
    Q_pval_lnRR = "Q p-val",
    R2_marginal_lnRR = "R²_mar",
    Q_SMDH = "Q",
    Q_pval_SMDH = "Q p-val",
    R2_marginal_SMDH = "R²_mar",
    level = "Level",
    estimate_lnRR = "Estimate",
    CI_lnRR = "95% CI",
    PI_lnRR = "95% PI",
    p_value_lnRR = "P-val",
    k_lnRR = "k",
    n_lnRR = "n",
    estimate_SMDH = "Estimate",
    CI_SMDH = "95% CI",
    PI_SMDH = "95% PI",
    p_value_SMDH = "P-val",
    k_SMDH = "k",
    n_SMDH = "n"
  ) %>%
   sub_missing(
    columns = everything(),
    missing_text = ""
  ) %>%
  opt_table_outline() %>%
  tab_options(
    table.font.size = "small",
    heading.title.font.size = 12,
    table.border.top.width = px(2),
    column_labels.font.weight = "medium",
    row_group.font.weight = "bold"
  )   %>%
  tab_style(
    style = cell_borders(
      sides = "bottom",
      color = "black",
      weight = px(1)
    ),
    locations = cells_body()
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(
      rows = level == "Heterogeneity"
    ))%>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(
      columns = c(p_value_lnRR),
      rows = p_value_lnRR < 0.05
    )
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(
      columns = c(p_value_SMDH),
      rows = p_value_SMDH < 0.05
    )
  )%>%
  tab_spanner(label = "lnRR", columns = contains("_lnRR")) %>%
  tab_spanner(label = "SMDH", columns = contains("_SMDH"))

#saving as a docx file
gtsave(MLMR_table,here::here("tables/EA_MLMR_table.docx"))
gtsave(MLMR_table,here::here("tables/EA_MLMR_table.html"))
```